---
title: "Eliciting causal models -- Analysis"
format: 
    pdf:
        fig-format: png
---

``` {r}
#| include: false
# Old yaml - put back for de-anonymised repo
# execute:
#   dev: svglite
# format: 
#     html:
#         code-fold: true
#         warning: false
#         message: false
#         fig-format: svg
```

``` {r library-load}
#options(device = svglite::svglite)

library(svglite)
library(tidyverse)
library(dagitty)
library(tidygraph)
library(scales)
library(patchwork)
library(ggsci)
library(ggokabeito)
library(ggdist)
# library(ggpubfigs)
#library(cowplot)
```

# Data import

``` {r loading-models}
# imports all causal models as DAGitty objects
# They are names dag_s1, ..., dag_s8 corresponding to the eight sessions
# A table of session data is also included
source("load-GCMs.R")
```

## Graphical Causal Models

Modelling sessions were originally run on Loopy, then transcribed into DAGitty (online) to generate the following `dot` language syntax for importing into `R`. A better way of doing this is under development. 


### Tidy graphs

Data imported as DAGitty objects, but need to be in tidygraph format for many calculations and visualisations.



``` {r}
# helper to get back door paths straight away - not working for all graphs yet
path_to_edges <- function(path) {
  tokens <- str_split(path, " ")[[1]]
  nodes  <- tokens[c(TRUE, FALSE)]
  arrows <- tokens[c(FALSE, TRUE)]
  
  tibble(
    from = ifelse(arrows == "->", nodes[-length(nodes)], nodes[-1]),
    to   = ifelse(arrows == "->", nodes[-1], nodes[-length(nodes)])
  )
}

open_backdoor_edgelist <- function(dag) {
  as_tibble(paths(backDoorGraph(dag), limit = 1e4)) |>
        filter(open) |> 
      pull(paths) |> 
      map_df(path_to_edges) |> 
      distinct()
}
```

Creating a list of all the DAGs in tidy format:

``` {r}
source("dagitty_tidygraph_translators.R")
# making list of dagitty dags (used later)
ddags <- list(
    "S1" = dag_s1, 
    "S2" = dag_s2,
    "S3" = dag_s3,
    "S4" = dag_s4,
    "S5" = dag_s5,
    "S6" = dag_s6,
    "S7" = dag_s7,
    "S8" = dag_s8
    )

tdag_s1 <- dag_to_tidy_graph(dag_s1)
tdag_s2 <- dag_to_tidy_graph(dag_s2)
tdag_s3 <- dag_to_tidy_graph(dag_s3)
tdag_s4 <- dag_to_tidy_graph(dag_s4)
tdag_s5 <- dag_to_tidy_graph(dag_s5)
tdag_s6 <- dag_to_tidy_graph(dag_s6)
tdag_s7 <- dag_to_tidy_graph(dag_s7)
tdag_s8 <- dag_to_tidy_graph(dag_s8)

tdags <- list(
    "S1" = tdag_s1, 
    "S2" = tdag_s2, 
    "S3" = tdag_s3, 
    "S4" = tdag_s4,
    "S5" = tdag_s5, 
    "S6" = tdag_s6, 
    "S7" = tdag_s7, 
    "S8" = tdag_s8)

quick_list_of_all_nodes <- unique(c(
    tdag_s1 |> pull(name), 
    tdag_s2 |> pull(name),
    tdag_s3 |> pull(name),
    tdag_s4 |> pull(name),
    tdag_s5 |> pull(name),
    tdag_s6 |> pull(name),
    tdag_s7 |> pull(name),
    tdag_s8 |> pull(name)
)
)
```

# Coding

## Node Classifiers

Coding the nodes into rough groups. There are two levels:

- `node_group` is the first aggregation into slightly coarser groups of nodes.
- `node_type` is an even more coarse grouping into a handful of types.

Also some nicer names of the fields:
- `node_nice_name` is a longer more explanatory name for plotting and putting in tables.
- `node_group_nice_name` a longer more explanatory name for node groups, as above.


``` {r}
name_node_nice_name <- tribble(
    ~name, ~nice_name,
    "CD", "Any system data",
    "CD.Att", "Attendance",
    "CD.LMS", "LMS activity",
    "CD.LMS.code", "Code artifacts",
    "CD.LMS.write", "Writing artifacts", 
    "CD.Str", "Course Data",
    "Grade", "Assessment grade",
    "Grade.Coms", "Mark - Communication",
    "Grade.Exam", "Mark - Exam",
    "Grade.Interp", "Mark - Interpretation",
    "Grade.Proc", "Mark - Learning process",
    "Grade.Quiz", "Mark - Quiz",
    "Grade.Ref", "Mark - Learning reflection",
    "Grade.Tech", "Mark - Technical",
    "H", "Help (general)",
    "H.AI", "Help from AI (any)",
    "H.AI.Agent", "Help from AI (agent)",
    "H.AI.evil", "Misuse of AI",
    "H.AI.evil.code", "Misuse of AI (code)",
    "H.AI.evil.write", "Misuse from AI (write)",
    "H.AI.Fb", "Help from AI (feedback)",
    "H.AI.good", "Help from AI (to learn)",
    "H.AI.Res", "Help from AI (research)",
    "H.AI.Study", "Help from AI (revision)",
    "H.Ment", "Help from mentor",
    "H.other.evil", "Cheating - non-AI",
    "H.Peer", "Help from peers",
    "H.Tut", "Help from teacher / tutor",
    "H.Tut.Fb", "Feedback from teacher / tutor",
    "LD", "Learning design",
    "LD.AIguide", "Guidance on AI use",
    "LD.Task", "Task design",
    "LD.Task.Code", "Coding task design",
    "LD.Task.Grp", "Groupwork task design",
    "LD.Task.Other", "Other task design",
    "LD.Task.Prac", "Practical task design",
    "LD.Task.Write", "Writing task design",
    "S.Act", "Student decisions / actions",
    "S.Act.AIstrat", "Student AI strategy",
    "S.Act.Fb", "Student acts on feedback",
    "S.Act.Fb.Post", "Student acts on feedback between subjects",
    "S.Act.Frm", "Student does formative tasks",
    "S.Act.Perf", "Student performance on task",
    "S.Act.Time", "Student commits time",
    "S.Aff.Integ", "Student acting with integrity",
    "S.Aff.Joy", "Student enjoyment",
    "S.Aff.MindSet", "Student mindset",
    "S.Aff.Safe", "Student feels safe",
    "S.Aff.WrkLd", "Student workload pressure",
    "S.Att.Eq", "Student equity group",
    "S.Att.HSk", "Student seeks help",
    "S.Att.Nro", "Student neuro-divergent",
    "S.Att.SocCap", "Student social capital",
    "S.Cp.Aca", "Student academic skill",
    "S.Cp.Aca.Post", "Student academic skill: Post grade",
    "S.Cp.EvJdg", "Student evaluative judgement",
    "S.Cp.Fb", "Student feedback literacy",
    "S.Cp.Grp", "Student groupwork capability",
    "S.Cp.Ind", "Student individual work capability",
    "S.Cp.Lrn", "Student metacognative learning capabilities",
    "S.Cp.SRL", "Student SRL",
    "S.Kn", "Student subject knowledge",
    "S.Kn.Post", "Student subject knowledge: Post grade",
    "S.Kn.Coms", "Student communications knowledge",
    "S.Kn.Interp", "Student interpretation knowledge",
    "S.Kn.Prior", "Student prior knowledge",
    "S.Kn.Tech", "Student technical knowledge",
    "T.Act.UseAI", "Teacher use of AI",
    "T.Aff.Integ", "Teacher acts with integrity",
    "T.Aff.WrkLd", "Teacher workload pressure",
    "T.Cp.Inst", "Teacher instruction quality",
    "T.Cp.KnStu", "Teacher knows students",
    "T.Cp.Mark", "Teacher reliably marks",
    "T.Kn.AI", "Teacher AI literacy",
    "T.Kn.Ass", "Teacher assessment literacy",
    "T.Kn.Sub", "Teacher subject knowledge"
)

node_group_nice_name <- tribble(
    ~node_group, ~nice_name,
    "Assessments", "Preliminary assessments", 
    "CD.Att", "Student attendance",
    "CD.LMS", "LMS activity",
    "CD.Str", "Course Data",
    "Grade.Outcome", "Assessment grade",
    "H.AI.evil", "Misuse of AI",
    "H.AI.good", "Help from AI",
    "H.Cheat", "Cheating - non-AI",
    "H.Peer", "Help from peers",
    "H.Tut", "Help from teacher / tutor",
    "LD", "Learning design",
    "S.Act", "Student decisions / actions",
    "S.Att", "Student attributes",
    "S.Cp", "Student metacognitive skill",
    "S.EM", "Student emotion / motivation",
    "S.Kn", "Student knowledge",
    "S.Kn.Prior", "Student prior knowledge",
    "T.Act", "Teacher use of AI",
    "T.Cp", "Teacher capabilities", 
    "T.EM", "Teacher emotion / motivation",
    "T.Kn", "Teacher knowledge"
)
```

### Coding node groups

``` {r node-group}
code_node_group <- function(v) {
    case_when(
        # Course data - Str, LMS --------------
        v == "CD" ~ "CD.Str", # was blending LMS and Str, but original graph was more about structure
        str_detect(v, "CD\\.LMS") ~ "CD.LMS",
        v == "CD.Att" ~ "CD.Att",
        v == "CD.Str" ~ "CD.Str",
        # Grade data --------------------------
        v == "Grade" ~ "Grade.Outcome",
        str_detect(v, "Grade\\.") ~ "Assessments",
        # Kinds of help -----------------------
        # - - - Help from teacher or tutor
        v == "H" ~ "H.Tut", # closest match for the only broad category one
        v == "H.Tut" ~ "H.Tut",
        v == "H.Ment" ~ "H.Tut",
        v == "H.Tut.Fb" ~ "H.Tut",
        # - - - Help from peers
        v == "H.Peer" ~ "H.Peer",
        # - - - Help from AI (good or bad usage)
        str_detect(v, "H\\.AI\\.evil") ~ "H.AI.evil",
        str_detect(v, "H\\.AI") ~ "H.AI.good",
        # - - - Other nefarious help (i.e. cheating)
        v == "H.other.evil" ~ "H.Cheat",
        # Learning Design --------------------
        str_detect(v, "^LD") ~ "LD",
        # Student activity -------------------
        str_detect(v, "^S\\.Act") ~ "S.Act",
        # Student attributes (generally unchangable) ---
        str_detect(v, "^S\\.Att") ~ "S.Att",
        # Student metacognitive capability (metacognition)
        str_detect(v, "^S\\.Cp") ~ "S.Cp",
        # Student emotion / motivation
        str_detect(v, "^S\\.Aff") ~ "S.EM",
        # Student knowledge (cognition)
        v == "S.Kn.Prior" ~ "S.Kn.Prior", # Prior knowledge treated as it's own node
        str_detect(v, "^S\\.Kn") ~ "S.Kn",
        # Teacher knowledge (cognition)
        str_detect(v, "^T\\.Kn") ~ "T.Kn",
        # Teacher capabilities (metacognition)
        str_detect(v, "^T\\.Cp") ~ "T.Cp",
        # Teacher emotion / motivation
        str_detect(v, "^T\\.Aff") ~ "T.EM",
        # Teacher actions / decisions
        str_detect(v, "^T\\.Act") ~ "T.Act",
        TRUE ~ v
        
    )
}
```

### Coding node types

``` {r node-type}
code_node_type <- function(n) {
    fl <- str_extract(n, "^.{1}")
    case_when(
        fl == "S" ~ "Student",
        fl == "T" ~ "Teacher",
        fl == "C" ~ "Learning data",
        fl == "H" ~ "Help",
        fl == "L" ~ "Learning design",
        fl == "G" ~ "Assessment data"
    ) |> 
        factor(levels = c("Assessment data", "Learning data", "Learning design", "Teacher", "Help", "Student"))
}

```

### Coding node measurability


From the node groups, describe how easy (or not) they are to measure. This is completed by someone familiar to the data warehouse at the institution:

- **DW** - _Data Warehouse_: Data is already collected and stored in the data warehouse of the institution.
- **O** - _Observable_: Node is not currently collected data, but would be possible to measure and collect with resourcing.
- **PO** - _Potentially Observable_:  Node is not measured, and would be difficult to measure directly. It is possible to find a direct proxy and measure that instead, such as a survey instrument.
- **L** - _Latent_: Node is not measured and would be difficult to even find a direct proxy for the latent construct. 


``` {r node-measurability}
code_node_measurability <- function(node) {
    case_when(
        # Student
        node == "S.Att.SocCap" ~ "PO", # check
        node == "S.Att.Nro" ~ "O", # check
        node == "S.Att.Eq" ~ "DW", # check
        node == "S.Att.HSk" ~ "L", # check
        node == "S.Act.AIstrat" ~ "L",
        node == "S.Act.Fb" ~ "L",
        node == "S.Act.Fb.Post" ~ "L",
        str_detect(node, "^S\\.Act") ~ "PO", 
        str_detect(node, "^S\\.Kn") ~ "PO", 
        str_detect(node, "^S") ~ "L", # Latent
        # Help
        node == "H.other.evil" ~ "L",
        node == "H.Peer" ~ "L",
        node == "H.AI.evil.code" ~ "PO", # Potentially observable
        node == "H.AI.evil.write" ~ "PO",
        node == "H.AI.evil.write" ~ "PO",
        str_detect(node, "^H\\.AI") ~ "PO",
        str_detect(node, "^H") ~ "L",
        # Assessment
        node == "Grade" ~ "DW",
        node == "Grade.Exam" ~ "DW",
        str_detect(node, "^Grade") ~ "O",
        # Learning data
        node == "CD.LMS" ~ "DW",
        str_detect(node, "^CD\\.LMS") ~ "O",
        node == "CD.Att" ~ "PO",
        node == "CD.Str" ~ "DW", # check
        node == "CD" ~ "PO", # check
        # Learning design
        str_detect(node, "^LD") ~ "PO",
        # Teacher
        node == "T.Act.UseAI" ~ "PO",
        node == "T.Aff.WrkLd" ~ "PO", # check this, not act with integ as L
        str_detect(node, "^T") ~ "L",
        node == "Assessments" ~ "DW"
    ) |> 
        factor(levels = c("DW", "O", "PO", "L"))
}
```

# Graph processing

## Graph helper functions

Convention is that these all start with `g_` and accept a `tidygraph` object as input. Functions outlined in the `graph_helpers.R` script. 

### Graph metrics functions

Generating a range of edge metrics - comparisons across models. Want to label edges in the unified DAG as 'contested' if they switch direction between some models. Also want to examine the same idea for a list of descendants. This is handled as a new table of edges - including all possible edges. 

``` {r fancy-edge-weight}
all_nodes <- 
    imap(tdags, \(g,m) g_nodelist(g)) |> 
    list_rbind(names_to = "Model")

all_edges <- 
    imap(tdags, \(g, m) g_edgelist(g) |> mutate(Model = m)) |> 
    list_rbind() |> 
    switches_exist(from, to) |> 
    mutate(edge_alignment = if_else(switched, "contested", "aligned")) |> 
    select(-switched)

all_possible_edges <- 
    imap(tdags, \(g, m) g_possible_edges(g)) |> 
    list_rbind(names_to = "Model")

max_possible_edges <- 
    all_possible_edges |> 
    filter(from != to) |> 
    count(from, to) |> 
    rename(max_n = n)

all_descendants <- 
    imap(tdags, \(g,m) g_all_descendants_edge_list(g) |> mutate(Model = m)) |> 
    list_rbind() |> 
    switches_exist(from, to) |> 
    mutate(descendant_alignment = if_else(switched, "descendant:contested", "descendant:aligned")) |> 
    select(-switched)

all_edges_with_metrics <- 
    all_possible_edges |> 
    filter(from != to) |> 
    left_join(all_edges |> mutate(in_graph = TRUE)) |> 
    mutate(in_graph = replace_na(in_graph, FALSE)) |> 
    left_join(all_descendants |> mutate(descendant_in_graph = TRUE)) |> 
    mutate(descendant_in_graph = replace_na(descendant_in_graph, FALSE))

all_edges_sum <- 
    all_edges_with_metrics |>
    group_by(from, to) |> 
    summarise(
        n = sum(in_graph),
        n_descendant = sum(descendant_in_graph),
        N = n(),
        p_direct_alignment = mean(edge_alignment == "aligned", na.rm = T),
        p_descendant_alignment = mean(descendant_alignment == "descendant:aligned", na.rm = T),
        .groups = "drop") |> 
    mutate(
        p_direct = n / N, 
        p_descendant = n_descendant / N) |> 
    arrange(desc(n)) 


all_edges_in_models_with_metrics <- 
    all_edges_sum |> 
    filter(n > 0) |> 
    inner_join(max_possible_edges) |> 
    mutate(
        edge_p_w = n / max_n,
        edge_dp_w = n_descendant / max_n) |> 
    mutate(edge_w = n)

```


## Creating merged graph



``` {r making-merged-fine-graph}
library(ggraph)

merged_edges <- 
    all_edges_in_models_with_metrics

merged_nodes <- 
    all_nodes |> 
    count(name) |> 
    rename(node_w = n) |> 
    mutate(node_group = code_node_group(name)) |> 
    mutate(node_type = code_node_type(name)) |> 
    mutate(measurable = code_node_measurability(name)) |> 
    left_join(name_node_nice_name, by = "name")

# cognitive, metacognitive, emotion / motivation, collaboration
merged_all_graph <- tbl_graph(nodes = merged_nodes, edges = merged_edges)
# View(merged_nodes |> arrange(name))
```

## Coarsen graphs

This is based on the coding of the nodes into node groups. 

Here we merge nodes into node groups before generating plot later. This has the potential to introduce loops depending on how the constructs were built, so we need to check the causal consistency between the fine-grained graph and the clustered / merged / higher-level graph. 

``` {r}
node_group_cluster_graph_from_merged <- 
    g_cluster_graph(merged_all_graph, code_node_group)

tcdags_node_group <- 
    map(tdags, \(g)g_cluster_graph(g, code_node_group)) 

node_group_cluster_graph_from_clusters <- 
    g_merge(
        g_merge(
            g_merge(tcdags_node_group[[1]], tcdags_node_group[[2]]),
            g_merge(tcdags_node_group[[3]], tcdags_node_group[[4]])
        ),
        g_merge(
            g_merge(tcdags_node_group[[5]], tcdags_node_group[[6]]),
            g_merge(tcdags_node_group[[7]], tcdags_node_group[[8]])
        )
    )

```

# Graph metrics

## Measurability, complexity

### Model level metrics

``` {r}
all_paths_summary <- 
    imap(ddags, \(d, n) d_count_paths(d)) |> 
    list_rbind(names_to = "Model") 

all_paths <- 
    imap(
        ddags,
        \(d, n) bind_cols(
            tibble(Model = n),
            dagitty::paths(d, limit = 5000)) |> 
            group_by(Model) |> 
            summarise(
                paths = list(paths),
                n_paths = n(),
                n_open = sum(open),
                p_open = mean(open)
            ))|> 
    list_rbind()


dag_complexity_metrics <- 
    imap(tdags, \(g, n)  as_tibble(g_summary_metrics(g))
    ) |> 
    list_rbind(names_to = "Model") |> 
    inner_join(
        imap(
            ddags, 
            \(d, n) 
            tibble(Model = n) |> 
                mutate(
                    n_adj_sets = length(dagitty::adjustmentSets(d)))) |> 
            list_rbind(), 
        by = "Model"
    ) |> 
    inner_join(all_paths_summary, by = "Model")
```

### Graph dissimilarity??


``` {r}
coarsened_tdags <- imap(
    tdags, 
    \(g,n)
    g_cluster_graph(g, code_node_group))

coarsest_tdags <- imap(
    tdags, 
    \(g,n)
    g_cluster_graph(g, code_node_type))

```

### Node group level metrics

``` {r}
all_nodes_with_metrics <- 
    imap(tdags, \(g,n) g_node_metrics(g)) |> 
    list_rbind(names_to = "Model")
```

## Joining metrics and session data

``` {r}
gcm_metrics <- 
    inner_join(session_data, dag_complexity_metrics, by = "Model") 
```
# Causal consistency

## Preservation of L1, L2

L1 (association) and L2 (intervention / do) levels on the causal hierarchy can be examined at a graph level through:

- $\mathcal{L}_1$ Examining all conditional probability implications. If these are preserved between graphs then the two graphs are $\mathcal{L}_1$-consistent.
- $\mathcal{L}_2$ Examining all 'do' operator implications. If these are preserved between graphs then the two graphs are $\mathcal{L}_2$-consistent. 
- $\mathcal{L}_3$ Is not explored - this would probably require a SCM, not just the DAG.

``` {r}
# building function to generate L1, L2, algebraic statements

nds <- tdag_s1 %N>% pull(name)
pairs <- tidyr::expand_grid(from = nds, to = nds) |> 
        filter(from != to)

# Getting conditional independencies
# This is a list structure - so better as that. X, Y and Z (Z is the set of vars)
L1_statements <- 
    ddags |> 
    imap(\(d,m) d_implied_conditional_independencies(d)) |> 
    list_rbind(names_to = "Model")

# Getting adjustments sets
pairs |> 
    rowwise() |> 
    mutate(
        adj_g = list(dagitty::adjustmentSets(dag_s1, exposure = from, outcome = to)),
        n_sets_g   = length(adj_g),
        identifiable = n_sets_g > 0
    )
```

# Visualisations

## Idea on layouts of graphs

- Sunflower layout: Each node group placed in $\phi$ ratio spiral out from centre, and each sub node spiraling out from these

## Session and other

#### Figure: Time spent with model and modelling utility

``` {r}
# denser option - radial arc bars for observable
library(scales)
library(ggforce)
library(patchwork)
library(cowplot)

hue1 <- palette.colors(3, palette = "Okabe-Ito")[[2]]
hue2 <- palette.colors(3, palette = "Okabe-Ito")[[3]]

# Main plot - hide all legends
p_validity_and_time <- gcm_metrics |> 
    mutate(
        x_duration = as.numeric(Modelling.duration, units = "minutes"),
        y_validity = GCM.validity
    ) |> 
    mutate(acyclic = if_else(acyclic, "Yes", "No")) |> 
    ggplot(aes(
        color = acyclic,
        size = n_adj_sets
    )) +
    geom_point(aes(
        x = x_duration,
        y = y_validity)) +
    scale_y_continuous(
        limits = c(1, 4.05),
        breaks = c(1, 2, 3, 4),
        labels = c("Trivially", "Simply", "Adequately", "Well"),
        name = "Representation of GCM") +
    scale_x_continuous(
        limits = c(33, 52),  # was 55 to  Keep original limits for the data
        breaks = c(35, 40, 45,50),
        name = "Time spent constructing model (minutes)") +
    scale_size(range = c(2, 7),
               breaks = range(gcm_metrics$n_adj_sets),
               name = "Adjustment\nsets (count)") +
    scale_color_manual(values = c(hue1, hue2), name = "Acyclic?") +
    scale_fill_manual(values = c(hue1, hue2), guide = "none") +
    theme_minimal() +
    # coord_fixed(ratio = 2
    #             # , clip = "off"
    #             ) +
    theme(
        legend.position = "left",  # Hide legends from main plot
        # panel.grid.major.x = element_blank(),  # Remove vertical gridlines
        panel.grid.minor = element_blank(),
        plot.margin = margin(10, 10, 10, 10)
    )


p_visibility <- gcm_metrics |>
    mutate(
        cp_dw = p_dw,
        cp_o = p_dw + p_o,
        cp_po = p_dw + p_o + p_po,
        cp_l = p_dw + p_o + p_po + p_l # should equal 1
    ) |> 
    mutate(GCM.validity = case_when(
        GCM.validity == 3 & acyclic ~ GCM.validity - 0.03,
        GCM.validity == 3 & !acyclic ~ GCM.validity + 0.03,
        GCM.validity == 2.5 & acyclic ~ GCM.validity - 0.03,
        GCM.validity == 2.5 & !acyclic ~ GCM.validity + 0.03,
        TRUE ~ GCM.validity
    )) |> 
    pivot_longer(c(starts_with("cp_"), starts_with("p_")), names_to = "node_vis", values_to = "proportion") |> 
    mutate(node_visibility = case_when(
        str_detect(node_vis, "p_dw") ~ "Data Warehouse",
        str_detect(node_vis, "p_o") ~ "Observable",
        str_detect(node_vis, "p_po") ~ "Potentially observable",
        str_detect(node_vis, "p_l") ~ "Latent"
    ) |> 
        ordered(levels = c(
            "Latent",
            "Potentially observable",
            "Observable",
            "Data Warehouse"
        ))) |> 
    mutate(prop_type = if_else(str_detect(node_vis, "^cp_"), "cp", "p")) |> 
    pivot_wider(names_from = prop_type, values_from = proportion, id_cols = c(Model,acyclic, GCM.validity, node_visibility)) |> 
    mutate(acyclic = if_else(acyclic, "Yes", "No")) |>  
    ggplot(aes(
        color = acyclic
    )) +
    geom_rect(aes(
        xmin = cp - p,
        xmax = cp,
        ymin = GCM.validity - 0.03,
        ymax = GCM.validity + 0.03,
        fill = acyclic,
        alpha = node_visibility),
        color = "white") +
    scale_y_continuous(
        limits = c(1, 4.05),
        breaks = c(1, 2, 3, 4),
        labels = c("Trivially", "Simply", "Adequately", "Well"),
        name = "Representation of GCM",
        position = "right") +
    scale_alpha_discrete(name = "Data visibility") +
    scale_fill_manual(values = c(hue1, hue2), guide = "none") +
    theme_minimal() +
    # coord_fixed(ratio = 0.15
    #             # , clip = "off"
    #             ) +
    xlab("Proportion of nodes") +
    theme(
        legend.position = "right",  # Hide legends from main plot
        # panel.grid.major.x = element_blank(),  # Remove vertical gridlines
        panel.grid.minor = element_blank(),
        plot.margin = margin(10, 10, 10, 10),
        axis.title.y = element_blank(),
        axis.text.y = element_blank()
    )


p_time_and_representation_1 <- p_validity_and_time + p_visibility
p_time_and_representation_1

ggsave("figures/FIG--time-modelling-and-representation.jpg", 
       plot = p_time_and_representation_1,
       width = 10, height = 5, 
       dpi = 300)

ggsave("figures/FIG--time-modelling-and-representation.svg", 
       plot = p_time_and_representation_1,
       width = 25, height = 10, units = "cm")
```



This version displays all in one (is a little dense)

``` {r}
# denser option - radial arc bars for observable
library(scales)
library(ggforce)
library(patchwork)
library(cowplot)

hue1 <- palette.colors(3, palette = "Okabe-Ito")[[2]]
hue2 <- palette.colors(3, palette = "Okabe-Ito")[[3]]

# Main plot - hide all legends
p_main <- gcm_metrics |> 
    mutate(
        x_duration = as.numeric(Modelling.duration, units = "minutes"),
        y_validity = GCM.validity * 10 + 20
    ) |> 
    mutate(acyclic = if_else(acyclic, "Yes, DAG", "No, contained cycles")) |> 
    ggplot(aes(
        color = acyclic,
        size = n_adj_sets
    )) +
    geom_arc_bar(aes(
        x0 = x_duration,
        y0 = y_validity,
        r0 = 0,
        r = 1.1,
        start = pi + 2 * (p_dw + p_o + p_po) * pi,
        end = 3 * pi
    ),
        fill = "white",
        linewidth = 0,
    alpha = 0.8, show.legend = FALSE) +
    geom_arc_bar(aes(
        x0 = x_duration, 
        y0 = y_validity,
        fill = acyclic,
        size = 1,
        r0 = 0,
        r = 1.5,
        start = pi, 
        end = pi + 2 * p_dw * pi
    ), 
        linewidth = 0,
    alpha = 1, show.legend = FALSE) + 
    geom_arc_bar(aes(
        x0 = x_duration, 
        y0 = y_validity,
        fill = acyclic,
        r0 = 0,
        r = 1.3,
        start = pi + 2 * p_dw * pi, 
        end = pi + 2 * (p_dw + p_o) * pi
    ), 
        linewidth = 0,
    alpha = 0.6, show.legend = FALSE) + 
    geom_arc_bar(aes(
        x0 = x_duration, 
        y0 = y_validity,
        fill = acyclic,
        r0 = 0,
        r = 1.1,
        start = pi + 2 * (p_dw + p_o) * pi, 
        end = pi + 2 * (p_dw + p_o + p_po) * pi
    ), 
        linewidth = 0,
    alpha = 0.4, show.legend = FALSE) + 
    geom_point(aes(
        x = x_duration,
        y = y_validity)) +
    scale_y_continuous(
        limits = c(30, 61),
        breaks = c(30, 40, 50, 60),
        labels = c("Trivially", "Simply", "Adequately", "Well"),
        name = "Representation of GCM") +
    scale_x_continuous(
        limits = c(30, 55),  # was 55 to  Keep original limits for the data
        breaks = c(30, 40, 50),
        name = "Time spent constructing model (minutes)") +
    scale_size(range = c(2, 7),
               breaks = range(gcm_metrics$n_adj_sets),
               name = "Adjustment\nsets") +
    scale_color_manual(values = c(hue1, hue2), name = "Is DAG?") +
    scale_fill_manual(values = c(hue1, hue2), guide = "none") +
    theme_minimal() +
    coord_fixed(ratio = 1, clip = "off") +
    theme(
        legend.position = "none",  # Hide legends from main plot
        # panel.grid.major.x = element_blank(),  # Remove vertical gridlines
        panel.grid.minor = element_blank(),
        plot.margin = margin(10, 10, 10, 10)
    )

# Create a dummy plot just to extract the legends
p_for_legend <- gcm_metrics |> 
    mutate(
        x_duration = as.numeric(Modelling.duration, units = "minutes"),
        y_validity = GCM.validity * 10 + 20
    ) |> 
    mutate(acyclic = if_else(acyclic, "Yes, DAG", "No, contained cycles")) |> 
    ggplot(aes(
        x = x_duration,
        y = y_validity,
        color = acyclic,
        size = n_adj_sets
    )) +
    geom_point() +
    scale_size(range = c(2, 7),
               breaks = range(gcm_metrics$n_adj_sets),
               name = "N adjustment sets") +
    scale_color_manual(values = c(hue1, hue2), name = "Is DAG?") +
    theme_minimal() +
    theme(legend.box = "vertical")

# Extract the legend
legend_extracted <- cowplot::get_legend(p_for_legend)

# Custom legend plot showing the arc explanation
p_legend_arcs <- ggplot() +
    geom_arc_bar(aes(x0 = 0, y0 = 0, r0 = 0.1, r = 0.8, start = pi, end = pi + pi/3),
             fill = "black", alpha = 1, linewidth = 0) +
    geom_arc_bar(aes(x0 = 0, y0 = 0, r0 = 0.1, r = 0.6, start = pi + pi/3, end = pi + (2*pi/3)),
             fill = "black", alpha = 0.6, linewidth = 0) +
    geom_arc_bar(aes(x0 = 0, y0 = 0, r0 = 0.1, r = 0.4, start = pi + (2*pi/3), end = 2 * pi),
             fill = "black", alpha = 0.4, linewidth = 0) +
    annotate("text", 
             x = 0.5, y = -0.9,
             label = "In data\nwarehouse", 
             hjust = 0, size = 3, color = "black") +
    annotate("text", 
             x = 0.5, y = 0, 
             label = "Observable", 
             hjust = 0, size = 3, color = adjustcolor("black", alpha.f = 0.8)) +
    annotate("text", x = 0.5, y = 0.9, label = "Potentially\nobservable", 
             hjust = 0, size = 3, color = adjustcolor("black", alpha.f = 0.6)) +
    annotate("segment", x = 0.1, xend = 0.4, y = -0.5, yend = -0.6, color = "black") +
    annotate("segment", x = 0, xend = 0.4, y = 0, yend = 0, 
             color = "black", alpha = 0.6) +
    annotate("segment", x = 0.1, xend = 0.4, y = 0.4, yend = 0.6, 
             color = "black", alpha = 0.4) +
    coord_fixed() +
    xlim(-1, 2.5) +
    ylim(-1.3, 1.3) +
    labs(title = "Proportion of nodes:") +
    theme_void() +
    theme(plot.title = element_text(size = 9, face = "bold"))

# Combine the arc legend and extracted legends vertically
legend_column <- cowplot::plot_grid(
    legend_extracted,
    p_legend_arcs,
    ncol = 2,
    rel_heights = c(1, 1)
)

# Combine main plot with legend column
p_time_and_representation <- cowplot::plot_grid(
    p_main,
    legend_column,
    ncol = 2,
    rel_widths = c(1, 1)  # Adjust ratio as needed
)

p_time_and_representation

ggsave("figures/FIG--time-modelling-and-representation-arc-bars.jpg", 
       plot = p_time_and_representation,
       width = 10, height = 5, 
       dpi = 900)
```



As a table:

``` {r}
gcm_metrics |> 
    mutate(acyclic = if_else(acyclic, "Yes, DAG", "No, contained cycles")) |> 
    arrange(desc(Modelling.duration)) |> 
    select(Modelling.duration, GCM.validity, acyclic, n_adj_sets, p_dw, p_o, p_po, p_l)
```

#### Figure (OPTIONAL): Time spent with model and modelling utility - including graph similarity

Graph similarity computed with d-Measure. 

Only showing the upper 50% of edges 



``` {r}
d_between_models_median <- median(pull(d_edges, edge_d))

print(str_c("Median d-measure between models is ", round(d_between_models_median, 3)))

d_edges |> 
    slice(1)

d_edges |> 
    arrange(edge_w) |> 
    slice(1)


d_between_models_by_time_and_utility <- 
    d_edges |> 
    inner_join(
        gcm_metrics |> 
            mutate(from = Model) |> 
            select(from, xfrom = Modelling.duration, yfrom = GCM.validity)
    ) |> 
    inner_join(
        gcm_metrics |> 
            mutate(to = Model) |> 
            select(to, xto = Modelling.duration, yto = GCM.validity)
    ) 

p_time_and_representaion_with_distance <- 
    gcm_metrics |> 
    mutate(acyclic = if_else(acyclic, "Yes, DAG", "No, contained cycles")) |> 
    ggplot() +
    geom_segment(
        data = d_between_models_by_time_and_utility |> 
            filter(edge_d < d_between_models_median),
        aes(x = xfrom, xend = xto, y = yfrom, yend = yto,
            alpha = edge_w), size = 0.5, color = "black") +
    geom_point(data = gcm_metrics, 
               aes(x = Modelling.duration, y = GCM.validity, size = n_adj_sets, colour = acyclic)) +
    # geom_text(aes(label = Model), nudge_y = -.1) +
    scale_y_continuous(limits = c(1, 4),
                       breaks = c(1:4),
                       labels = c("Trivially", "Simply", "Adequately", "Well"),
                       name = "Representation of GCM") +
     scale_x_time(labels = label_time(format = "%M:%S"),
                  name = "Time spent constructing model (minutes)") +
    scale_size(range = c(2,7),
               breaks = range(gcm_metrics$n_adj_sets),
               # labels = c("None", )
               name = "N adjustment sets") +
    scale_alpha_continuous(range = c(0.1, 0.7), guide = "none") +
    scale_color_npg(name = "Is DAG?") +
    theme_minimal() +
    theme(legend.position = "bottom",
          legend.direction = "vertical",
          panel.grid.major = element_line(color = "grey85", linewidth = 0.2),
          panel.grid.minor = element_line(color = "grey92", linewidth = 0.1)
    )

p_time_and_representaion_with_distance
# ggsave("figures/FIG--time-modelling-and-representation-with-distances.jpg", 
#        width = 5, height = 4, 
#        dpi = 900)
```


## DAG networks

### Visualising model dissimilarities

``` {r}
tbl_graph(
    nodes = dag_complexity_metrics |> rename(name = Model),
    edges = d_edges) |> 
    activate("edges") |> 
    filter(edge_d < 0.33) |> 
    ggraph(layout = "fr") +
    geom_edge_fan(aes(alpha = edge_w)) +
    geom_node_point(aes(size = density, color = factor(acyclic))) +
    geom_node_text(aes(label = name), repel = T) +
    theme_graph() +
    ggtitle("Graph dissimilarities")
```

Same again, but hierarchical clustering

``` {r}
hc <- hclust(d_matrix |> select(-from) |> as.dist())
plot(hc)
```


### Edge coherence


``` {r}
tb_edge_coherence <- 
    edges_with_metrics |> 
    summarise(
        n = sum(n),
        n_desc = sum(n_desc),
        max_n = sum(max_n),
        aligned = mean(switched == "aligned"),
        contested = mean(switched != "aligned")
    ) |> 
    mutate(edges_included = n / max_n,
           desc_incllude = n_desc / max_n)
tb_edge_coherence
```

Generally when two nodes in the model were selected by different participants the same edge was drawn in `r round(100*tb_edge_coherence$edges_included , 2)`% of the time. A handful of potential edges (`r round(100*(1-tb_edge_coherence$aligned),2)`%) were included, but asigned a different direction of causal flow. 
In `r round(100*tb_edge_coherence$desc_incllude , 2)`% of the cases where a path $A \rightarrow B$ was included in a model, then there existed a path from $A$ to $B$ in every other model that included those two nodes, just sometimes through a mediator(s) variable. 

### Merging graphs

``` {r plotting-merged-graph}

merged_all_graph_layout <- create_layout(merged_all_graph, layout = "sugiyama")
merged_all_graph_layout_adjusted <- 
    merged_all_graph_layout |> 
    mutate(
        # x = if_else(name == "S.Post.Fb",x + 20,x),
        y = if_else(name == "S.Act.Fb.Post",y + 4,y)
           ) |> 
    mutate(
        x = if_else(name == "S.Act",x + 20,x),
        y = if_else(name == "S.Act",y - 4,y)
           ) 
ptemp <- merged_all_graph_layout_adjusted |> 
    ggraph() +
        geom_edge_fan(aes(alpha = edge_w), strength = 0.2,
            arrow = arrow(length = unit(2, 'mm'), type = "closed"),
            start_cap = circle(3, 'mm'),
            end_cap = circle(3, 'mm')) +
        geom_node_point(aes(colour = node_type, size = node_w)) +
        # geom_node_text(aes(label = name, size = node_w, alpha = node_w), nudge_y = -.5) +
        theme_graph() +
        scale_edge_alpha_continuous(range = c(0.15,0.8)) +
        scale_alpha_continuous(range = c(0.3,1)) +
        theme(legend.position = "none")

b <- ggplot_build(ptemp)
x_limits_fine <- b$layout$panel_params[[1]]$x.range
y_limits_fine <- b$layout$panel_params[[1]]$y.range



plot_fine_graph_from_layout <- function(lyt, weight_nodes = TRUE) {
    ppp <- 
        lyt |> 
        ggraph() 
    
    if (weight_nodes) {
        pp <- ppp +
            geom_node_point(aes(colour = node_type,
                                size = node_w)) +
            scale_size(name = "point-size") +
            geom_edge_fan(aes(
                alpha = edge_p_w * edge_w + edge_w, 
                edge_linetype = factor(switched)), 
                strength = 0.2,
                arrow = arrow(length = unit(1, 'mm'),
                              type = "closed"),
                start_cap = circle(3, 'mm'),
                end_cap = circle(3, 'mm'))
    } else {
        pp <- ppp +
            geom_node_point(aes(colour = node_type)) +
            geom_edge_fan(strength = 0.2, alpha = 0.5,
                          arrow = arrow(length = unit(1, 'mm'),
                                        type = "closed"),
                          start_cap = circle(1, 'mm'),
                          end_cap = circle(1, 'mm'))
            
    }
    pp +
        scale_colour_okabe_ito() +
        # geom_node_text(aes(label = name, size = node_w, alpha = node_w), nudge_y = -.5) +
        theme_graph() +
        scale_edge_alpha_continuous(range = c(0.15,0.8)) +
        scale_alpha_continuous(range = c(0.3,1)) +
        theme(legend.position = "none",
              plot.margin = margin(t = 0, r = 0, b = 0, l = 0, unit = "pt")) +
        xlim(x_limits_fine) +
        ylim(y_limits_fine)
}

g_plot_fine <- function(g, lyt = merged_all_graph_layout_adjusted, weight_nodes = TRUE) {
    create_layout(g, layout = "sugiyama") |> 
        left_join(lyt |> select(x,y,name,node_group,node_type,node_w), by = "name",
                  suffix = c(".default", "")) |> 
        plot_fine_graph_from_layout(weight_nodes = weight_nodes)
}

pMergedFineLegend <- merged_all_graph_layout_adjusted |> 
    g_plot_fine() + 
    theme(legend.position = "bottom") +
    guides(size = "none", edge_width = "none", alpha = "none", edge_alpha = "none",
           color = guide_legend(nrow = 1)) +
    theme(legend.title = element_blank())
pMergedFineLegend

pMergedFineLegendTall <- merged_all_graph_layout_adjusted |> 
    g_plot_fine() + 
    theme(legend.position = "right") +
    guides(size = "none", alpha = "none", edge_alpha = "none", edge_width = "none",
           color = guide_legend(ncol = 2)) +
    theme(legend.title = element_blank())
pMergedFine <- pMergedFineLegend +
    theme(legend.position = "none")

pMergedFineLabelled <- pMergedFine + 
    geom_node_text(
        aes(label = nice_name, size = node_w, colour = node_type), alpha = 0.6, 
        data = merged_all_graph_layout_adjusted |>  filter(node_w > 0), repel = TRUE, max.overlaps = 4) +
    ggnewscale::new_scale("size") +
    scale_size(range = c(3,6), name = "text-size")
pMergedFineLabelled

# Extract legend
leg <- cowplot::get_legend(pMergedFineLegend)
# Turn it into a ggplot object for saving
legplot <- cowplot::ggdraw(leg)
# Save just the legend
ggsave("dag_legend.jpg", legplot, width = 6, height = 1)

leg_tall <- cowplot::get_legend(pMergedFineLegendTall)
leg_tall_plot <- cowplot::ggdraw(leg_tall)


# ggsave("dag_merged_fine.svg", pMergedFine, width = 6, height = 4)
# ggsave("dag_merged_fine_labelled.svg", pMergedFineLabelled, width = 6, height = 4)
```


``` {r}
p1 <- g_plot_fine(tdag_s1, weight_nodes = FALSE)
p2 <- g_plot_fine(tdag_s2, weight_nodes = FALSE)
p3 <- g_plot_fine(tdag_s3, weight_nodes = FALSE)
p4 <- g_plot_fine(tdag_s4, weight_nodes = FALSE)
p5 <- g_plot_fine(tdag_s5, weight_nodes = FALSE)
p6 <- g_plot_fine(tdag_s6, weight_nodes = FALSE)
p7 <- g_plot_fine(tdag_s7, weight_nodes = FALSE)
p8 <- g_plot_fine(tdag_s8, weight_nodes = FALSE)

# Plot border 
    # theme(
    #   plot.background = element_rect(
    #     colour = "grey70", 
    #     fill = NA,
    #     linewidth = 1
    #   )
    # )

hl_text_size <- 1.5
hl_text_alpha <- 0.7

p2peerHighlight <- p2 + 
    geom_node_point(size = 3, shape = 1, alpha = 0.6,
    data = p2$data |>  filter(name == "H.Peer")) +
    geom_node_point(size = 5, shape = 1, alpha = 0.3,
    data = p2$data |>  filter(name == "H.Peer")) +
    geom_node_point(size = 7, shape = 1, alpha = 0.1,
    data = p2$data |>  filter(name == "H.Peer")) +
    geom_node_text(aes(label = nice_name, colour = node_type), 
                   size = hl_text_size , alpha = hl_text_alpha,
    data = p2$data |>  filter(name == "H.Peer") |> 
        inner_join(name_node_nice_name, by = "name"), repel = T) 
p3peerHighlight <- p3 + 
    geom_node_point(size = 3, shape = 1, alpha = 0.6,
    data = p3$data |>  filter(name == "H.Peer")) +
    geom_node_point(size = 5, shape = 1, alpha = 0.3,
    data = p3$data |>  filter(name == "H.Peer")) +
    geom_node_point(size = 7, shape = 1, alpha = 0.1,
    data = p3$data |>  filter(name == "H.Peer")) +
    geom_node_text(aes(label = nice_name, colour = node_type), 
                   size = hl_text_size , alpha = hl_text_alpha,
    data = p3$data |>  filter(name == "H.Peer") |> 
        inner_join(name_node_nice_name, by = "name"), repel = T) 
p4peerHighlight <- p4 +
    geom_node_point(size = 3, shape = 1, alpha = 0.6,
    data = p4$data |>  filter(name == "H.Peer")) +
    geom_node_point(size = 5, shape = 1, alpha = 0.3,
    data = p4$data |>  filter(name == "H.Peer")) +
    geom_node_point(size = 7, shape = 1, alpha = 0.1,
    data = p4$data |>  filter(name == "H.Peer")) +
    geom_node_text(aes(label = nice_name, colour = node_type), 
                   size = hl_text_size , alpha = hl_text_alpha,
    data = p4$data |>  filter(name == "H.Peer") |> 
        inner_join(name_node_nice_name, by = "name"), repel = T) 

p7KnHighlight <- p7 +
    geom_node_point(size = 3, shape = 5, alpha = 0.6,
    data = p7$data |>  filter(str_detect(name, "Kn"))) +
    geom_node_point(size = 5, shape = 5, alpha = 0.3,
    data = p7$data |>  filter(str_detect(name, "Kn"))) +
    geom_node_point(size = 7, shape = 5, alpha = 0.1,
    data = p7$data |>  filter(str_detect(name, "Kn"))) +
    geom_node_text(aes(label = nice_name, colour = node_type), 
                   size = hl_text_size , alpha = hl_text_alpha,
    data = p7$data |>  filter(str_detect(name, "Kn")) |> 
        inner_join(name_node_nice_name, by = "name"), repel = T) 

p5KnHighlight <- p5 +
    geom_node_point(size = 3, shape = 5, alpha = 0.6,
    data = p5$data |>  filter(str_detect(name, "Kn"))) +
    geom_node_point(size = 5, shape = 5, alpha = 0.3,
    data = p5$data |>  filter(str_detect(name, "Kn"))) +
    geom_node_point(size = 7, shape = 5, alpha = 0.1,
    data = p5$data |>  filter(str_detect(name, "Kn"))) +
    geom_node_text(aes(label = nice_name, colour = node_type), 
                   size = hl_text_size , alpha = hl_text_alpha,
    data = p5$data |>  filter(str_detect(name, "Kn")) |> 
        inner_join(name_node_nice_name, by = "name"), repel = T) 

p6KnHighlight <- p6 +
    geom_node_point(size = 3, shape = 5, alpha = 0.6,
    data = p6$data |>  filter(str_detect(name, "Kn"))) +
    geom_node_point(size = 5, shape = 5, alpha = 0.3,
    data = p6$data |>  filter(str_detect(name, "Kn"))) +
    geom_node_point(size = 7, shape = 5, alpha = 0.1,
    data = p6$data |>  filter(str_detect(name, "Kn"))) +
    geom_node_text(aes(label = nice_name, colour = node_type), 
                   size = hl_text_size , alpha = hl_text_alpha,
    data = p6$data |>  filter(str_detect(name, "Kn")) |> 
        inner_join(name_node_nice_name, by = "name"), repel = T) 
# 

```

``` {r}
pAllFine <- (p1 | p2 | p3 | p4) / (p5 | p6 | p7 | p8)
pAllFine
# ggsave("dags_All_fine.svg", pAllFine, width = 18, height = 9)
```

#### Figure: Model graphs, all and merged

``` {r}
library(grid)
tag_plot <- function(p, tag, x = 0, y = 0.9, fontsize = 8) {
  p + annotation_custom(
    grid::textGrob(tag, x = x, y = y, 
                   hjust = 0, vjust = 1, gp = gpar(fontsize = fontsize))
  )
}

pAll3by3 <- 
    (
        tag_plot(p2peerHighlight,"a")| 
            tag_plot(p8, "b") | 
            tag_plot(p7KnHighlight, "c")) / 
    (
        tag_plot(p4peerHighlight, "d") | 
            tag_plot(p1, "e") | 
            tag_plot(p5KnHighlight, "f")) / 
    (
        tag_plot(p3peerHighlight, "g") | 
            leg_tall_plot | 
            tag_plot(p6KnHighlight, "h")) 
     

pFineTogetherLabelled <- pAll3by3 / tag_plot(pMergedFineLabelled, "i") + plot_layout(heights = c(1,1,1,3))
pFineTogether <- pAll3by3 / pMergedFine + plot_layout(heights = c(1,1,1,3))
pFineTogetherLabelled

# ggsave("dags_fine_all_and_merged.png", pFineTogether, width = 9, height = 10, dpi = 900)
# ggsave("dags_fine_all_and_merged.svg", pFineTogether, width = 9, height = 10)
ggsave("figures/FIG--dags_fine_all_and_merged_labelled.jpg", pFineTogetherLabelled, width = 9, height = 10, dpi = 900)
# ggsave("dags_fine_all_and_merged_labelled.svg", pFineTogetherLabelled, width = 9, height = 10)

print("a, d, g are models S2, S3, S4; b, e are S8 and S1; c, f, h are S7, S5, S6. Closest models are therefore d and g (D-measure 0.105) and the furthest are e and 7 (D-measure 0.324)")
```

#### Common patterns - fine

Most common three-node patterns. 

``` {r}
merged_all_graph |> 
    activate("nodes") |> 
    filter(node_w > 3) |> 
    activate("edges") |>
    filter(edge_w > 1) |>
    g_plot_fine() +
    geom_node_text(aes(label = nice_name, colour = node_type), 
                   alpha = hl_text_alpha, repel = TRUE)

```

### Figure - all graphs, single highlight

``` {r}
plot_highlighted_fine_graph <- function(model = 1, LEGEND_POSITION = "none") {
    MODEL <- paste0("s", model)
    nodes_in_graph <- all_nodes |> 
        filter(Model == MODEL) |> 
        pull(name)
    edges_in_graph <- all_edges |> 
        filter(Model == MODEL) |> 
        select(from, to)
    
    edges_in_graph_indexed <- 
        edges_in_graph |> 
        mutate(
            from = match(from, merged_all_graph %N>% as_tibble() |> pull(name)),
            to = match(to, merged_all_graph %N>% as_tibble() |> pull(name))
        )
    
    g_highlight <- merged_all_graph %N>%
        mutate(
            node_highlight = if_else(
                name %in% nodes_in_graph, 2, 1
            )) %E>%
        left_join(
            edges_in_graph_indexed |> 
                mutate(edge_highlight = 2)
        ) |> 
        mutate(edge_highlight = replace_na(edge_highlight, 1)) |> 
        create_layout("sugiyama")
    
    g_highlight |>
        ggraph() +
        geom_edge_fan(aes(
            # edge_linetype = factor(switched),
            alpha = edge_highlight), 
            strength = 0.2,
            arrow = arrow(length = unit(1, 'mm'),
                          type = "closed"),
            start_cap = circle(3, 'mm'),
            end_cap = circle(3, 'mm')) +
        geom_node_label(
            aes(label = nice_name, 
                color = node_type,
                size = node_w),
            alpha = 0.95,
            data = g_highlight |> 
                filter(node_highlight == 2),
            repel = T, max.overlaps = 30,
            show.legend = FALSE) +
        geom_node_point(aes(colour = node_type,
                            size = node_w,
                            alpha = node_highlight)) +
        scale_size(name = "point-size", range = c(2,5)) +
        scale_color_okabe_ito(order = c(9, 7,1, 2,3, 6, 8,5, 4)) +
        # scale_color_npg() +
        theme_graph() +
        scale_edge_alpha_continuous(range = c(0.08,0.8)) +
        scale_alpha_continuous(range = c(0.1,1.0)) +
        # theme(legend.position = "none",
        #       plot.margin = margin(t = 0, r = 0, b = 0, l = 0, unit = "pt")) +
        theme(legend.position = LEGEND_POSITION) +
        guides(size = "none", edge_width = "none", alpha = "none", edge_alpha = "none",
               color = guide_legend(ncol = 1)) +
        theme(legend.title = element_blank())
}

plot_highlighted_fine_graph(4)
# +
#     xlim(x_limits_fine) +
#     ylim(y_limits_fine)



plot_highlighted_coarse_graph <- function(model = 1, LEGEND_POSITION = "none") {
    MODEL <- paste0("s", model)
    node_groups_in_graph <- all_nodes |> 
        filter(Model == MODEL) |> 
        mutate(name = code_node_group(name)) |> 
        distinct(name) |> 
        pull(name) 
    edges_in_coarse_graph <- all_edges |> 
        filter(Model == MODEL) |> 
        select(from, to) |> 
        mutate(across(from:to, code_node_group))
    
    edges_in_graph_indexed <- 
        edges_in_coarse_graph |> 
        mutate(
            from = match(from, merged_coarse_models %N>% as_tibble() |> pull(name)),
            to = match(to, merged_coarse_models %N>% as_tibble() |> pull(name))
        )
    
    edge_integers <- 
        edges_in_graph_indexed |> 
        mutate(p2p3 = (2 ** from) * (3 ** to) ) |> 
        pull(p2p3)
    
    node_types <- merged_all_graph %N>%
        as_tibble() |> 
        distinct(node_group, node_type) 
    
    g_highlight <- merged_coarse_models %N>%
        inner_join(node_types |> rename(name = node_group),
                   by = "name") |>
        mutate(
            node_highlight = if_else(
                name %in% node_groups_in_graph, 2, 1
            )) %E>%
        mutate(edge_highlight = if_else(
            ((2 ** from )*(3 ** to)) %in% edge_integers, 2, 1
        )) |> 
        create_layout("sugiyama")
    
    g_highlight |>
        ggraph() +
        geom_edge_fan(aes(
            # edge_linetype = factor(switched),
            alpha = edge_highlight), 
            strength = 0.2,
            arrow = arrow(length = unit(1, 'mm'),
                          type = "closed"),
            start_cap = circle(3, 'mm'),
            end_cap = circle(3, 'mm')) +
        geom_node_label(
            aes(label = nice_name, 
                color = node_type,
                size = node_w),
            alpha = 0.95,
            data = g_highlight |> 
                filter(node_highlight == 2),
            repel = T, max.overlaps = 30,
            show.legend = FALSE) +
        geom_node_point(aes(colour = node_type,
                            size = node_w,
                            alpha = node_highlight)) +
        scale_size(name = "point-size", range = c(2,5)) +
        # scale_color_npg() +
        scale_color_okabe_ito(order = c(9, 7,1, 2,3, 6, 8,5, 4)) +
        theme_graph() +
        scale_edge_alpha_continuous(range = c(0.08,0.8)) +
        scale_alpha_continuous(range = c(0.1,1.0)) +
        # theme(legend.position = "none",
        #       plot.margin = margin(t = 0, r = 0, b = 0, l = 0, unit = "pt")) +
        theme(legend.position = LEGEND_POSITION) +
        guides(size = "none", edge_width = "none", alpha = "none", edge_alpha = "none",
               label = "none",
               color = guide_legend(ncol = 1)) +
        theme(legend.title = element_blank())
}

# Model used in paper
p_merged_highlighted_fine <- plot_highlighted_fine_graph(5, LEGEND_POSITION = "right")
p_merged_highlighted_coarse <- plot_highlighted_coarse_graph(5, LEGEND_POSITION = "right")
p_merged_highlighted_fine + ggtitle("Model used in paper (5)")
p_merged_highlighted_coarse 
ggsave(plot = p_merged_highlighted_fine, 
       filename = "figures/FIG--merged-fine-with-1-highlighted.jpg",
       width = 8, height = 5, dpi = 600)

ggsave(plot = p_merged_highlighted_coarse, 
       filename = "figures/FIG--merged-coarse-with-1-highlighted.jpg",
       width = 8, height = 5, dpi = 600)
```

``` {r}
# Generating all
p_fine_show_and_save <- function(n){
    p_temp <- plot_highlighted_fine_graph(n, LEGEND_POSITION = "right") + ggtitle(paste0("Model ", n, " highlighted"))
    ggsave(plot = p_temp, 
           filename = paste0("figures/plt--merged-fine-with-", n, "-highlighted.png"),
           width = 8, height = 5, dpi = 600)
    return(p_temp)
}
p_coarse_show_and_save <- function(n){
    p_temp <- plot_highlighted_coarse_graph(n, LEGEND_POSITION = "right") + ggtitle(paste0("Coarsened model ", n, " highlighted"))
    ggsave(plot = p_temp, 
           filename = paste0("figures/plt--merged-coarse-with-", n, "-highlighted.png"),
           width = 8, height = 5, dpi = 600)
    return(p_temp)
}
```

Highlighting all models in turn
``` {r render-all-highlighted-fine-models}
p_fine_show_and_save(1)
p_fine_show_and_save(2)
p_fine_show_and_save(3)
p_fine_show_and_save(4)
p_fine_show_and_save(5)
p_fine_show_and_save(6)
p_fine_show_and_save(7)
p_fine_show_and_save(8)
```

Highlighting all coarsened models in turn
``` {r render-all-highlighted-coarse-models}
p_coarse_show_and_save(1)
p_coarse_show_and_save(2)
p_coarse_show_and_save(3)
p_coarse_show_and_save(4)
p_coarse_show_and_save(5)
p_coarse_show_and_save(6)
p_coarse_show_and_save(7)
p_coarse_show_and_save(8)
```




### Visualising coarsened graphs

``` {r plotting-coarse-graph}
merged_coarse_layout <- 
    merged_all_graph |> 
    g_coarsen_dag_by_ng() |> 
    create_layout(layout = "sugiyama") |> 
    mutate(
        x = case_when(
            name == "CD.Att" ~ 18.5,
            name == "T.Aff.WrkLd" ~ x - 5,
            name == "H.AI.good" ~ 27,
            TRUE ~ x),
        y = case_when(
            name == "CD.Att" ~ y + 1,
            name == "Grade" ~ y - 2.5,
            name == "CD.Str" ~ 3,
            name == "S.EM" ~ 5,
            TRUE ~ y)
           )

p_temp <- merged_coarse_layout |> 
     ggraph() +
        geom_edge_fan(aes(alpha = edge_w), strength = 0.2,
            arrow = arrow(length = unit(2, 'mm')),
            start_cap = circle(3, 'mm'),
            end_cap = circle(3, 'mm')) +
        geom_node_point(aes(colour = node_type, size = node_w)) +
    scale_color_okabe_ito(order = c(9, 7,1, 2,3, 6, 8,5, 4)) +
        # scale_color_npg(drop = FALSE) +
        # geom_node_text(aes(label = name, size = node_w, alpha = node_w), nudge_y = -.5) +
        theme_graph() +
        scale_edge_alpha_continuous(range = c(0.15,0.8)) +
        scale_alpha_continuous(range = c(0.2,0.6)) +
        theme(legend.position = "none") 

bc <- ggplot_build(p_temp)
x_limits_course <- bc$layout$panel_params[[1]]$x.range
y_limits_course <- bc$layout$panel_params[[1]]$y.range

coarse_plot_from_layout <- function(lyt) {
    lyt |> 
        ggraph() +
        geom_edge_fan(aes(alpha = edge_w), strength = 0.2,
            arrow = arrow(length = unit(2, 'mm')),
            start_cap = circle(3, 'mm'),
            end_cap = circle(3, 'mm')) +
        geom_node_point(aes(colour = node_type, size = node_w)) +
        scale_color_okabe_ito(order = c(9, 7,1, 2,3, 6, 8,5, 4), drop = FALSE) +
        # ggsci::scale_color_npg(drop = FALSE) +
        # geom_node_text(aes(label = name, size = node_w, alpha = node_w), nudge_y = -.5) +
        theme_graph() +
        scale_edge_alpha_continuous(range = c(0.15,0.8)) +
        scale_alpha_continuous(range = c(0.2,0.6)) +
        theme(legend.position = "none") +
        xlim(x_limits_course) +
        ylim(y_limits_course)
}

g_plot_coarse <- function(g, lyt = merged_coarse_layout) {
    create_layout(g, layout = "sugiyama") |> 
        left_join(lyt |> select(x,y,name, node_type,node_w), by = "name",
                  suffix = c(".default", "")) |> 
    coarse_plot_from_layout()    
}

pCoarseMerged <- merged_coarse_layout |> 
    coarse_plot_from_layout()
# ggsave("dag_merged_course.svg", pCoarseMerged, , width = 6, height = 6)
```


Note the appearance of feedback loops - these are between S.Cp and S.Kn, but also S.Cp and H.AI.evil. These metacognitive skills are clearly seen as important feedback processes in this system.

``` {r}
pCoarseMerged + geom_node_text(aes(label = name, size = node_w, alpha = node_w), nudge_x = 1)
```

``` {r}
p1course <- g_coarsen_dag_by_ng(tdag_s1) |> g_plot_coarse()
p2course <- g_coarsen_dag_by_ng(tdag_s2) |> g_plot_coarse()
p3course <- g_coarsen_dag_by_ng(tdag_s3) |> g_plot_coarse()
p4course <- g_coarsen_dag_by_ng(tdag_s4) |> g_plot_coarse()
p5course <- g_coarsen_dag_by_ng(tdag_s5) |> g_plot_coarse()
p6course <- g_coarsen_dag_by_ng(tdag_s6) |> g_plot_coarse()
p7course <- g_coarsen_dag_by_ng(tdag_s7) |> g_plot_coarse()
p8course <- g_coarsen_dag_by_ng(tdag_s8) |> g_plot_coarse()

pAllcourse <- (p1course | p2course | p3course | p4course) / (p5course | p6course | p7course | p8course)
# ggsave("dags_all_course.svg", pAllcourse, width = 18, height = 9)

pAllcourse
```

``` {r}
# all node group edges
all_edges |> 
    mutate(
        from_g = code_node_group(from), 
        to_g = code_node_group(to)) |> 
    filter(to_g == "LD") |> 
    filter(from_g == "CD.LMS")
```

#### Figure: Consensus DAG

Most common node group patterns - nodes in at least half the models, then looking at the resulting graph of node-groups.

``` {r}
node_list_majority_of_models <- 
    all_nodes |> 
    count(name) |> 
    filter(n >= 4) |> 
    pull(name)

coarse_filtered_graph_temp <- 
    merged_all_graph |> 
    activate("nodes") |> 
    filter(name %in% node_list_majority_of_models)

coarse_filtered_graph <- tbl_graph(
    nodes = coarse_filtered_graph_temp %N>%
        as_tibble() |> 
        mutate(name = code_node_group(name),
               node_type = code_node_type(name)) ,
    edges = coarse_filtered_graph_temp %E>%
        as_tibble() %>%
        mutate(
            from = coarse_filtered_graph_temp %N>% pull(name) %>% .[from],
            to   = coarse_filtered_graph_temp %N>% pull(name) %>% .[to]
        ) %>%
        select(from, to, edge_w) |> 
        mutate(across(from:to, code_node_group))) |> 
    activate(edges) |> 
    mutate(
        switched = map_lgl(row_number(), function(i) {
            from_i <- .E()$from[i]
            to_i   <- .E()$to[i]
            any(from == to_i & to == from_i)
        })
    ) 

```

``` {r}
# Old - not used
lyt_top_node_groups <- create_layout(coarse_filtered_graph, layout = "sugiyama") |> 
    mutate(
        x = case_when(
            nice_name == "Student subject knowledge" ~ 0.5,
            nice_name == "Assessment grade" ~ 0.5,
            nice_name == "Help from teacher / tutor" ~ 1,
            nice_name == "Task design" ~ 0.5,
            TRUE ~ x
        ),
        y = case_when(
            nice_name == "Student subject knowledge" ~ 2.5,
            # nice_name == "Assessment grade" ~ 0.5,
            nice_name == "LMS activity" ~ 1,
            nice_name == "Task design" ~ 4,
            TRUE ~ y
        )
    )

p_top_node_groups <- ggraph(lyt_top_node_groups) +
    geom_edge_fan(aes(edge_linetype = factor(switched), alpha = edge_w, ), # alpha = edge_w, 
                  strength = 0.8,
                  arrow = arrow(length = unit(2, 'mm'), type = "closed"),
                  start_cap = circle(3, 'mm'),
                  end_cap = circle(3, 'mm')) +
    geom_node_point(aes(colour = node_type, size = node_w)) +
        ggsci::scale_color_npg(drop = FALSE) +
        # geom_node_text(aes(label = name, size = node_w, alpha = node_w), nudge_y = -.5) +
        theme_graph() +
        scale_edge_alpha_continuous(range = c(0.15,0.8)) +
        scale_alpha_continuous(range = c(0.2,0.9)) +
        theme(legend.position = "none") +
    geom_node_text(aes(label = nice_name, colour = node_type), 
                   alpha = 0.9, nudge_y = -.3, repel = T)

# p_top_node_groups
# ggsave("figures/FIG--consensus-DAG.png", p_top_node_groups, dpi = 600, height = 4, width = 6)
```

Weighting by num of models instead
This one uses only nodes that are in the majority of models, then coarsens the graph, then aggregates edges based on the number of models that they appear in. 
The edges must also appear in at least 2 models

``` {r}
node_list_majority_of_models <- 
    all_nodes |> 
    count(name) |> 
    filter(n >= 4) |> 
    pull(name)

coarse_filtered_graph_ng_filled <- 
    tbl_graph( # only nodes that are in most models (4 or more)
        nodes = all_nodes |> 
            filter(name %in% node_list_majority_of_models) |> 
            mutate( # but then coarsen into node groups
                node_group = code_node_group(name),
                   node_type = code_node_type(name)) |> 
            count(node_group, node_type) |> 
            rename(node_w = n),
        edges =     all_edges |> # Only edges between node groups...
            mutate(across(from:to, code_node_group)) |>
            distinct(Model, from, to) |> 
            count(from, to) |>
            # filter(n > 1) |> # ...that occur in at least two models
            filter(from %in% code_node_group(node_list_majority_of_models)) |>
            filter(to %in% code_node_group(node_list_majority_of_models)) |>
            rename(edge_w = n)) |> 
    activate("edges") |> 
    mutate(
        switched = map_lgl(row_number(), function(i) {
            from_i <- .E()$from[i]
            to_i   <- .E()$to[i]
            any(from == to_i & to == from_i)
        })
    ) |> 
    activate("nodes") |> 
    inner_join(node_group_nice_name)
```

Filtered - edges must be in at least 2 models

``` {r}
lyt_coarse_filtered_ng_filtered <- create_layout(
    coarse_filtered_graph_ng_filled |> 
        activate("edges") |> 
        filter(edge_w > 1) |> 
        mutate(
            switched = map_lgl(row_number(), function(i) {
                from_i <- .E()$from[i]
                to_i   <- .E()$to[i]
                any(from == to_i & to == from_i)
            })
        ) , 
    layout = "sugiyama")

p_top_node_groups_filtered <- ggraph(lyt_coarse_filtered_ng_filtered) +
    geom_edge_fan(aes(edge_linetype = factor(switched), alpha = edge_w, ), # alpha = edge_w, 
                  strength = 0.8,
                  arrow = arrow(length = unit(2, 'mm'), type = "closed"),
                  start_cap = circle(3, 'mm'),
                  end_cap = circle(3, 'mm')) +
    geom_node_point(aes(colour = node_type, size = node_w)) +
    scale_color_okabe_ito(order = c(9, 7,1, 2,3, 6, 8,5, 4), drop = FALSE) +
    # ggsci::scale_color_npg(drop = FALSE) +
    # geom_node_text(aes(label = name, size = node_w, alpha = node_w), nudge_y = -.5) +
    theme_graph() +
        scale_edge_alpha_continuous(range = c(0.15,0.8)) +
        scale_alpha_continuous(range = c(0.2,0.9)) +
        theme(legend.position = "none") +
    scale_size(range = c(3,6)) +
    geom_node_text(aes(label = nice_name, colour = node_type), 
                   alpha = 0.9, nudge_y = -.13, repel = F) +
    xlim(c(min(lyt_coarse_filtered_ng_filtered$x) - 0.15, max(lyt_coarse_filtered_ng_filtered$x) + 0.15))

p_top_node_groups_filtered

ggsave("figures/FIG--consensus-sub-model-filtered.jpg", 
       p_top_node_groups_filtered, 
       dpi = 900, 
       height = 4, width = 6)
```

All edges between node groups from the majority nodes

``` {r}
lyt_coarse_filtered_ng_full <- create_layout(
    coarse_filtered_graph_ng_filled, 
    layout = "sugiyama") |> 
    select(-x, -y) |> 
    inner_join(lyt_coarse_filtered_ng_filtered |> select(x, y, node_group))

p_top_node_groups_full <- ggraph(lyt_coarse_filtered_ng_full) +
    geom_edge_fan(aes(edge_linetype = factor(switched), alpha = edge_w, ), # alpha = edge_w, 
                  strength = 0.8,
                  arrow = arrow(length = unit(2, 'mm'), type = "closed"),
                  start_cap = circle(3, 'mm'),
                  end_cap = circle(3, 'mm')) +
    geom_node_point(aes(colour = node_type, size = node_w)) +
    scale_color_okabe_ito(order = c(9, 7,1, 2,3, 6, 8,5, 4), drop = FALSE) +
        # ggsci::scale_color_npg(drop = FALSE) +
        # geom_node_text(aes(label = name, size = node_w, alpha = node_w), nudge_y = -.5) +
        theme_graph() +
        scale_edge_alpha_continuous(range = c(0.15,0.8)) +
        scale_alpha_continuous(range = c(0.2,0.9)) +
        theme(legend.position = "none") +
    scale_size(range = c(3,6)) +
    geom_node_text(aes(label = nice_name, colour = node_type), 
                   alpha = 0.9, nudge_y = -.13, repel = F) +
    xlim(c(
        min(lyt_coarse_filtered_ng_full$x) - 0.15, 
        max(lyt_coarse_filtered_ng_full$x) + 0.15))

p_top_node_groups_full

ggsave("figures/FIG--consensus-sub-model-full.jpg",
       p_top_node_groups_full,
       dpi = 900,
       height = 4, width = 6)
```

Node groups close to the constructs:
- Belonging
- Mindset
- Collaboration

``` {r}
# all nodes
all_nodes_with_metrics |> 
    inner_join(name_node_nice_name) |> 
    inner_join(node_group_nice_name |> rename(g_nice_name = nice_name)) |> 
    select(name, nice_name, g_nice_name, node_group, Model, pagerank_no_outcome, pagerank, btw_rel_rank) |> 
    arrange(name, Model) |> 
    knitr::kable()

```






## Node importance and measurability

``` {r}
all_node_groups_aggregated_metrics <- 
    all_nodes_with_metrics |> 
    group_by(Model, node_group, node_type) |> 
    summarise(
        p_traced = mean(measurability == "DW"),
        measurability = (3 * mean(measurability == "DW") +
                             2 * mean(measurability == "O") +
                             mean(measurability == "PO")) / 3,
        pagerank = sum(pagerank),
              pagerank_no_outcome = sum(pagerank_no_outcome),
              btw_rel = sum(btw_rel)) |> 
    group_by(node_group, node_type) |> 
    mutate(
        n = n(),
        overall_pagerank = sum(pagerank / 8),
           overall_pagerank_no_outcome = sum(pagerank_no_outcome / 8),
           overall_btw = sum(btw_rel)) |> 
    ungroup()
```

#### Figure: Node importance and data availability

``` {r}
d_ng_measurability_importance <- 
    all_node_groups_aggregated_metrics |> 
    filter(node_group != "Grade.Outcome") |>
    inner_join(node_group_nice_name, by = "node_group") |>
    mutate(node_group = nice_name) |>
    group_by(node_group) |>
    mutate(m_group = mean(measurability)) |>
    ungroup() |> 
    mutate(node_group = fct_reorder(node_group, pagerank_no_outcome)) |> 
    arrange(desc(node_group)) |> 
    filter(n > 3) # need at least half the models with node group


p_top_node_pagerank_no_outcome_measurability <- 
    d_ng_measurability_importance |> 
    # arrange(Model, node_group) |> 
    ggplot(aes(x = pagerank_no_outcome, y = node_group)) +
    geom_boxplot(aes(fill = m_group), colour = "lightgray", size = 0.4, alpha = 0.7) +
    # geom_line(aes(x = pagerank_no_outcome, y = node_group, group = Model), alpha = 0.6) +
    geom_point(aes(colour = measurability), alpha = 0.9) +
    scale_color_gradient(
        low = "darkslateblue", 
        # mid = "darkturquoise",
        high = "aquamarine",
        # midpoint = mean(range(d_ng_measurability_importance$measurability)),
        limits = c(0,1),
        breaks = c(0, 1/3, 2/3, 1),
        labels = c("Latent","Potentially\nobservable", "Observable", "Data\nWarehouse"),
        name = "",
        guide = guide_colorbar(reverse = FALSE)
    ) +
    scale_fill_gradient(
        # midpoint = mean(range(d_ng_measurability_importance$m_group)),
        limits = c(0,1),
                low = "darkslateblue", 
        # mid = "darkturquoise",
        high = "aquamarine"
        ) +
    theme_minimal() +
    ylab("Node group") +
    xlab("Node groups combined Centrality (Pagerank)") +
    # geom_rug(aes(colour = m_dist_to_trace), alpha = 0.6, sides = "b") +
    # coord_cartesian(clip = "off") +
    guides(fill = "none") +
    theme(legend.position = "bottom",
          legend.justification = "center",
    legend.box.just = "center",
          legend.text  = element_text(size = 8),
          legend.key.width = unit(1, "cm"),
          axis.ticks.y = element_blank(),
          axis.line.y = element_blank(),
          panel.grid.major.y = element_blank(),
          panel.grid.minor.y = element_blank())

# p_top_node_pagerank_no_outcome_measurability

# extract legend
leg <- get_legend(p_top_node_pagerank_no_outcome_measurability)

# arrange with relative heights (legend gets a small slice)
p_node_importance_with_measurability <- (wrap_elements(leg) /
    (p_top_node_pagerank_no_outcome_measurability + theme(legend.position = "none"))) +
  plot_layout(heights = c(0.2, 1))

# p_node_importance_with_measurability
# ggsave("figures/FIG--node_pagerank_no_outcome_measurability.png", p_node_importance_with_measurability,
#        width = 5,
#        height = 4,
#        dpi = 900)

```

With discrete scale instead

``` {r}
m_pal <- c("darkslateblue",
    "dodgerblue","orange","deeppink")
    

p_top_node_pagerank_no_outcome_measurability_discrete <- 
    d_ng_measurability_importance |>
    mutate(
        measurability = case_when(
            measurability < (1/6) ~ "Latent", 
            measurability < 0.5 ~ "Potentially observable",
            measurability < (5/6) ~ "Observable",
            TRUE ~ "Data warehouse") |> 
            factor(levels = c("Latent", "Potentially observable", "Observable", "Data warehouse"))
    ) |> 
    arrange(Model, node_group) |>
    ggplot(aes(x = pagerank_no_outcome, y = node_group)) +
    geom_boxplot(aes(fill = m_group), colour = "lightgray", size = 0.4, alpha = 0.7) +
    # geom_line(aes(x = pagerank_no_outcome, y = node_group, group = Model), alpha = 0.6) +
    geom_point(aes(colour = measurability), alpha = 0.9) +
    scale_color_manual(values = m_pal, # or retro four
                       labels = c("Latent","Potentially\nobservable", "Observable", "Data\nWarehouse"),
                       name = ""
    ) +
    scale_fill_gradientn(
        colours = m_pal,
        breaks = c(0, 1/3, 2/3, 5/6)
        ) +
        theme_minimal() +
            ylab("Node group") +
            xlab("Node groups combined Centrality (Pagerank)") +
            # geom_rug(aes(colour = m_dist_to_trace), alpha = 0.6, sides = "b") +
            # coord_cartesian(clip = "off") +
            guides(fill = "none",
                   color = guide_legend(nrow = 2, byrow = FALSE)) +
            theme(legend.position = "bottom",
          legend.direction = "vertical",
                  legend.justification = "left",
                  legend.text  = element_text(size = 8),
                  legend.key.width = unit(1, "cm"),
                  axis.ticks.y = element_blank(),
                  axis.line.y = element_blank(),
                  panel.grid.major.y = element_blank(),
          panel.grid.minor.y = element_blank())
    
p_top_node_pagerank_no_outcome_measurability_discrete

# 
# # extract legend
# leg <- get_legend(p_top_node_pagerank_no_outcome_measurability_discrete)
# 
# # arrange with relative heights (legend gets a small slice)
# p_node_importance_with_measurability <- (wrap_elements(leg) /
#     (p_top_node_pagerank_no_outcome_measurability_discrete + theme(legend.position = "none"))) +
#   plot_layout(heights = c(0.2, 1))


ggsave("figures/FIG--node_pagerank_no_outcome_measurability.jpg",
       p_top_node_pagerank_no_outcome_measurability_discrete, 
       width = 5, 
       height = 4, 
       dpi = 900)
```

As table

``` {r}
all_nodes_with_metrics |> 
    distinct(name, measurability) |> 
    janitor::tabyl(measurability) |> 
    janitor::adorn_totals("row")
```


``` {r}
# Pagerank no outcome - dist to traced version - old

d_ng_trace_importance <- 
    all_node_groups_aggregated_metrics |> 
    inner_join(all_node_groups_distance_to_traced, by = c("Model", "node_group")) |> 
    filter(node_group != "Grade.Outcome") |>
    inner_join(node_group_nice_name, by = "node_group") |>
    mutate(node_group = nice_name) |>
    group_by(node_group) |>
    mutate(p_traced_group = mean(m_dist_to_trace)) |>
    ungroup() |> 
    mutate(node_group = fct_reorder(node_group, pagerank_no_outcome)) |> 
    arrange(desc(node_group)) |> 
    filter(n > 3) # need at least half the models with node group


p_top_node_pagerank_no_outcome_trace_dist <- 
    d_ng_trace_importance |> 
    # arrange(Model, node_group) |> 
    ggplot(aes(x = pagerank_no_outcome, y = node_group)) +
    geom_boxplot(aes(fill = p_traced_group), colour = "lightgray", size = 0.4, alpha = 0.7) +
    # geom_line(aes(x = pagerank_no_outcome, y = node_group, group = Model), alpha = 0.6) +
    geom_point(aes(colour = m_dist_to_trace), alpha = 0.9) +
    scale_color_gradient2(
        low = "darkslateblue", 
        mid = "coral",
        high = "deeppink",
        midpoint = mean(range(d_ng_trace_importance$m_dist_to_trace)),
                         breaks = range(d_ng_trace_importance$m_dist_to_trace),
                         labels = c("Observed","Latent"),
                         name = "",
        guide = guide_colorbar(reverse = TRUE)
        ) +
    scale_fill_gradient2(
        midpoint = mean(range(d_ng_trace_importance$m_dist_to_trace)),
        low = "darkslateblue", 
        mid = "coral",
        high = "deeppink"
        ) +
    theme_minimal() +
    ylab("Node group") +
    xlab("Node groups combined Centrality (Pagerank)") +
    # geom_rug(aes(colour = m_dist_to_trace), alpha = 0.6, sides = "b") +
    # coord_cartesian(clip = "off") +
    guides(fill = "none") +
    theme(legend.position = "bottom",
          axis.ticks.y = element_blank(),
          axis.line.y = element_blank(),
          panel.grid.major.y = element_blank(),
          panel.grid.minor.y = element_blank())

p_top_node_pagerank_no_outcome_trace_dist

ggsave("figures/FIG--node_pagerank_no_outcome_trace_distance.jpg", p_top_node_pagerank_no_outcome_trace_dist, 
       width = 5, 
       height = 4, 
       dpi = 900)

d_ng_trace_importance |> 
    group_by(node_group) |> 
    summarise(m_dist_to_trace = mean(m_dist_to_trace)) |> 
    arrange(desc(m_dist_to_trace)) 
    
```

Above plot but lines joining models that are like?



#### Table: Node lists, with types, groups and importance metrics

``` {r appendix-node-groups-table}
# could add distance to traced
node_groups_table <- 
    all_nodes_with_metrics |>
    inner_join(name_node_nice_name, by = "name") |> 
    mutate(name = if_else(
        node_group == "H.AI.good" & measurability == "Traced",
        str_c(nice_name, "*"),
        nice_name
    )) |> 
    inner_join(node_group_nice_name |> rename(ng_nice_name = nice_name), by = "node_group") |> 
    mutate(node_group = ng_nice_name) |> 
    group_by(node_type, node_group) |>
    mutate(measurability = case_when(
        all(measurability == "Traced") ~ "Traced",
        all(measurability == "Untraced") ~ "Untraced",
        TRUE ~ str_c("Some traced* (", sum(measurability == "Traced"), " of ", n(), ")")
    )) |> 
    group_by(node_type, node_group, measurability) |>
    summarise(
        nodes = str_c(unique(name), collapse = "; "),
        n = sum(node_w),
        pagerank = round(sum(pagerank_no_outcome) / 8, 2)
        # ,btw = round(sum(btw_rel), 3)
    ) |> 
    mutate(pagerank = if_else(node_group == "Grade.Outcome", NA, pagerank)) |> 
    # mutate(node_type = case_when(
    #     node_type == "L" ~ "Learning Design",
    #     node_type == "S" ~ "Student",
    #     node_type == "T" ~ "Teacher",
    #     node_type == "C" ~ "Course Data",
    #     node_type == "G" ~ "Grade Data",
    #     node_type == "H" ~ "Help"
    # )) |> 
    arrange(desc(n), desc(pagerank)) |> 
    select(node_group, node_type, nodes, n, measurability, pagerank) |> 
    ungroup()

node_groups_table
```

``` {r}
all_nodes_with_metrics |> 
    distinct(name) |> 
    nrow() |> 
    paste("distinct nodes")

all_nodes_with_metrics |> 
    distinct(node_group) |> 
    nrow() |> 
    paste("node groups")

all_nodes_with_metrics |> 
    distinct(name, measurability) |> count(measurability)
```

``` {r}
#| eval: false
tb1 <- node_groups_table |> filter(n >= 7) |> arrange(desc(n))
tb2 <- node_groups_table |> filter(n < 7) |> arrange(desc(n))
tb1 |> clipr::write_clip()
tb2 |> clipr::write_clip()

```

What about grouping by the node category first, so it's hierarchical?

``` {r}
node_types_table <- 
    node_groups_table |> 
    group_by(node_type) |> 
    mutate(nt = sum(n)) |> 
    ungroup() |> 
    arrange(desc(nt), desc(n)) |> 
    select(node_type, nt, node_group, nodes, n, measurability, pagerank) |> 
    mutate(node_type = str_c(node_type, " (", nt, ")")) |> 
    select(-nt)

node_types_table
```

``` {r}
#| eval: false
tbt1 <- node_types_table |> filter(str_detect(node_type, "Student|Help"))
tbt2 <- node_types_table |> filter(!str_detect(node_type, "Student|Help"))
tbt1 |> clipr::write_clip()
tbt2 |> clipr::write_clip()

```



## Edge / path importance



``` {r}
all_edge_group_with_metrics <-
    edges_with_metrics |> 
    mutate(from = code_node_group(from),
           to = code_node_group(to)) |> 
    group_by(from, to) |> 
    summarise(
        n = sum(n),
        max_n = sum(max_n),
        edge_p_w = mean(edge_p_w),
        p_aligned = mean(switched == "aligned")
    )

edge_group_analysis_edgelist <- all_edge_group_with_metrics |> 
    mutate(edge_w = n + edge_p_w * n) |> 
    arrange(desc(edge_w))


# the node group one here isn't one row per node group, and doesn't have the node type!
edge_group_analysis_graph <- tbl_graph(
    nodes = all_node_groups_aggregated_metrics |> 
        rename(name = node_group) |> 
        group_by(name, node_type) |> 
        summarise(
            p_traced = mean(p_traced),
            pgrnk = mean(pagerank_no_outcome),
            node_w = sum(n)
        ) , 
    edges = edge_group_analysis_edgelist)

edge_group_analysis_graph %N>%
    arrange(name) |> 
    # activate(edges) |>
    # filter(edge_w > 8) |>
    ggraph(
        layout = "circle"
        # layout = "sugiyama"
        ) +
            geom_edge_diagonal(
                aes(alpha = edge_w), 
                          strength = 0.5,
                          arrow = arrow(length = unit(3, 'mm'),
                                        type = "closed"),
                          start_cap = circle(2, 'mm'),
                          end_cap = circle(3, 'mm')) +
    geom_node_point(aes(size = pgrnk, color = node_type)) +
    geom_node_text(aes(label = name), repel = TRUE) +
    scale_edge_alpha(range = c(0.1,9)) +
    theme_graph()
```

Path importance - what about the sets of implied conditional independancies? Is there a way to look at the sets of these and see if any contradict each other??

# Simulation


Now we examine the feasibility of learning these models from data. A typical counter to going to the effort of eliciting these models from domain experts is that they can be learned from the data alone. However, we have seen here that much of the data is not present in the institutions ecosystem, with the nodes being deemed inherently latent (`r round(100*(sum(all_nodes_with_metrics$measurability == "L") / nrow(all_nodes_with_metrics)), 1)`%), potentially observable (`r round(100*(sum(all_nodes_with_metrics$measurability == "PO") / nrow(all_nodes_with_metrics)), 1)`%), observable but not currently in the data (`r round(100*(sum(all_nodes_with_metrics$measurability == "O") / nrow(all_nodes_with_metrics)), 1)`%), or already measured and stored in the current data ecosystem (`r round(100*(sum(all_nodes_with_metrics$measurability == "DW") / nrow(all_nodes_with_metrics)), 1)`%).

## Creating a simulated data set

We design this experiment as follows. We ask the question can a causal graph discovered from available data generate similar answers to the problem of assuring assessment? This translates to getting an unbiased estimate of Student Knowledge on Grade. As such 3 of the models are excluded, lacking sufficient adjustment sets (2 models) or having broken student knowledge into multiple nodes (1 model). 
For the "potentially observable" nodes a single descendant node is added, so that $N \rightarrow N_P$, a proxy measurement node that is not the exact node measured but whose value depends only on the potentially observable node and some additional noise. The set of proxy nodes for the potentially observable ($PO$) nodes we denote ($\hat{PO}$). 
Each of the 5 remaining GCMs, with their additional proxy measurement nodes, is used to generate a synthetic data set. We assuming each node has an equal linear effect on its descendants ($\beta=0.4$), generated using `simulateSEM` of the `dagitty` package in R. This treats all variables as continuous with mean 0 and variance 1. 

``` {r}
sim_signal_strength <- list("idyllic" = 0.9, "typical" = 0.4, "noisy" = 0.1)

tibble(Scenario = names(sim_signal_strength),
       Beta = as.numeric(sim_signal_strength)) |>
    mutate(
       `Variance due to each parent node` = Beta^2 / (1 + Beta^2)
       ) |> 
    knitr::kable(digits = 3)
```

We then apply a structure learning algorithm to each models simulated data in three phases: learning from the currently measured data only (only DW nodes), learning from all observable nodes (DW and O) nodes, and finally learning from all potentially available data (DW, O and $\hat{PO}$). This represents increasing levels of resourcing applied to learning the underlying causal structure, given the model. 

Note we are working under a local assumption that each model is the 'ground truth', but fully acknowledge that we do not know for this system what the true model is. However, if the structure learning claims of application to discover latent nodes are to be believed (CITE) then we should see some reasonable convergence towards each locally true model. 


``` {r simulate-data-from-dags}
# maybe use dagitty::simulateSEM() or dagitty::simulateLogistic()
# dagitty suggests using lavaan instead for SEM

# should add layout!!
# use same x, y from main graph

m_build_model_with_proxy_nodes <- function(
        m = "S1", 
        x_p_nudge = 0, y_p_nudge = -2,
        S.Kn_to_Obs = TRUE) {
    model <- m
    if (S.Kn_to_Obs) { # for running sims
        all_nodes_adj <- 
            all_nodes_with_metrics |> 
            mutate(measurability = if_else(
                # just picking S.Kn.Tech for S7
                name %in% c("S.Kn", "S.Kn.Tech"), "O",
                measurability
            ) |> 
                factor(levels = levels(all_nodes_with_metrics$measurability)))
    } else {
        all_nodes_adj <- all_nodes_with_metrics
    }
    nodes <- all_nodes_adj |>
        inner_join(merged_all_graph_layout_adjusted |> 
                       select(name, x, y),
                   by = "name") |> 
        filter(Model == toupper(model)) |> 
        select(name, x, y, node_group, node_type, measurability) |> 
        bind_rows( # adding proxy nodes for PO nodes
            all_nodes_adj |> 
                inner_join(merged_all_graph_layout_adjusted |> 
                       select(name, x, y),
                   by = "name") |> 
                filter(Model == model, measurability == "PO") |> 
                select(name, node_type, x, y) |> 
                mutate(name = str_c(name, "_p"),
                       # Need to place proxy nodes close - might cause issues
                       x = x + x_p_nudge,
                       y = y + y_p_nudge  
                       ) |> 
                mutate(node_group = "Proxy") |> 
                mutate(measurability = "O") # the assumption for the proxy
        ) |> 
        mutate(measurability = factor(
            measurability, 
            levels = levels(all_nodes_with_metrics$measurability)))
    
    edges <- all_edges |> 
        filter(Model == tolower(model)) |> 
        select(-Model) |> 
        bind_rows(
            tibble(
                from = nodes |> 
                    filter(measurability == "PO") |> 
                    pull(name)) |> 
                mutate(to = str_c(from, "_p"))
        )
    
    new_tidydag <- tidygraph::tbl_graph(
        nodes = nodes, edges = edges, directed = TRUE
    )
    new_dag <- dagitty_from_edge_df(edges)
    
    # setting coordinates of dagitty object - needs to be a named list
    node_names <- nodes$name
    node_coord_list <- list(x = nodes$x, y = nodes$y)
    names(node_coord_list$x) <- node_names
    names(node_coord_list$y) <- node_names
    coordinates(new_dag) <- node_coord_list
    
    node_list_dw <- nodes |> filter(measurability == "DW") |> pull(name)
    node_list_dwo <- nodes |> 
        filter(!str_detect(name, "_p")) |> 
        filter(measurability %in% c("DW", "O")) |> 
        pull(name)
    node_list_dwoop <- nodes |>  
        filter(measurability %in% c("DW", "O")) |> 
        pull(name)
    
    return(list(
        dag = new_dag, # new dag in dagitty format
        tdag = new_tidydag, # new dag as tidydag
        node_list_dw = node_list_dw, # list of nodes in DW
        node_list_dwo = node_list_dwo, # list of nodes in DW or Observable
        node_list_dwoop = node_list_dwoop # list of all non-latent nodes
        ))
}



simSEM <- partial(simulateSEM, eps = 1, standardized = F, N = 1000)

# d.noisy <- simSEM(new_dag, b.default = sim_signal_strength[["noisy"]])
# d.typical <- simSEM(new_dag, b.default = sim_signal_strength[["typical"]])
# d.ideal <- simSEM(new_dag, b.default = sim_signal_strength[["ideal"]])

bn_fit_model_list <- function(m_proxy_list, Beta = 0.6) {
    
    # getting simulated data, and subsets
    d.sim.sem <- simSEM(m_proxy_list$dag, b.default = Beta)
    d.dw <- as.data.frame(d.sim.sem[,m_proxy_list$node_list_dw])
    d.dwo <- as.data.frame(d.sim.sem[,m_proxy_list$node_list_dwo])
    d.dwoop <- as.data.frame(d.sim.sem[,m_proxy_list$node_list_dwoop])
    
    # fitting BNets
    bn.full =  bnlearn::bn.fit(bnlearn::hc(d.sim.sem), d.sim.sem)
    bn.dw =    bnlearn::bn.fit(bnlearn::hc(d.dw), d.dw)
    bn.dwo =   bnlearn::bn.fit(bnlearn::hc(d.dwo), d.dwo)
    bn.dwoop = bnlearn::bn.fit(bnlearn::hc(d.dwoop), d.dwoop)
    
    return( # returns a named list of simmed data and models
        list(
            # The full data set
            d.full   = d.sim.sem,
            bn.full  = bn.full,
            # Only nodes in the Data Warehouse (DW)
            d.dw     = d.dw,
            bn.dw    = bn.dw,
            # DW and Observable (O) nodes
            d.dwo    = d.dwo,
            bn.dwo   = bn.dwo,
            # Including Potentially Observable (PO), so ALL non-latent nodes
            d.dwoop  = d.dwoop,
            bn.dwoop = bn.dwoop
    ))
}

bn_to_dagitty <- function(bn) {
  edges <- bnlearn::arcs(bn) |> 
    as_tibble() |> 
    mutate(edge_str = str_c(from, " -> ", to))
  
  dagitty(str_c(
    "dag{",
    str_c(edges$edge_str, collapse = "; "),
    "}"
  ))
}

# Thinking of a simple dag X -> Y, we have 
# Var(Y) = b^2 * Var(X) + Var(eps) -- eps is the noise, and here it is set to 1 by default
# then Var(Y) = b^2 + 1
# Since the R^2 is the ratio of the variance explained by X:
# R^2 = Var(explained) / Var(total)
# R^2 = b^2 / (b^2 + 1)

dagitty_to_bn <- function(dag) {
  
  edges <- dagitty::edges(dag) |> 
    as_tibble() |> 
    filter(e == "->") |>  # directed edges only, excludes bidirected
    select(v, w) |> 
    rename(from = v, to = w)
  
  nodes <- unique(c(edges$from, edges$to))
  
  # Build empty network then set arcs
  net <- bnlearn::empty.graph(nodes)
  bnlearn::arcs(net) <- as.matrix(edges)
  net
}


causal_estimate_over_all_adj_sets <- function(
        dag, data, latent_nodes = NULL,
        exposure = "S.Kn", outcome = "Grade",
        intervention_val = 1) {
    # This estimates P(Grade|S.Kn, Adjustment Set) over all adjustment sets
    # This will give a range of that the model suggests is the true effect
    adj_sets <- adjustmentSets(dag, exposure = exposure, outcome = outcome,
                               type = "all")
    if (!is.null(latent_nodes)) {
        # remove any with latent nodes 
        adj_sets <- keep(adj_sets, \(x) !any(x %in% latent_nodes))
    }
    # Throw error if none!
    estimate_per_set <- map(adj_sets, function(adj_set) {
        if (length(adj_set) == 0) {
            message("No adjustment sets found")
            formula_str <- str_c(outcome, " ~ ", exposure)
        } else {
            formula_str <- str_c(outcome, " ~ ", exposure, " + ",
                                 str_c(adj_set, collapse = " + "))
        }
        fit <- lm(as.formula(formula_str), data = data)
        preds <- predict(fit,
                     newdata = data |>
                         mutate({{ exposure }} := intervention_val),
                     se.fit = TRUE)
    
    list(
        mean = mean(preds$fit),
        se   = sqrt(mean(preds$se.fit^2))
    )
    }
    )
    estimates <- map_dbl(estimate_per_set, "mean")
    std_errors <- map_dbl(estimate_per_set, "se")
    return(list(adjustment_set = adj_sets,
                estimate = estimates,
                std_error = std_errors))
}

generate_causal_estimate_comparisons <- function(
        Model = "S1",
        Beta = 0.6, # strength of signal along paths
        # might need to change for different models depending on present nodes
        exposure = "S.Kn", outcome = "Grade",
        intervention_val = 1,
        # for placiing proxy nodes
        x_p_nudge = 0, y_p_nudge = -2
) {
    ground_truth_dag <- ddags[[Model]]
    model_list <- m_build_model_with_proxy_nodes(
        m = Model, 
        x_p_nudge = x_p_nudge, 
        y_p_nudge = y_p_nudge)
    
    bn_list <- bn_fit_model_list(model_list, Beta = Beta)
    # Ground truth uses full simmed data
    est <- partial(causal_estimate_over_all_adj_sets,
                   exposure = exposure, outcome = outcome,
                   intervention_val = intervention_val)
    
    ground_truth <- est(dag = ground_truth_dag, data = bn_list$d.full)
    expert_dag <- est(
        dag = ground_truth_dag, 
        data = bn_list$d.full,
        latent_nodes = 
            all_nodes_adj |> 
            filter(Model == Model, measurability == "L") |> 
            pull(name))
    can_estimate_dw <- 
        (exposure %in% model_list$node_list_dw) & (outcome %in% model_list$node_list_dw)
    can_estimate_dwo <- 
        (exposure %in% model_list$node_list_dwo) & (outcome %in% model_list$node_list_dwo)
    can_estimate_dwoop <- 
        (exposure %in% model_list$node_list_dwoop) & (outcome %in% model_list$node_list_dwoop)
    if (can_estimate_dw) {
        dw <- est(   
            dag = bn_to_dagitty(bn_list$bn.dw) , 
            data = bn_list$d.dw)
    } else {
        dw <- list("adjustment_set" = NA, "estimate" = NA)
        message("Nodes not found in DW node list")
    }
    
    if (can_estimate_dwo) {
        dwo <- est(  
            dag = bn_to_dagitty(bn_list$bn.dwo) , 
            data = bn_list$d.dwo)
    } else {
        dwo <- list("adjustment_set" = NA, "estimate" = NA)
        message("Nodes not found in DWO node list")
    }
    
    if (can_estimate_dwoop) {
        dwoop <- est(
            dag = bn_to_dagitty(bn_list$bn.dwoop) , 
            data = bn_list$d.dwoop)
    } else {
        dwoop <- list("adjustment_set" = NA, "estimate" = NA)
        message("Nodes not found in DWOOP node list")
    }
    
    estimates <- tibble(
        Model = Model,
        DAG = c("Ground Truth", "Expert DAG", "Data Warehouse", "Observable", "Potentially Observable") |> 
            factor(levels = c("Ground Truth", "Expert DAG","Data Warehouse", "Observable", "Potentially Observable")),
        exposure = exposure, 
        outcome = outcome,
        adjustment_sets = list(
            ground_truth$adjustment_set,
            expert_dag$adjustment_set,
            dw$adjustment_set,
            dwo$adjustment_set,
            dwoop$adjustment_set
        ),
        estimate_mean = c(
            mean(ground_truth$estimate),
            mean(expert_dag$estimate),
            mean(dw$estimate),
            mean(dwo$estimate),
            mean(dwoop$estimate)
        ),
        estimate_std_errors = c(
            mean(ground_truth$std_error),
            mean(expert_dag$std_error),
            mean(dw$std_error),
            mean(dwo$std_error),
            mean(dwoop$std_error)
        ),
        estimate_sd = c(
            sd(ground_truth$estimate),
            sd(expert_dag$estimate),
            sd(dw$estimate),
            sd(dwo$estimate),
            sd(dwoop$estimate)
        )
    )
    
    return( # Monster list of everything we need
        list(
            Model = Model,
            DAG = ground_truth_dag,
            augmented_DAGs = model_list,
            fitted_BN = bn_list,
            estimates = estimates
        )
    )
}

s1_estimates <- 
    generate_causal_estimate_comparisons(
        Model = "S1", 
        Beta = sim_signal_strength[["typical"]])
s2_estimates <- 
    generate_causal_estimate_comparisons(
        Model = "S2", 
        Beta = sim_signal_strength[["typical"]])
# s3_estimates <- # ERROR! This one is cycle, and no adj sets
#     generate_causal_estimate_comparisons(
#         Model = "S3", 
#         Beta = sim_signal_strength[["typical"]])
s4_estimates <- # Acyclic: THIS IS SLOW!!
    generate_causal_estimate_comparisons(
        Model = "S4", 
        Beta = sim_signal_strength[["typical"]])
s5_estimates <- 
    generate_causal_estimate_comparisons(
        Model = "S5", 
        Beta = sim_signal_strength[["typical"]])
# s6_estimates <- # cyclic, no adj sets
#     generate_causal_estimate_comparisons(
#         Model = "S6", 
#         Beta = sim_signal_strength[["typical"]])
# s7_estimates <- # Different exposure - not sure what to do
#     generate_causal_estimate_comparisons(
#         Model = "S7", 
#         exposure = "S.Kn.Tech",
#         Beta = sim_signal_strength[["typical"]])
s8_estimates <- 
    generate_causal_estimate_comparisons(
        Model = "S8", 
        Beta = sim_signal_strength[["typical"]])

all_estimates <-
    s1_estimates$estimates |> 
    bind_rows(s2_estimates$estimates) |> 
    # bind_rows(s3_estimates$estimates) |> 
    bind_rows(s4_estimates$estimates) |> 
    bind_rows(s5_estimates$estimates) |> 
    # bind_rows(s6_estimates$estimates) |> 
    # bind_rows(s7_estimates$estimates) |> 
    bind_rows(s8_estimates$estimates) |> 
    filter(!is.na(estimate_mean)) |> 
    group_by(Model) |> 
    mutate(
        gt_est_diff = estimate_mean - estimate_mean[DAG == "Ground Truth"] )|> 
    ungroup() |> 
    mutate(error = estimate_std_errors + replace_na(estimate_sd, 0)) |> 
    select(Model, DAG, estimate_mean, error, gt_est_diff, everything())

dodge_amt <- 0.1
all_estimates |>
    filter(DAG != "Observable") |> 
    mutate(
        y_pos = as.numeric(factor(Model)) + 
            (as.numeric(factor(DAG)) - mean(as.numeric(factor(DAG)))) * dodge_amt
    ) |> 
    ggplot(aes(color = DAG, y = y_pos)) + 
    # geom_pointinterval(
    #     aes(x = gt_est_diff, 
    #         xmin = gt_est_diff - 2 * error, 
    #         xmax = gt_est_diff + 2 * error),
    #        alpha = 0.3, size = 1
    # ) +
    geom_pointinterval(
        aes(x = gt_est_diff,
            xmin = gt_est_diff - 1 *error,
            xmax = gt_est_diff + 1 *error),
           alpha = 0.8, size = 3
    ) +
    scale_y_continuous(
        breaks = seq_along(levels(factor(all_estimates$Model))),
        labels = levels(factor(all_estimates$Model)),
        name = "Model"
    ) +
    ggokabeito::scale_color_okabe_ito() +
    xlab("Deviation from ground truth estimate") +
    theme_minimal()

all_estimates |>
    filter(DAG != "Potentially Observable") |> 
    mutate(
        y_pos = 
            as.numeric(factor(Model)) + 
            (as.numeric(factor(DAG)) - mean(as.numeric(factor(DAG)))) * dodge_amt
    ) |> 
    ggplot(aes(color = DAG, y = y_pos)) + 
    geom_pointinterval(
        aes(x = estimate_mean, 
            xmin = estimate_mean - 1 *error, 
            xmax = estimate_mean + 1 *error),
           alpha = 0.8, size = 3
    ) +
    ggokabeito::scale_color_okabe_ito() +
    scale_y_continuous(
        breaks = seq_along(levels(factor(all_estimates$Model))),
        labels = levels(factor(all_estimates$Model)),
        name = "Model"
    ) +
    xlab("Estimated effect") +
    theme_minimal()

```


``` {r}
plot_bnlearn_comparisons <- function(s_estimates) {
    p_orig <- s_estimates$DAG |> 
        dag_to_tidy_graph() %N>%
        mutate(name = str_remove(name, "_p$")) |> 
        g_plot_fine(weight_nodes = F)
    p_dwo <- s_estimates$fitted_BN$bn.dwo |>
        bn_to_dagitty() |> 
        dag_to_tidy_graph() %N>%
        mutate(name = str_remove(name, "_p$")) |> 
        g_plot_fine(weight_nodes = F)
    p_dwoop <- s_estimates$fitted_BN$bn.dwoop |> 
        bn_to_dagitty() |> 
        dag_to_tidy_graph() %N>%
        mutate(name = str_remove(name, "_p$")) |> 
        g_plot_fine(weight_nodes = F)
    return(p_orig + p_dwo + p_dwoop)
}

plot_bnlearn_coarse_comparison <- function(s_estimates) {
    p_orig <- s_estimates$DAG |> 
        dag_to_tidy_graph() %N>%
        mutate(name = str_remove(name, "_p$")) |>
        g_coarsen_dag_by_ng() |> 
        g_plot_coarse()
    p_dwo <- s_estimates$fitted_BN$bn.dwo |>
        bn_to_dagitty() |> 
        dag_to_tidy_graph() %N>%
        mutate(name = str_remove(name, "_p$")) |> 
        g_coarsen_dag_by_ng() |> 
        g_plot_coarse()
    p_dwoop <- s_estimates$fitted_BN$bn.dwoop |> 
        bn_to_dagitty() |> 
        dag_to_tidy_graph() %N>%
        mutate(name = str_remove(name, "_p$")) |> 
        g_coarsen_dag_by_ng() |> 
        g_plot_coarse()
    return(p_orig  / p_dwo )
    
}

plot_bnlearn_coarse_comparison(s_estimates = s1_estimates) + ggtitle("Model1: Actual coarsened DAG (top) vs coarsened learned DAG from observable nodes")
plot_bnlearn_coarse_comparison(s_estimates = s2_estimates)+ ggtitle("Model 2: Actual coarsened DAG (top) vs coarsened learned DAG from observable nodes")
plot_bnlearn_coarse_comparison(s_estimates = s4_estimates)+ ggtitle("Model 4: Actual coarsened DAG (top) vs coarsened learned DAG from observable nodes")
 plot_bnlearn_coarse_comparison(s_estimates = s5_estimates)+ ggtitle("Model 5: Actual coarsened DAG (top) vs coarsened learned DAG from observable nodes")
plot_bnlearn_coarse_comparison(s_estimates = s6_estimates)+ ggtitle("Model 6: Actual coarsened DAG (top) vs coarsened learned DAG from observable nodes")
 plot_bnlearn_coarse_comparison(s_estimates = s8_estimates)+ ggtitle("Model 8: Actual coarsened DAG (top) vs coarsened learned DAG from observable nodes")



plot_bnlearn_comparisons(s_estimates = s1_estimates)
plot_bnlearn_comparisons(s_estimates = s2_estimates)
plot_bnlearn_comparisons(s_estimates = s4_estimates)
plot_bnlearn_comparisons(s_estimates = s5_estimates)
plot_bnlearn_comparisons(s_estimates = s6_estimates)
plot_bnlearn_comparisons(s_estimates = s8_estimates)

```
