---
title: "Eliciting causal models -- Analysis"
execute:
  dev: svglite
format:
    html:
        code-fold: true
        warning: false
        message: false
        fig-format: svg
---

# TODO


## Write functions to help node coding with causal abstraction theory

- Given a graph, suggest L1-L2 permissible abstractions (to some degree)
- Given a coarsening, check for L1-L2 validity
- Would be good to have some workflow documented

``` {r}
#| include: false
# Old yaml - put back for de-anonymised repo
# execute:
#   dev: svglite
# format: 
#     html:
#         code-fold: true
#         warning: false
#         message: false
#         fig-format: svg

# Anonymised repo yaml
# format: 
#    pdf:
#        fig-format: png
```

``` {r library-load}
#options(device = svglite::svglite)

library(svglite)
library(tidyverse)
library(dagitty)
library(tidygraph)
library(scales)
library(patchwork)
library(ggsci) # drop??
library(ggokabeito) # drop??
library(ggdist)
library(ggraph)

# library(ggpubfigs)
#library(cowplot)
```

``` {r helper-function}
source("graph_helpers.R")
```

# Notation

We will use curvy uppercase ($\mathcal{M,G,V,E}$) script for sets across all the models, $\mathbf{BOLD}$ uppercase for sets, and $UPPER$ case standard for instances in the models. Lowercase like $v$ will be reserved for imagining actual values (so a node taking on a particular value).  

The models are indexed as $\mathcal{M}=\{M_1,M_2,\dots,M_8 \}$ with their corresponding graphical form as $\mathcal{G}=\{G_1,G_2,\dots,G_8 \}$, and each graphical model $G_k$ comprises of a set of vertices $\mathbf{V}_{G_k}$ and edges $\mathbf{E}_{G_k}$, so $G_k=(\mathbf{V}_{G_k}, \mathbf{E}_{G_k})$, with the set of all nodes being $\mathcal{V}$ edges being $\mathcal{E}$. $E_{ij}=V_i \rightarrow V_j$ indicates an edge between two nodes. If two nodes are connected through a mediated path, such as $V_i \rightarrow \dots \rightarrow V_j$ then we write $\mathrm{De}_{ij}=V_i \rightsquigarrow V_j$ or $V_i \in \mathrm{an}(V_j)$. To specify model context for ancestors, descendants and mediated parts we subscript the model: $V_i \in \mathrm{an}_{G_k}(V_j)$ or $V_i \rightsquigarrow _{G_k} V_j$. Nodes $V_i$ are indexed over all models (so across $\mathcal{V}$), but if we want to talk about $V_i$ as it stands in $G_k$ we would use $V_{i:G_k}$, and likewise for an edge $E_{ij:G_k}$ residing in the context of model $G_k$.

We will also be applying coarsening functions to create cluster DAGs. A typical coarsening functions $\varphi$ is applied to the graph, and within it all the nodes and edges. So $\varphi (G_{k})=\varphi G_{k}=(\mathbf{V}_{\varphi G_k},\mathbf{E}_{\varphi G_k})$, which will also refer to as a new cluster model $G^C_{k}=\varphi G_k$ for slightly cleaner notation, and including the $C$ for cluster. The coarsening function $\varphi$ induces an equivalence relation $\sim _\varphi$ on $\mathbf{V}_{G_k}$ where $V_i \sim_\varphi V_j$ iff $\varphi (V_i) = \varphi (V_j)$, i.e. both nodes map to the same cluster.  In this cluster model we would find $\mathbf{V}_{\varphi G_k}$. The coarsened / cluster nodes would be $[V_{i}]_{\varphi}=\{V_j \in \mathcal{V}:V_j \sim_{\varphi} V_i\}$, the equivalence class of nodes $[V_i]_{\varphi}$, or simply $[V_i]$ if $\varphi$ is clear from the context. These clusters can be relabled and reindexed to $C_a = [V_i]_{\varphi}$ for legibility.
The edges in $\varphi G_{k}=G_{k}^{C}$ are then $E^{\varphi}_{[i][j]}=[V_i]\rightarrow [V_j]=C_a \rightarrow C_b=E^{C}_{ab}$. Choosing to use the $C$ or $\varphi$ syntax depends on the context - are we mainly discussing the cluster model $G_k^C$ or the processes of coarsening $\varphi G_{k}$?
There is also the *causal degradation* produced by $\varphi$, which we denote as $\mathrm{deg}_\varphi$ generally, $\mathrm{deg}_\varphi({G_k})$ as applied to a particular model, and $\mathcal{L}_i{\text{-}}\mathrm{deg}_\varphi(G_k)$ to specify a particular level of the causal hierarchy, $\mathcal{L}_i:i\in \{1,2,3\}$. 

A measure of *causal consistency* is also computed, but this is between two models, rather than under the process of a clustering. So 

Notation for adjustment sets? Things like that?



# Data import

``` {r loading-models}
# imports all causal models as DAGitty objects
# They are names dag_s1, ..., dag_s8 corresponding to the eight sessions
# A table of session data is also included
source("load-GCMs.R")
```

``` {r colour-settings}
# Some colours
# \definecolor{myochre}{RGB}{204,119,34}
# \definecolor{mygum}{RGB}{87,180,66}
myochre <- rgb(204, 119, 34, maxColorValue = 255)
mygum <- rgb(87, 180, 66, maxColorValue = 255)
nt_pal <- palette.colors()[c(1,9,8,4,2,3)]
```

## Graphical Causal Models

Modelling sessions were originally run on Loopy, then transcribed into DAGitty (online) to generate the following `dot` language syntax for importing into `R`. A better way of doing this is under development. 

Data imported as `DAGitty` objects, but need to be in `tidygraph` format for many calculations and visualisations.


Creating a list of all the DAGs in tidy format:

``` {r}
source("dagitty_tidygraph_translators.R")
# making list of dagitty dags (used later)
ddags <- list(
    "M1" = dag_m1, 
    "M2" = dag_m2,
    "M3" = dag_m3,
    "M4" = dag_m4,
    "M5" = dag_m5,
    "M6" = dag_m6,
    "M7" = dag_m7,
    "M8" = dag_m8
    )

tdags <- list(
    "M1" = dag_to_tidy_graph(dag_m1), 
    "M2" = dag_to_tidy_graph(dag_m2), 
    "M3" = dag_to_tidy_graph(dag_m3), 
    "M4" = dag_to_tidy_graph(dag_m4),
    "M5" = dag_to_tidy_graph(dag_m5), 
    "M6" = dag_to_tidy_graph(dag_m6) |> 
    filter( # this removes feedback link to next learning cycle
        !str_detect(name, "Post")), 
    "M7" = dag_to_tidy_graph(dag_m7), 
    "M8" = dag_to_tidy_graph(dag_m8))

ddags.raw <- ddags

ddags <- 
    imap(tdags, \(g, n)
         tidy_graph_to_dag(g)) |> 
    map(\(d) { 
        dagitty::outcomes(d) <- "Grade"
        d}) |> 
    imap(\(d, n) {
        if (n != "M7") {
            dagitty::exposures(d) <- 
                "S.Kn"
        } else {
            dagitty::exposures(d) <- 
                c("S.Kn.Coms", "S.Kn.Interp", "S.Kn.Tech")
        }
        d})

quick_list_of_all_nodes <- 
    map(tdags, \(g) pull(g, name)) |> 
    unlist() |> 
    unique() |> 
    sort()
```

# Coding

## Node Classifiers

Coding the nodes into rough groups. There are two levels:

- `node_group` is the first aggregation into slightly coarser groups of nodes.
- `node_type` is an even more coarse grouping into a handful of types.

Also some nicer names of the fields:
- `node_nice_name` is a longer more explanatory name for plotting and putting in tables.
- `node_group_nice_name` a longer more explanatory name for node groups, as above.


``` {r nicer-names}
name_node_nice_name <- tribble(
    ~name, ~nice_name,
    "CD", "Any system data",
    "CD.Att", "Attendance",
    "CD.LMS", "LMS activity",
    "CD.LMS.code", "Code artifacts",
    "CD.LMS.write", "Writing artifacts", 
    "CD.Str", "Course Data",
    "Grade", "Assessment grade",
    "Grade.Coms", "Mark - Communication",
    "Grade.Exam", "Mark - Exam",
    "Grade.Interp", "Mark - Interpretation",
    "Grade.Proc", "Mark - Learning process",
    "Grade.Quiz", "Mark - Quiz",
    "Grade.Ref", "Mark - Learning reflection",
    "Grade.Tech", "Mark - Technical",
    "H", "Help (general)",
    "H.AI", "Help from AI (any)",
    "H.AI.Agent", "Help from AI (agent)",
    "H.AI.evil", "Misuse of AI",
    "H.AI.evil.code", "Misuse of AI (code)",
    "H.AI.evil.write", "Misuse from AI (write)",
    "H.AI.Fb", "Help from AI (feedback)",
    "H.AI.good", "Help from AI (to learn)",
    "H.AI.Res", "Help from AI (research)",
    "H.AI.Study", "Help from AI (revision)",
    "H.Ment", "Help from mentor",
    "H.other.evil", "Cheating - non-AI",
    "H.Peer", "Help from peers",
    "H.Tut", "Help from teacher / tutor",
    "H.Tut.Fb", "Feedback from teacher / tutor",
    "LD", "Learning design",
    "LD.Other", "Other learning design",
    "LD.AIguide", "Guidance on AI use",
    "LD.Task", "Task design",
    "LD.Task.Code", "Coding task design",
    "LD.Task.Grp", "Groupwork task design",
    "LD.Task.Other", "Other task design",
    "LD.Task.Prac", "Practical task design",
    "LD.Task.Write", "Writing task design",
    "S.Act", "Student decisions / actions",
    "S.Act.AIstrat", "Student AI strategy",
    "S.Act.Fb", "Student acts on feedback",
    "S.Act.Fb.Post", "Student acts on feedback between subjects",
    "S.Act.Frm", "Student does formative tasks",
    "S.Act.Perf", "Student performance on task",
    "S.Act.Time", "Student commits time",
    "S.Aff.Integ", "Student acting with integrity",
    "S.Aff.Joy", "Student enjoyment",
    "S.Aff.MindSet", "Student mindset",
    "S.Aff.Safe", "Student feels safe",
    "S.Aff.WrkLd", "Student workload pressure",
    "S.Cp.WrkLd", "Student workload pressure",
    "S.Aff.Hsk", "Student help seeker",
    "S.Att.Eq", "Student equity group",
    "S.Cp.Hsk", "Student help seeker",
    "S.Att.Nro", "Student neuro-divergent",
    "S.Att.SocCap", "Student social capital",
    "S.Kn.Aca", "Student academic skill",
    "S.Cp.Aca", "Student academic skill",
    "S.Cp.Aca.Post", "Student academic skill: Post grade",
    "S.Cp.EvJdg", "Student evaluative judgement",
    "S.Cp.Fb", "Student feedback literacy",
    "S.Cp.Grp", "Student groupwork capability",
    "S.Cp.Ind", "Student individual work capability",
    "S.Cp.Lrn", "Student metacognative learning capabilities",
    "S.Cp.SRL", "Student SRL",
    "S.Kn", "Student subject knowledge",
    "S.Kn.Post", "Student subject knowledge: Post grade",
    "S.Kn.Coms", "Student communications knowledge",
    "S.Kn.Interp", "Student interpretation knowledge",
    "S.Kn.Prior", "Student prior knowledge",
    "S.Kn.Tech", "Student technical knowledge",
    "T.Act.UseAI", "Teacher use of AI",
    "T.Aff.Integ", "Teacher acts with integrity",
    "T.Aff.WrkLd", "Teacher workload pressure",
    "T.Cp.WrkLd", "Teacher workload pressure",
    "T.Cp.Inst", "Teacher instruction quality",
    "T.Cp.KnStu", "Teacher knows students",
    "T.Cp.Mark", "Teacher reliably marks",
    "T.Kn.AI", "Teacher AI literacy",
    "T.Cp.AI", "Teacher AI expertise",
    "T.Kn.Ass", "Teacher assessment literacy",
    "T.Kn.Sub", "Teacher subject knowledge",
    "T.Cp.Sub", "Teacher subject expertise"
)

node_group_nice_name <- tribble(
    ~node_group, ~nice_name,
    "Assessments", "Preliminary assessments", 
    "CD.Att", "Student attendance",
    "CD.LMS", "LMS activity",
    "CD.Str", "Course Data",
    "Grade", "Assessment grade",
    "H.AI.evil", "Misuse of AI",
    "H.AI.good", "Help from AI",
    "H.Cheat", "Cheating - non-AI",
    "H.Peer", "Help from peers",
    "H.Gen", "Help (General)",
    "H.Tut", "Help from teacher / tutor",
    "LD.Gen", "Learning design",
    "LD.Task", "Task design",
    "S.Act", "Student decisions / actions",
    "S.Att", "Student attributes",
    "S.Cp", "Student metacognitive skill",
    "S.EM", "Student emotion / motivation",
    "S.Kn", "Student knowledge",
    "S.Kn.Prior", "Student prior knowledge",
    # "T.Act", "Teacher use of AI",
    "T.Cp", "Teacher capabilities", 
    "T.EM", "Teacher emotion / motivation",
    "T.Kn", "Teacher knowledge"
)
```

### Coding node groups

This process involved iterating between:

1. Examining the transcript, the node description (label in the diagram), and forming an initial classification of the node. A natural place for a node-group to form was where in one model a node was broken into sub-components, where in others it was kept whole. `Student knowledge` was a key example of this, where several models split this into finer grained nodes, where others did not. 
2. After this the graph coarsening according to the map $\varphi(V_i)=\mathrm{nodegroup}(V_i)$, calculating the $\mathcal{L}_1$ and $\mathcal{L}_2$ cost of clustering the nodes into their node groups. From this two operations were performed on the cluster graph: (a) identifying which node clusters had contributed most to the degradation of the causal consistency, and (b) parts of the cluster graph that could potentially be coarsened further. 

``` {r node-group}
code_node_group <- function(v) {
    case_when(
        # Course data - Str, LMS --------------
        v == "CD" ~ "CD.Str", # was blending LMS and Str, but original graph was more about structure
        str_detect(v, "CD\\.LMS") ~ "CD.LMS",
        v == "CD.Att" ~ "CD.Att", # Attendance data
        v == "CD.Str" ~ "CD.Str",
        # Grade data --------------------------
        v == "Grade" ~ "Grade",
        str_detect(v, "Grade\\.") ~ "Grade", # Assessments
        # Kinds of help / assistance -----------------------
        # - Help from teacher or tutor
        # - - H.Gen(eral) for the broad H category
        v == "H" ~ "H.Gen", # This should actually be the node_type...somehow
        v == "H.Tut" ~ "H.Tut",
        v == "H.Ment" ~ "H.Tut",
        v == "H.Tut.Fb" ~ "H.Tut",
        # - Help from peers - collapsing with Tut for "help from people" - change from first coding
        v == "H.Peer" ~ "H.Tut",
        # - Help from AI (good or bad usage)
        str_detect(v, "H\\.AI\\.evil") ~ "H.AI.evil",
        str_detect(v, "H\\.AI") ~ "H.AI.good",
        # - Other nefarious help (i.e. cheating)
        v == "H.other.evil" ~ "H.Cheat",
        # Learning Design --------------------
        str_detect(v, "^LD\\.Task") ~ "LD.Task",
        str_detect(v, "^LD") ~ "LD.Gen",
        # Student nodes ---------------------------------
        # - Student activity / decisions 
        # - - Action post grade a seperate node
        v == "S.Act.Fb.Post" ~ "S.Act.Post",
        v == "S.Cp.Aca.Post" ~ "S.Cp.Post",
        str_detect(v, "^S\\.Act") ~ "S.Act",
        # - Student attributes (generally unchangable) 
        str_detect(v, "^S\\.Att") ~ "S.Att",
        # - Student metacognitive capability (metacognition) / capacity / workload
        str_detect(v, "^S\\.Cp") ~ "S.Cp",
        # - Student emotion / motivation
        str_detect(v, "^S\\.Aff") ~ "S.EM",
        # - Student knowledge (cognition)
        v == "S.Kn.Prior" ~ "S.Kn.Prior", 
        v == "S.Kn.Post" ~ "S.Kn.Post", 
        str_detect(v, "^S\\.Kn") ~ "S.Kn",
        # Teacher nodes ---------------------------------
        # - Teacher knowledge / meta-cognition: What they bring to the table
        str_detect(v, "^T\\.Kn") ~ "T.Kn",
        # - Teacher capabilities - usually executed during learning cycle
        str_detect(v, "^T\\.Cp") ~ "T.Cp",
        # - Teacher emotion / motivation
        str_detect(v, "^T\\.Aff") ~ "T.EM",
        TRUE ~ v
        
    )
}
```

### Coding node types

``` {r node-type}
code_node_type <- function(n) {
    fl <- str_extract(n, "^.{1}")
    case_when(
        fl == "S" ~ "Student",
        fl == "T" ~ "Teacher",
        fl == "C" ~ "Learning data",
        fl == "H" ~ "Help",
        fl == "L" ~ "Learning design",
        fl == "G" ~ "Assessment data"
    ) |> 
        factor(levels = rev(c(
            "Learning design", 
            "Teacher", 
            "Student",
            "Help", 
            "Learning data",
            "Assessment data"
            )))
}

```

### Coding node measurability


From the node groups, describe how easy (or not) they are to measure. This is completed by someone familiar to the data warehouse at the institution:

- **DW** - _Data Warehouse_: Data is already collected and stored in the data warehouse of the institution.
- **O** - _Observable_: Node is not currently collected data, but would be possible to measure and collect with resourcing.
- **PO** - _Potentially Observable_:  Node is not measured, and would be difficult to measure directly. It is possible to find a direct proxy and measure that instead, such as a survey instrument.
- **L** - _Latent_: Node is not measured and would be difficult to even find a direct proxy for the latent construct. 


``` {r node-measurability}
code_node_measurability <- function(node) {
    case_when(
        # Student
        node == "S.Att.SocCap" ~ "PO", # check
        node == "S.Att.Nro" ~ "O", # check
        node == "S.Att.Eq" ~ "DW", # check
        node == "S.Att.HSk" ~ "L", # check
        node == "S.Act.AIstrat" ~ "L",
        node == "S.Act.Fb" ~ "L",
        node == "S.Act.Fb.Post" ~ "L",
        str_detect(node, "^S\\.Act") ~ "PO", 
        str_detect(node, "^S\\.Kn") ~ "PO", 
        str_detect(node, "^S") ~ "L", # Latent
        # Help
        node == "H.other.evil" ~ "L",
        node == "H.Peer" ~ "L",
        node == "H.AI.evil.code" ~ "PO", # Potentially observable
        node == "H.AI.evil.write" ~ "PO",
        node == "H.AI.evil.write" ~ "PO",
        str_detect(node, "^H\\.AI") ~ "PO",
        str_detect(node, "^H") ~ "L",
        # Assessment
        node == "Grade" ~ "DW",
        node == "Grade.Exam" ~ "DW",
        str_detect(node, "^Grade") ~ "O",
        # Learning data
        node == "CD.LMS" ~ "DW",
        str_detect(node, "^CD\\.LMS") ~ "O",
        node == "CD.Att" ~ "PO",
        node == "CD.Str" ~ "DW", # check
        node == "CD" ~ "O", # check
        # Learning design
        str_detect(node, "^LD") ~ "PO",
        # Teacher
        node == "T.Cp.UseAI" ~ "PO",
        node == "T.Aff.WrkLd" ~ "PO", 
        node == "T.Cp.Mark" ~ "PO", 
        str_detect(node, "^T\\.K") ~ "PO",
        str_detect(node, "^T") ~ "L",
        node == "Assessments" ~ "DW"
    ) |> 
        factor(levels = c("DW", "O", "PO", "L"))
}

coarsen_node_measurability <- function(ng_list) {
    # Takes in a list of values of the nodes in a cluster node
    case_when(
            all(ng_list %in%  c("O", "DW")) ~ "Observable",
            all(ng_list %in% c("PO")) ~ "Potentially observable",
            all(ng_list == "L") ~ "Latent",
            all(ng_list %in% c("PO", "L")) ~ "Likely latent",
            all(ng_list %in% c("DW","O","PO")) ~ "Likely observable",
            TRUE ~ str_c("Some observable (", 
                         sum(ng_list %in% c("O","DW")), " of ", 
                         n(), ")")) |> 
            factor(levels = c(
                "Latent", 
                "Likely latent", 
                "Potentially observable",
                "Likely observable",
                "Observable"))
}


```

# Graph processing

## Graph helper functions

Functions outlined in the `graph_helpers.R` script. Functions starting with `gcm_` should operate on either a `dagitty` or `tidygraph` object. 

### Graph metrics functions

Generating a range of edge metrics - comparisons across models. Want to label edges in the unified DAG as 'contested' if they switch direction between some models. Also want to examine the same idea for a list of descendants. This is handled as a new table of edges - including all possible edges. 

``` {r fancy-edge-weight}
all_nodes <- 
    imap(tdags, \(g,m) gcm_nodelist(g)) |> 
    list_rbind(names_to = "Model")

all_edges <- 
    imap(tdags, \(g, m) gcm_edgelist(g) |> mutate(Model = m)) |> 
    list_rbind() |> 
    switches_exist(from, to) |> 
    mutate(edge_alignment = if_else(switched, "contested", "aligned"))

all_possible_edges <- 
    imap(tdags, \(g, m) gcm_possible_edges(g)) |> 
    list_rbind(names_to = "Model")

max_possible_edges <- 
    all_possible_edges |> 
    filter(from != to) |> 
    count(from, to) |> 
    rename(max_n = n)

all_descendants <- 
    imap(tdags, \(g,m) gcm_all_descendants_edge_list(g)) |> 
    list_rbind(names_to = "Model") |> 
    switches_exist(from, to) |> 
    mutate(descendant_alignment = if_else(switched, "descendant:contested", "descendant:aligned")) |> 
    select(-switched)

all_possible_edges_with_metrics <- 
    all_possible_edges |> 
    filter(from != to) |> 
    left_join(
        all_edges |> 
            mutate(in_graph = TRUE),
        by = c("Model", "from", "to")) |> 
    mutate(in_graph = replace_na(in_graph, FALSE)) |> 
    left_join(
        all_descendants |> 
            mutate(descendant_in_graph = TRUE),
        by = c("Model", "from", "to")) |> 
    mutate(descendant_in_graph = replace_na(
        descendant_in_graph, FALSE))

all_edges_sum <- 
    all_possible_edges_with_metrics |>
    group_by(from, to) |> 
    summarise(
        n = sum(in_graph),
        n_descendant = sum(descendant_in_graph),
        N = n(),
        direct_alignment = mean(edge_alignment == "aligned", na.rm = T),
        descendant_alignment = mean(
            descendant_alignment == "descendant:aligned", na.rm = T),
        .groups = "drop") |> 
    mutate(
        p_direct = n / N, 
        p_descendant = n_descendant / N) |> 
    arrange(desc(n)) 


all_edges_in_models_with_metrics <- 
    all_edges_sum |> 
    filter(n > 0) |> 
    inner_join(max_possible_edges,
               by = c("from", "to")) |> 
    mutate(
        edge_p_w = n / max_n,
        edge_dp_w = n_descendant / max_n) |> 
    mutate(edge_w = n) |> 
    switches_exist(from, to)

```


## Creating merged graph

``` {r making-merged-fine-graph}

merged_edges <- 
    all_edges_in_models_with_metrics

merged_nodes <- 
    all_nodes |> 
    count(name) |> 
    rename(node_w = n) |> 
    mutate(node_group = code_node_group(name)) |> 
    mutate(node_type = code_node_type(name)) |> 
    mutate(measurable = code_node_measurability(name)) |> 
    left_join(name_node_nice_name, by = "name")

merged_all_graph <- tbl_graph(nodes = merged_nodes, edges = merged_edges)

```

## Coarsen graphs

This is based on the coding of the nodes into node groups. 

Here we merge nodes into node groups before generating plot later. This has the potential to introduce loops depending on how the constructs were built, so we need to check the causal consistency between the fine-grained graph and the clustered / merged / higher-level graph. 

``` {r}
clean_node_group_cluster_graph <- function(cdag) {
    # After coarsening with code_node_group cleans up some of the 
    # names and nodes
    cdag |> 
        activate(nodes) |> 
        mutate(node_type = code_node_type(name)) |>
        select(-any_of(c("nice_name", "node_group"))) |> 
        left_join(
            node_group_nice_name |> 
                rename(name = node_group),
            by = "name") 
}

node_group_cluster_graph_from_merged <- 
    gcm_cluster_graph(merged_all_graph, code_node_group) |> 
    clean_node_group_cluster_graph()
    

tcdags_node_group <- 
    map(tdags, 
        \(g)
        gcm_cluster_graph(g, code_node_group) |> 
            clean_node_group_cluster_graph()) 

node_group_cluster_graph_from_clusters <- 
    gcm_merge(
        gcm_merge(
            gcm_merge(tcdags_node_group[[1]], tcdags_node_group[[2]]),
            gcm_merge(tcdags_node_group[[3]], tcdags_node_group[[4]])
        ),
        gcm_merge(
            gcm_merge(tcdags_node_group[[5]], tcdags_node_group[[6]]),
            gcm_merge(tcdags_node_group[[7]], tcdags_node_group[[8]])
        )
    )

```

# Graph metrics

## Measurability, complexity

### Model level metrics

``` {r}
all_paths_summary <-
    imap(ddags, \(d, n) gcm_count_paths(d)) |>
    list_rbind(names_to = "Model")

all_paths <- 
    imap(
        ddags,
        \(d, n) bind_cols(
            tibble(Model = n),
            dagitty::paths(d, limit = 5000)) |> 
            group_by(Model) |> 
            summarise(
                paths = list(paths),
                n_paths = n(),
                n_open = sum(open),
                p_open = mean(open)
            ))|> 
    list_rbind()


dag_complexity_metrics <- 
    imap(tdags, \(g, n)  as_tibble(gcm_summary_metrics(g))
    ) |> 
    list_rbind(names_to = "Model") |> 
    inner_join(
        imap(
            ddags, 
            \(d, n) 
            tibble(Model = n) |> 
                mutate(
                    n_adj_sets = length(dagitty::adjustmentSets(d)))) |> 
            list_rbind(), 
        by = "Model"
    ) |> 
    inner_join(all_paths_summary, by = "Model")
```

### Node group level metrics

``` {r}
all_nodes_with_metrics <- 
    imap(tdags, \(g,n) gcm_node_metrics(g)) |> 
    list_rbind(names_to = "Model")
```

## Joining metrics and session data

``` {r}
gcm_metrics <- 
    inner_join(session_data, dag_complexity_metrics, by = "Model") 
```

# Visualisations

## Idea on layouts of graphs

- Sunflower layout: Each node group placed in $\phi$ ratio spiral out from centre, and each sub node spiraling out from these. A good order for this would be in temporal (roughly) of the node types: Learning design, Teacher, Student, Help, Learning data, Assessment data. 

``` {r}
all_nodes_with_metrics |> 
    count(node_type) |> 
    arrange(node_type)
```

## Modelling sessions, time and perceptions

### Figure: Time spent with model and modelling utility

``` {r}
# denser option - radial arc bars for observable
library(scales)
library(ggforce)
library(patchwork)
library(cowplot)

hue1 <- palette.colors(3, palette = "Okabe-Ito")[[2]]
hue2 <- palette.colors(3, palette = "Okabe-Ito")[[3]]

# Main plot - hide all legends
p_validity_and_time <- gcm_metrics |> 
    mutate(
        x_duration = as.numeric(Modelling.duration, units = "minutes"),
        y_validity = GCM.validity
    ) |> 
    mutate(acyclic = if_else(acyclic, "Yes", "No")) |> 
    ggplot(aes(
        color = acyclic,
        size = n_adj_sets
    )) +
    geom_point(aes(
        x = x_duration,
        y = y_validity)) +
    scale_y_continuous(
        limits = c(1, 4.05),
        breaks = c(1, 2, 3, 4),
        labels = c("Trivially", "Simply", "Adequately", "Well"),
        name = "Representation of GCM") +
    scale_x_continuous(
        limits = c(33, 52),  # was 55 to  Keep original limits for the data
        breaks = c(35, 40, 45,50),
        name = "Time spent constructing model (minutes)") +
    scale_size(range = c(2, 7),
               breaks = range(gcm_metrics$n_adj_sets),
               name = "Adjustment\nsets (count)") +
    scale_color_manual(values = c(hue1, hue2), name = "Acyclic?") +
    scale_fill_manual(values = c(hue1, hue2), guide = "none") +
    theme_minimal() +
    # coord_fixed(ratio = 2
    #             # , clip = "off"
    #             ) +
    theme(
        legend.position = "left",  # Hide legends from main plot
        # panel.grid.major.x = element_blank(),  # Remove vertical gridlines
        panel.grid.minor = element_blank(),
        plot.margin = margin(10, 10, 10, 10)
    )


p_visibility <- gcm_metrics |>
    mutate(
        cp_dw = p_dw,
        cp_o = p_dw + p_o,
        cp_po = p_dw + p_o + p_po,
        cp_l = p_dw + p_o + p_po + p_l # should equal 1
    ) |> 
    mutate(GCM.validity = case_when(
        GCM.validity == 3 & acyclic ~ GCM.validity - 0.03,
        GCM.validity == 3 & !acyclic ~ GCM.validity + 0.03,
        GCM.validity == 2.5 & acyclic ~ GCM.validity - 0.03,
        GCM.validity == 2.5 & !acyclic ~ GCM.validity + 0.03,
        TRUE ~ GCM.validity
    )) |> 
    pivot_longer(c(starts_with("cp_"), starts_with("p_")), names_to = "node_vis", values_to = "proportion") |> 
    mutate(node_visibility = case_when(
        str_detect(node_vis, "p_dw") ~ "Data Warehouse",
        str_detect(node_vis, "p_o") ~ "Observable",
        str_detect(node_vis, "p_po") ~ "Potentially observable",
        str_detect(node_vis, "p_l") ~ "Latent"
    ) |> 
        ordered(levels = c(
            "Latent",
            "Potentially observable",
            "Observable",
            "Data Warehouse"
        ))) |> 
    mutate(prop_type = if_else(str_detect(node_vis, "^cp_"), "cp", "p")) |> 
    pivot_wider(names_from = prop_type, values_from = proportion, id_cols = c(Model,acyclic, GCM.validity, node_visibility)) |> 
    mutate(acyclic = if_else(acyclic, "Yes", "No")) |>  
    ggplot(aes(
        color = acyclic
    )) +
    geom_rect(aes(
        xmin = cp - p,
        xmax = cp,
        ymin = GCM.validity - 0.03,
        ymax = GCM.validity + 0.03,
        fill = acyclic,
        alpha = node_visibility),
        color = "white") +
    scale_y_continuous(
        limits = c(1, 4.05),
        breaks = c(1, 2, 3, 4),
        labels = c("Trivially", "Simply", "Adequately", "Well"),
        name = "Representation of GCM",
        position = "right") +
    scale_alpha_discrete(name = "Data visibility") +
    scale_fill_manual(values = c(hue1, hue2), guide = "none") +
    theme_minimal() +
    # coord_fixed(ratio = 0.15
    #             # , clip = "off"
    #             ) +
    xlab("Proportion of nodes") +
    theme(
        legend.position = "right",  # Hide legends from main plot
        # panel.grid.major.x = element_blank(),  # Remove vertical gridlines
        panel.grid.minor = element_blank(),
        plot.margin = margin(10, 10, 10, 10),
        axis.title.y = element_blank(),
        axis.text.y = element_blank()
    )


p_time_and_representation_1 <- p_validity_and_time + p_visibility
p_time_and_representation_1

ggsave("figures/FIG--time-modelling-and-representation.jpg", 
       plot = p_time_and_representation_1,
       width = 10, height = 5, 
       dpi = 300)

```




As a table:

``` {r}
gcm_metrics |> 
    mutate(acyclic = if_else(acyclic, "Yes, DAG", "No, contained cycles")) |> 
    arrange(desc(Modelling.duration)) |> 
    select(Modelling.duration, GCM.validity, acyclic, n_adj_sets, p_dw, p_o, p_po, p_l)
```






## Merging graphs

``` {r plotting-merged-graph}

merged_all_graph_layout <- create_layout(
    merged_all_graph %E>%
        mutate(switched = if_else(
            switched, "contested", "alligned")), # needs switched boolean
    layout = "sugiyama")

merged_all_graph_layout_adjusted <- 
    merged_all_graph_layout |> 
    mutate(
        # x = if_else(name == "S.Post.Fb",x + 20,x),
        y = if_else(name == "S.Act.Fb.Post",y + 4,y)
           ) |> 
    mutate(
        x = if_else(name == "S.Act",x + 20,x),
        y = if_else(name == "S.Act",y - 4,y)
           ) 
ptemp <- merged_all_graph_layout_adjusted |> 
    ggraph() +
        geom_edge_fan(aes(alpha = edge_w), strength = 0.2,
            arrow = arrow(length = unit(2, 'mm'), type = "closed"),
            start_cap = circle(3, 'mm'),
            end_cap = circle(3, 'mm')) +
        geom_node_point(aes(colour = node_type, size = node_w)) +
        # geom_node_text(aes(label = name, size = node_w, alpha = node_w), nudge_y = -.5) +
        theme_graph() +
        scale_edge_alpha_continuous(range = c(0.15,0.8)) +
        scale_alpha_continuous(range = c(0.3,1)) +
        theme(legend.position = "none")

b <- ggplot_build(ptemp)
x_limits_fine <- b$layout$panel_params[[1]]$x.range
y_limits_fine <- b$layout$panel_params[[1]]$y.range



plot_fine_graph_from_layout <- function(lyt, weight_nodes = TRUE) {
    ppp <- 
        lyt |> 
        ggraph() 
    
    if (weight_nodes) {
        pp <- ppp +
            geom_node_point(aes(colour = node_type,
                                size = node_w)) +
            scale_size(name = "point-size") +
            geom_edge_fan(aes(
                alpha = edge_p_w * edge_w + edge_w, 
                edge_linetype = factor(switched)), 
                strength = 0.2,
                arrow = arrow(length = unit(1, 'mm'),
                              type = "closed"),
                start_cap = circle(3, 'mm'),
                end_cap = circle(3, 'mm'))
    } else {
        pp <- ppp +
            geom_node_point(aes(colour = node_type)) +
            geom_edge_fan(strength = 0.2, alpha = 0.5,
                          arrow = arrow(length = unit(1, 'mm'),
                                        type = "closed"),
                          start_cap = circle(1, 'mm'),
                          end_cap = circle(1, 'mm'))
            
    }
    pp +
        scale_colour_discrete(palette = nt_pal) +
        # geom_node_text(aes(label = name, size = node_w, alpha = node_w), nudge_y = -.5) +
        theme_graph() +
        scale_edge_alpha_continuous(range = c(0.15,0.8)) +
        scale_alpha_continuous(range = c(0.3,1)) +
        theme(legend.position = "none",
              plot.margin = margin(t = 0, r = 0, b = 0, l = 0, unit = "pt"))  +
        xlim(x_limits_fine) +
        ylim(y_limits_fine)
}

g_plot_fine <- function(g, 
                        lyt = merged_all_graph_layout_adjusted, 
                        weight_nodes = TRUE) {
    create_layout(g, layout = "sugiyama") |> 
        left_join(
            lyt |> 
                select(x,y,name,node_group,node_type,node_w), 
            by = "name",
            suffix = c(".default", "")) |> 
        plot_fine_graph_from_layout(weight_nodes = weight_nodes)
}


pMergedFineLegendTall <- merged_all_graph_layout_adjusted |> 
    g_plot_fine() + 
    theme(legend.position = "right") +
    guides(size = "none", alpha = "none", edge_alpha = "none", edge_width = "none",
           color = guide_legend(ncol = 2)) +
    theme(legend.title = element_blank())

leg_tall <- cowplot::get_legend(pMergedFineLegendTall)
leg_tall_plot <- cowplot::ggdraw(leg_tall)

```


``` {r plotting-coarse-graph}
merged_coarse_layout <- 
    merged_all_graph |> 
    gcm_cluster_graph(code_node_group) |> 
    clean_node_group_cluster_graph() |> 
    create_layout(layout = "sugiyama") |> 
    mutate(
        x = case_when(
            name == "CD.Att" ~ 18.5,
            name == "T.Aff.WrkLd" ~ x - 5,
            name == "H.AI.good" ~ 27,
            TRUE ~ x),
        y = case_when(
            name == "CD.Att" ~ y + 1,
            name == "Grade" ~ y - 2.5,
            name == "CD.Str" ~ 3,
            name == "S.EM" ~ 5,
            TRUE ~ y)
           )

p_temp <- merged_coarse_layout |> 
     ggraph() +
        geom_edge_fan(aes(alpha = edge_w), strength = 0.2,
            arrow = arrow(length = unit(2, 'mm')),
            start_cap = circle(3, 'mm'),
            end_cap = circle(3, 'mm')) +
        geom_node_point(aes(colour = node_type, size = node_w)) +
    scale_color_okabe_ito(order = c(9, 7,1, 2,3, 6, 8,5, 4)) +
        # scale_color_npg(drop = FALSE) +
        # geom_node_text(aes(label = name, size = node_w, alpha = node_w), nudge_y = -.5) +
        theme_graph() +
        scale_edge_alpha_continuous(range = c(0.15,0.8)) +
        scale_alpha_continuous(range = c(0.2,0.6)) +
        theme(legend.position = "none") 

bc <- ggplot_build(p_temp)
x_limits_course <- bc$layout$panel_params[[1]]$x.range
y_limits_course <- bc$layout$panel_params[[1]]$y.range

coarse_plot_from_layout <- function(lyt) {
    lyt |> 
        ggraph() +
        geom_edge_fan(aes(alpha = edge_w), strength = 0.2,
            arrow = arrow(length = unit(2, 'mm')),
            start_cap = circle(3, 'mm'),
            end_cap = circle(3, 'mm')) +
        geom_node_point(aes(colour = node_type
                            # , size = node_w
                            )) +
        scale_color_okabe_ito(order = c(9, 7,1, 2,3, 6, 8,5, 4), drop = FALSE) +
        # ggsci::scale_color_npg(drop = FALSE) +
        # geom_node_text(aes(label = name, size = node_w, alpha = node_w), nudge_y = -.5) +
        theme_graph() +
        scale_edge_alpha_continuous(range = c(0.15,0.8)) +
        scale_alpha_continuous(range = c(0.2,0.6)) +
        theme(legend.position = "none") +
        xlim(x_limits_course) +
        ylim(y_limits_course)
}

g_plot_coarse <- function(g, lyt = merged_coarse_layout) {
    create_layout(g, layout = "sugiyama") |> 
        left_join(lyt |> select(x,y,name, node_type,node_w), by = "name",
                  suffix = c(".default", "")) |> 
    coarse_plot_from_layout()    
}

pCoarseMerged <- merged_coarse_layout |> 
    coarse_plot_from_layout()
# ggsave("dag_merged_course.svg", pCoarseMerged, , width = 6, height = 6)
```


### Common patterns - fine

Most common three-node patterns. 

``` {r}
merged_all_graph |> 
    activate("nodes") |>
    filter(node_w > 3) |>
    # activate("edges") |>
    # filter(edge_w > 1) |>
    g_plot_fine() +
    geom_node_text(aes(label = nice_name, colour = node_type), 
                   alpha = 0.7, repel = TRUE)

```

### Figure - all graphs, single highlight

``` {r}
plot_highlighted_fine_graph <- function(model = 1, LEGEND_POSITION = "none") {
    MODEL <- paste0("M", model)
    nodes_in_graph <- all_nodes |> 
        filter(Model == MODEL) |> 
        pull(name)
    edges_in_graph <- all_edges |> 
        filter(Model == MODEL) |> 
        select(from, to)
    
    edges_in_graph_indexed <- 
        edges_in_graph |> 
        mutate(
            from = match(from, merged_all_graph %N>% as_tibble() |> pull(name)),
            to = match(to, merged_all_graph %N>% as_tibble() |> pull(name))
        )
    
    g_highlight <- merged_all_graph %N>%
        mutate(
            node_highlight = if_else(
                name %in% nodes_in_graph, 2, 1
            )) %E>%
        left_join(
            edges_in_graph_indexed |> 
                mutate(edge_highlight = 2),
            by = c("from", "to")
        ) |> 
        mutate(edge_highlight = replace_na(edge_highlight, 1)) |> 
        create_layout("sugiyama")
    
    g_highlight |>
        ggraph() +
        geom_edge_fan(aes(
            # edge_linetype = factor(switched),
            alpha = edge_highlight), 
            strength = 0.2,
            arrow = arrow(length = unit(1, 'mm'),
                          type = "closed"),
            start_cap = circle(3, 'mm'),
            end_cap = circle(3, 'mm')) +
        geom_node_label(
            aes(label = nice_name, 
                color = node_type,
                size = node_w),
            alpha = 0.7,
            data = g_highlight |> 
                filter(node_highlight == 2),
            repel = T, max.overlaps = 30,
            show.legend = FALSE) +
        geom_node_point(aes(colour = node_type,
                            size = node_w,
                            alpha = node_highlight)) +
        scale_size(name = "point-size", range = c(2,3)) +
        # scale_color_okabe_ito(order = c(9, 7,1, 2,3, 6, 8,5, 4)) +
        scale_color_discrete(palette = nt_pal,
                             drop = FALSE) +
        # scale_color_npg() +
        theme_graph() +
        scale_edge_alpha_continuous(range = c(0.08,0.8)) +
        scale_alpha_continuous(range = c(0.1,1.0)) +
        # theme(legend.position = "none",
        #       plot.margin = margin(t = 0, r = 0, b = 0, l = 0, unit = "pt")) +
        theme(legend.position = LEGEND_POSITION) +
        guides(size = "none", edge_width = "none", alpha = "none", edge_alpha = "none",
               color = guide_legend(ncol = 1)) +
        theme(legend.title = element_blank())
}

merged_cluster_graph_ng_no_post_nodes <- 
    merged_all_graph |> 
    activate(nodes) |> 
    filter(!str_detect(name, "Post")) |> 
    gcm_cluster_graph(code_node_group) |> 
    clean_node_group_cluster_graph() |> 
    mutate(node_group = name)


plot_highlighted_coarse_graph <- function(model = 1, LEGEND_POSITION = "none",
                                          repel_labels = TRUE) {
    merged_coarse_models <- 
        merged_cluster_graph_ng_no_post_nodes
    
    MODEL <- paste0("M", model)
    node_groups_in_graph <- all_nodes |> 
        filter(Model == MODEL) |> 
        mutate(name = code_node_group(name)) |> 
        distinct(name) |> 
        pull(name) 
    edges_in_coarse_graph <- all_edges |> 
        filter(Model == MODEL) |> 
        select(from, to) |> 
        mutate(across(from:to, code_node_group))
    
    edges_in_graph_indexed <- 
        edges_in_coarse_graph |> 
        mutate(
            from = match(from, 
                         merged_coarse_models %N>% 
                             as_tibble() |> 
                             pull(name)),
            to = match(to, 
                       merged_coarse_models %N>% 
                           as_tibble() |> 
                           pull(name))
        )
    
    edge_integers <- 
        edges_in_graph_indexed |> 
        mutate(p2p3 = (2 ** from) * (3 ** to) ) |> 
        pull(p2p3)
    
    node_types <- merged_all_graph %N>%
        as_tibble() |> 
        distinct(node_group, node_type) 
    
    g_highlight <- merged_coarse_models %N>%
        mutate(
            node_highlight = if_else(
                name %in% node_groups_in_graph, 2, 1
            )) %E>%
        mutate(edge_highlight = if_else(
            ((2 ** from )*(3 ** to)) %in% edge_integers, 2, 1
        )) |> 
        create_layout("sugiyama")
    
    g_highlight |>
        ggraph() +
        geom_edge_fan(aes(
            # edge_linetype = factor(switched),
            alpha = edge_highlight), 
            strength = 0.2,
            arrow = arrow(length = unit(1, 'mm'),
                          type = "closed"),
            start_cap = circle(3, 'mm'),
            end_cap = circle(3, 'mm')) +
        geom_node_label(
            aes(label = nice_name, 
                color = node_type
                ,size = node_w
                ),
            alpha = 0.7,
            data = g_highlight |> 
                filter(node_highlight == 2),
            repel = repel_labels, max.overlaps = 30,
            show.legend = FALSE) +
        geom_node_point(aes(colour = node_type,
                            size = node_w,
                            alpha = node_highlight)) +
        scale_size(name = "point-size", range = c(2,3)) +
        # scale_color_npg() +
        scale_color_discrete(
            palette = nt_pal,
            drop = FALSE) +
        theme_graph() +
        scale_edge_alpha_continuous(range = c(0.08,0.8)) +
        scale_alpha_continuous(range = c(0.1,1.0)) +
        # theme(legend.position = "none",
        #       plot.margin = margin(t = 0, r = 0, b = 0, l = 0, unit = "pt")) +
        theme(legend.position = LEGEND_POSITION) +
        guides(size = "none", edge_width = "none", alpha = "none", edge_alpha = "none",
               label = "none",
               color = guide_legend(ncol = 1)) +
        theme(legend.title = element_blank())
}
```

### Figure: Highlighted graphs on merged background

``` {r}
# Model used in paper
md_for_paper <- 5
p_merged_highlighted_fine <- plot_highlighted_fine_graph(
    md_for_paper, 
                                                         LEGEND_POSITION = "right")
p_merged_highlighted_coarse <- plot_highlighted_coarse_graph(
    md_for_paper, LEGEND_POSITION = "right")
p_merged_highlighted_fine + ggtitle(str_c("Model  ",md_for_paper , ", original"))
p_merged_highlighted_coarse + ggtitle(str_c("Model  ",md_for_paper , ", coarsened")) 
# ggsave(plot = p_merged_highlighted_fine, 
#        filename = "figures/FIG--merged-fine-with-1-highlighted.jpg",
#        width = 8, height = 5, dpi = 600)
# 
# ggsave(plot = p_merged_highlighted_coarse, 
#        filename = "figures/FIG--merged-coarse-with-1-highlighted.jpg",
#        width = 8, height = 5, dpi = 600)
```

``` {r}
# Generating all
p_fine_show_and_save <- function(n){
    p_temp <- plot_highlighted_fine_graph(n, LEGEND_POSITION = "right") + ggtitle(paste0("Model ", n, " highlighted"))
    ggsave(plot = p_temp, 
           filename = paste0("figures/plt--merged-fine-with-", n, "-highlighted.png"),
           width = 8, height = 5, dpi = 600)
    return(p_temp)
}
p_coarse_show_and_save <- function(n){
    p_temp <- plot_highlighted_coarse_graph(n, LEGEND_POSITION = "right") + ggtitle(paste0("Coarsened model ", n, " highlighted"))
    ggsave(plot = p_temp, 
           filename = paste0("figures/plt--merged-coarse-with-", n, "-highlighted.png"),
           width = 8, height = 5, dpi = 600)
    return(p_temp)
}
```

Highlighting all models in turn
``` {r render-all-highlighted-fine-models}
p_fine_show_and_save(1)
p_fine_show_and_save(2)
p_fine_show_and_save(3)
p_fine_show_and_save(4)
p_fine_show_and_save(5)
p_fine_show_and_save(6)
p_fine_show_and_save(7)
p_fine_show_and_save(8)
```

Highlighting all coarsened models in turn
``` {r render-all-highlighted-coarse-models}
p_coarse_show_and_save(1)
p_coarse_show_and_save(2)
p_coarse_show_and_save(3)
p_coarse_show_and_save(4)
p_coarse_show_and_save(5)
p_coarse_show_and_save(6)
p_coarse_show_and_save(7)
p_coarse_show_and_save(8)
```




## Visualising coarsened graphs


Note the appearance of feedback loops - these are between S.Cp and S.Kn, but also S.Cp and H.AI.evil. These metacognitive skills are clearly seen as important feedback processes in this system.

``` {r}
pCoarseMerged + geom_node_text(aes(label = name, size = node_w, alpha = node_w), nudge_x = 1)
```

``` {r fig.height = 7}
plots_coarsend_ng <- 
    imap(
        tdags,
        \(g,n)
        g_plot_coarse(gcm_cluster_graph(g, code_node_group))
    ) 

pAllcoarse <- (plots_coarsend_ng$M1 | plots_coarsend_ng$M2) / (plots_coarsend_ng$M3 | plots_coarsend_ng$M4) / (plots_coarsend_ng$M5 | plots_coarsend_ng$M6) / ( plots_coarsend_ng$M7 | plots_coarsend_ng$M8)
# ggsave("dags_all_course.svg", pAllcourse, width = 18, height = 9)

pAllcoarse
```

### Figure: Consensus DAG

Most common node group patterns - nodes in at least half the models, then looking at the resulting graph of node-groups.

``` {r}
node_list_majority_of_models <- 
    all_nodes |> 
    count(name) |> 
    filter(n >= 4) |> 
    pull(name)

coarse_filtered_graph_temp <- 
    merged_all_graph |> 
    activate("nodes") |> 
    filter(name %in% node_list_majority_of_models)

coarse_filtered_graph <- 
    coarse_filtered_graph_temp |> 
    gcm_cluster_graph(code_node_group)

```


Weighting by num of models instead
This one uses only nodes that are in the majority of models, then coarsens the graph, then aggregates edges based on the number of models that they appear in. 
The edges must also appear in at least 2 models

``` {r}
node_list_majority_of_models <- 
    all_nodes |> 
    count(name) |> 
    filter(n >= 4) |> 
    pull(name)

coarse_filtered_graph_ng_filled <- 
    tbl_graph( # only nodes that are in most models (4 or more)
        nodes = all_nodes |> 
            filter(name %in% node_list_majority_of_models) |> 
            mutate( # but then coarsen into node groups
                node_group = code_node_group(name),
                   node_type = code_node_type(name)) |> 
            count(node_group, node_type) |> 
            rename(node_w = n),
        edges = all_edges |> # Only edges between node groups...
            mutate(across(from:to, code_node_group)) |>
            distinct(Model, from, to) |> 
            count(from, to) |>
            # filter(n > 1) |> # ...that occur in at least two models
            filter(from %in% code_node_group(node_list_majority_of_models)) |>
            filter(to %in% code_node_group(node_list_majority_of_models)) |>
            rename(edge_w = n)) |> 
    activate("edges") |> 
    switches_exist(from, to) |> 
    activate("nodes") |> 
    inner_join(node_group_nice_name)
```

All edges between node groups from the majority nodes

``` {r}
lyt_coarse_filtered_ng_full <- create_layout(
    coarse_filtered_graph_ng_filled, 
    layout = "sugiyama")

p_top_node_groups_full <- ggraph(lyt_coarse_filtered_ng_full) +
    geom_edge_fan(aes(edge_linetype = factor(switched), alpha = edge_w, ), # alpha = edge_w, 
                  strength = 0.8,
                  arrow = arrow(length = unit(2, 'mm'), type = "closed"),
                  start_cap = circle(3, 'mm'),
                  end_cap = circle(3, 'mm')) +
    geom_node_point(aes(colour = node_type, size = node_w)) +
    scale_color_discrete(palette = nt_pal, drop = FALSE) +
        # ggsci::scale_color_npg(drop = FALSE) +
        # geom_node_text(aes(label = name, size = node_w, alpha = node_w), nudge_y = -.5) +
        theme_graph() +
        scale_edge_alpha_continuous(range = c(0.15,0.8)) +
        scale_alpha_continuous(range = c(0.2,0.9)) +
        theme(legend.position = "none") +
    scale_size(range = c(3,6)) +
    geom_node_text(aes(label = nice_name, colour = node_type), 
                   alpha = 0.9, 
                   nudge_y = -.3,
                   repel = F) +
    xlim(c(
        min(lyt_coarse_filtered_ng_full$x) - 0.15, 
        max(lyt_coarse_filtered_ng_full$x) + 0.15))

p_top_node_groups_full

ggsave("figures/FIG--consensus-sub-model-full.jpg",
       p_top_node_groups_full,
       dpi = 900,
       height = 4, width = 6)
```


Filtered - edges must be in at least 2 models

``` {r}
lyt_coarse_filtered_ng_filtered <- create_layout(
    coarse_filtered_graph_ng_filled |> 
        activate("edges") |> 
        filter(edge_w > 1) |> 
        mutate(
            switched = map_lgl(row_number(), function(i) {
                from_i <- .E()$from[i]
                to_i   <- .E()$to[i]
                any(from == to_i & to == from_i)
            })
        ) , 
    layout = "sugiyama") |> 
    select(-x, -y) |> 
    inner_join(lyt_coarse_filtered_ng_full |> select(x, y, node_group))

p_top_node_groups_filtered <- ggraph(lyt_coarse_filtered_ng_filtered) +
    geom_edge_fan(aes(edge_linetype = factor(switched), alpha = edge_w, ), # alpha = edge_w, 
                  strength = 0.8,
                  arrow = arrow(length = unit(2, 'mm'), type = "closed"),
                  start_cap = circle(3, 'mm'),
                  end_cap = circle(3, 'mm')) +
    geom_node_point(aes(colour = node_type, size = node_w)) +
    scale_color_okabe_ito(order = c(9, 7,1, 2,3, 6, 8,5, 4), drop = FALSE) +
    # ggsci::scale_color_npg(drop = FALSE) +
    # geom_node_text(aes(label = name, size = node_w, alpha = node_w), nudge_y = -.5) +
    theme_graph() +
        scale_edge_alpha_continuous(range = c(0.15,0.8)) +
        scale_alpha_continuous(range = c(0.2,0.9)) +
        theme(legend.position = "none") +
    scale_size(range = c(3,6)) +
    geom_node_text(aes(label = nice_name, colour = node_type), 
                   alpha = 0.9, nudge_y = -.3, repel = F) +
    xlim(c(min(lyt_coarse_filtered_ng_filtered$x) - 0.15, max(lyt_coarse_filtered_ng_filtered$x) + 0.15))

p_top_node_groups_filtered

ggsave("figures/FIG--consensus-sub-model-filtered.jpg", 
       p_top_node_groups_filtered, 
       dpi = 900, 
       height = 4, width = 6)
```



``` {r}
# all nodes
all_nodes_with_metrics |> 
    inner_join(name_node_nice_name) |> 
    inner_join(node_group_nice_name |> rename(g_nice_name = nice_name)) |> 
    select(name, nice_name, g_nice_name, node_group, Model, pagerank_no_outcome, pagerank, btw_rel_rank) |> 
    arrange(name, Model) |> 
    knitr::kable()

```






## Node importance and measurability

``` {r}
all_node_groups_aggregated_metrics <- 
    all_nodes_with_metrics |> 
    group_by(node_group) |> 
    mutate(
        measurability_node_group = coarsen_node_measurability(measurability)
    ) |> 
    group_by(Model, node_group, node_type, measurability_node_group) |> 
    summarise(
        p_traced = mean(measurability == "DW"),
        measurability_metric = (3 * mean(measurability == "DW") +
                                    2 * mean(measurability == "O") +
                                    mean(measurability == "PO")) / 3,
        pagerank = sum(pagerank),
        pagerank_no_outcome = sum(pagerank_no_outcome),
        btw_rel = sum(btw_rel)) |> 
    group_by(node_group, node_type) |> 
    mutate(
        n = n(),
        overall_pagerank = sum(pagerank / 8),
           overall_pagerank_no_outcome = sum(pagerank_no_outcome / 8),
           overall_btw = sum(btw_rel)) |> 
    ungroup()
```

#### Figure: Node importance and data availability

``` {r}
d_ng_measurability_importance <- 
    all_node_groups_aggregated_metrics |> 
    filter(node_group != "Grade") |>
    inner_join(node_group_nice_name, by = "node_group") |>
    mutate(node_group = nice_name) |>
    group_by(node_group) |>
    mutate(m_group = mean(measurability_metric)) |>
    ungroup() |> 
    mutate(node_group = fct_reorder(node_group, pagerank_no_outcome)) |> 
    arrange(desc(node_group)) |> 
    filter(n > 3) # need at least half the models with node group


p_top_node_pagerank_no_outcome_measurability <- 
    d_ng_measurability_importance |> 
    # arrange(Model, node_group) |> 
    ggplot(aes(x = pagerank_no_outcome, y = node_group)) +
    geom_boxplot(aes(fill = m_group), colour = "lightgray", size = 0.4, alpha = 0.7) +
    # geom_line(aes(x = pagerank_no_outcome, y = node_group, group = Model), alpha = 0.6) +
    geom_point(aes(colour = measurability_metric), alpha = 0.9) +
    scale_color_gradient(
        low = "darkslateblue", 
        # mid = "darkturquoise",
        high = "aquamarine",
        # midpoint = mean(range(d_ng_measurability_importance$measurability)),
        limits = c(0,1),
        breaks = c(0, 1/3, 2/3, 1),
        labels = c("Latent","Potentially\nobservable", "Observable", "Data\nWarehouse"),
        name = "",
        guide = guide_colorbar(reverse = FALSE)
    ) +
    scale_fill_gradient(
        # midpoint = mean(range(d_ng_measurability_importance$m_group)),
        limits = c(0,1),
                low = "darkslateblue", 
        # mid = "darkturquoise",
        high = "aquamarine"
        ) +
    theme_minimal() +
    ylab("Node group") +
    xlab("Node groups combined Centrality (Pagerank)") +
    # geom_rug(aes(colour = m_dist_to_trace), alpha = 0.6, sides = "b") +
    # coord_cartesian(clip = "off") +
    guides(fill = "none") +
    theme(legend.position = "bottom",
          legend.justification = "center",
    legend.box.just = "center",
          legend.text  = element_text(size = 8),
          legend.key.width = unit(1, "cm"),
          axis.ticks.y = element_blank(),
          axis.line.y = element_blank(),
          panel.grid.major.y = element_blank(),
          panel.grid.minor.y = element_blank())

# p_top_node_pagerank_no_outcome_measurability

# extract legend
leg <- get_legend(p_top_node_pagerank_no_outcome_measurability)

# arrange with relative heights (legend gets a small slice)
p_node_importance_with_measurability <- (wrap_elements(leg) /
    (p_top_node_pagerank_no_outcome_measurability + theme(legend.position = "none"))) +
  plot_layout(heights = c(0.2, 1))

# p_node_importance_with_measurability
# ggsave("figures/FIG--node_pagerank_no_outcome_measurability.png", p_node_importance_with_measurability,
#        width = 5,
#        height = 4,
#        dpi = 900)

```

With discrete scale instead

``` {r}
m_pal <- c("darkslateblue",
    "dodgerblue","orange","deeppink")
    

p_top_node_pagerank_no_outcome_measurability_discrete <- 
    d_ng_measurability_importance |>
    mutate(
        measurability = case_when(
            measurability_metric < (1/6) ~ "Latent", 
            measurability_metric < 0.5 ~ "Potentially observable",
            measurability_metric < (5/6) ~ "Observable",
            TRUE ~ "Data warehouse") |> 
            factor(levels = c(
                "Latent", 
                "Potentially observable", 
                "Observable", 
                "Data warehouse"
                ))) |> 
    arrange(Model, node_group) |>
    ggplot(aes(x = pagerank_no_outcome, y = node_group)) +
    geom_boxplot(aes(fill = m_group), colour = "lightgray", size = 0.4, alpha = 0.7) +
    # geom_line(aes(x = pagerank_no_outcome, y = node_group, group = Model), alpha = 0.6) +
    geom_point(aes(colour = measurability), alpha = 0.9) +
    scale_color_manual(values = m_pal, # or retro four
                       labels = c("Latent","Potentially\nobservable", "Observable", "Data\nWarehouse"),
                       name = ""
    ) +
    scale_fill_gradientn(
        colours = m_pal,
        breaks = c(0, 1/3, 2/3, 5/6)
        ) +
        theme_minimal() +
            ylab("Node group") +
            xlab("Node groups combined Centrality (Pagerank)") +
            # geom_rug(aes(colour = m_dist_to_trace), alpha = 0.6, sides = "b") +
            # coord_cartesian(clip = "off") +
            guides(fill = "none",
                   color = guide_legend(nrow = 2, byrow = FALSE)) +
            theme(legend.position = "bottom",
          legend.direction = "vertical",
                  legend.justification = "left",
                  legend.text  = element_text(size = 8),
                  legend.key.width = unit(1, "cm"),
                  axis.ticks.y = element_blank(),
                  axis.line.y = element_blank(),
                  panel.grid.major.y = element_blank(),
          panel.grid.minor.y = element_blank())
    
p_top_node_pagerank_no_outcome_measurability_discrete

# 
# # extract legend
# leg <- get_legend(p_top_node_pagerank_no_outcome_measurability_discrete)
# 
# # arrange with relative heights (legend gets a small slice)
# p_node_importance_with_measurability <- (wrap_elements(leg) /
#     (p_top_node_pagerank_no_outcome_measurability_discrete + theme(legend.position = "none"))) +
#   plot_layout(heights = c(0.2, 1))


ggsave("figures/FIG--node_pagerank_no_outcome_measurability.jpg",
       p_top_node_pagerank_no_outcome_measurability_discrete, 
       width = 5, 
       height = 4, 
       dpi = 900)
```

As table

``` {r}
all_nodes_with_metrics |> 
    distinct(name, measurability) |> 
    janitor::tabyl(measurability) |> 
    janitor::adorn_totals("row")
```




#### Table: Node lists, with types, groups and importance metrics

Node group measurability was coded according to:

- Observed: All nodes were either in the data warehouse or observable. 
- Observable: No nodes were latent.
- Likely latent: Nodes were a mix of latent or potentially observable. 
- Latent: All nodes were latent.

``` {r appendix-node-groups-table}
# could add distance to traced
node_groups_table <- 
    all_nodes_with_metrics |>
    inner_join(name_node_nice_name, by = "name") |> 
    mutate(name =  nice_name) |> 
    left_join(
        node_group_nice_name |> 
            rename(ng_nice_name = nice_name), by = "node_group") |> 
    mutate(node_group = ng_nice_name) |> 
    group_by(node_type, node_group) |>
    mutate(measurability = coarsen_node_measurability(measurability)) |> 
    group_by(node_type, node_group, measurability) |>
    summarise(
        nodes = str_c(unique(name), collapse = "; "),
        n = sum(node_w),
        pagerank = round(sum(pagerank_no_outcome) / 8, 2)
        # ,btw = round(sum(btw_rel), 3)
    ) |> 
    mutate(pagerank = if_else(node_group == "Assessment grade", NA, pagerank)) |> 
    # mutate(node_type = case_when(
    #     node_type == "L" ~ "Learning Design",
    #     node_type == "S" ~ "Student",
    #     node_type == "T" ~ "Teacher",
    #     node_type == "C" ~ "Course Data",
    #     node_type == "G" ~ "Grade Data",
    #     node_type == "H" ~ "Help"
    # )) |> 
    arrange(desc(n), desc(pagerank)) |> 
    select(node_type, node_group, nodes, n, measurability, pagerank) |> 
    ungroup()

node_groups_table |> 
    arrange(node_type, node_group)
```

``` {r}
all_nodes_with_metrics |> 
    distinct(name) |> 
    nrow() |> 
    paste("distinct nodes")

all_nodes_with_metrics |> 
    distinct(node_group) |> 
    nrow() |> 
    paste("node groups")

all_nodes_with_metrics |> 
    distinct(name, measurability) |> count(measurability)
```

What about grouping by the node category first, so it's hierarchical?

``` {r}
node_types_table <- 
    node_groups_table |> 
    group_by(node_type) |> 
    mutate(nt = sum(n)) |> 
    ungroup() |> 
    arrange(desc(nt), desc(n)) |> 
    select(node_type, nt, node_group, nodes, n, measurability, pagerank) |> 
    mutate(node_type = str_c(node_type, " (", nt, ")")) |> 
    select(-nt)

node_types_table
```


# Causal consistency between models

## Informal consistency: Edge coherence

How often do causal paths point in the same direction? To examine this across all models we must look at relationships between 

``` {r}
n_possible_edges <- nrow(all_possible_edges)
n_edges <- nrow(all_edges_in_models_with_metrics)

prob_universal_inclusion_given_existence <- 
    mean(all_edges_in_models_with_metrics$p_direct == 1)

prob_universal_descendant_inclusion_given_existence <- 
    mean(all_edges_in_models_with_metrics$p_descendant == 1)

prob_universal_descendant_inclusion_given_descendant_existance <- 
    all_edges_sum |> 
    filter(p_descendant > 0) |> 
    mutate(universal_descendant = p_descendant == 1) |> 
    pull(universal_descendant) |> 
    mean()
    
```

We examine the `r n_edges` edges from `r n_possible_edges` possible edges across the 8 models, and in addition include the ancestor - descendant pairs. From this we define a few events:

- $E_{ij}\in M$ for and edge belonging to a model
- $E_{ij} \in \mathcal{M}$

- $P(E_{ij} \in M_k|i,j\in V(M_k), E_{ij} \in \mathcal{E})$ - the probability of universal inclusion of an edge in all models it could be included in, given it appears in at least one model. This is `r prob_universal_inclusion_given_existence`
- $P(De_{ij} \in M_k|i,j\in V(M_k), E_{ij} \in \mathcal{E})$ - the probability of universal inclusion of a descendant in all models it could be included in, given the edge appears in at least one model. This is `r prob_universal_descendant_inclusion_given_existence`. Could maybe use the squiggle notation here of mediated path?


## Causal $\mathcal{L}-consistancy$

THIS BIT NEEDS TO BE RE-WRITTEN FOR THE NEW CODE

L1 (association) and L2 (intervention / do) levels on the causal hierarchy can be examined at a graph level through:

- $\mathcal{L}_1$ Examining all conditional probability implications. If these are preserved between graphs then the two graphs are $\mathcal{L}_1$-consistent. ? An equivalent may be having similar / same CPDAGs - complete partial DAGs - which are the Markov equivalent class of DAGs. These induce the same conditional independencies (Thm 1.2.4 in Pearl 2009). These are the skeleton + all v-structures (colliders) arrowed in.
- $\mathcal{L}_2$ Examining all 'do' operator implications. If these are preserved between graphs then the two graphs are $\mathcal{L}_2$-consistent. Or - use the Structural Intervention Distance (Peters & Buhlmann, 2015)
- $\mathcal{L}_3$ Is not explored - this would probably require a SCM, not just the DAG.

## Causal Consistency when Clustering

### L1 consistency when clustering

This computes two metrics, the loss ratio of conditional independence relationships (ci_loss) and the rate of introduction of spurious ancestors when coarsening. CI loss indicates how often a conditional independence relationship in the fine graph did not have its analogue in the coarsened graph. Spurious ancestors indicates the rate at which merged nodes introduced new ancestors in the cluster DAG - violating L1 consistency (CITE - Annan???)

``` {r L1-consistency-of-clustering}
clustering_L1_consistency <- 
    tdags |> 
    imap(
        \(g, m)
        gcm_cluster_L1_consistency(
            g, code_node_group)$clustering_L1_consistency
    ) |> 
    list_rbind(names_to = "Model")
```

``` {r}
clustering_L1_consistency |> 
    knitr::kable(
        caption = "L1-consistency metrics when coarsening GCMs",
        digits = 4)
```

### L2 consistency when clustering

This computes a single error metric - the % of lost identifiable pairs when coarsening. So this is where the $P(Y|do(X))$ is identifiable in the fine graph but $P(\varphi(Y)|do(\varphi(X)))$ is not identifiable in the coarsened cluster graph. 

``` {r L2-consistency-of-clustering}
clustering_L2_consistency <- 
    tdags |> 
    imap(
        \(g,m)
        gcm_cluster_L2_consistency(g, code_node_group)) |> 
    list_rbind(names_to = "Model")
```

``` {r}
clustering_L2_consistency |> 
    knitr::kable(
        caption = "L2-consistency metrics when coarsening GCMs",
        digits = 4)
```

## Causal consistency between models

#### Computing L1, L2 distances between models

``` {r dist-matrix-helper-funcion}
compute_dist_matrix <- function(
        model_list, # list of models
        .df, # function that computes distance between two models
        symmetric = FALSE,
        impute_undefined = FALSE) {
    n  <- length(model_list)
    nm <- names(model_list) 
    D <- matrix(0, n, n, dimnames = list(nm, nm))
    
    if (symmetric) {
        # fill upper triangle, mirror, zero diagonal
        for (i in seq_len(n - 1)) {
            for (j in (i + 1):n) {
                d <- .df(model_list[[i]], model_list[[j]])
                D[i, j] <- d
                D[j, i] <- d
            }
        }
        Dsym <- D
    } else {
        for (i in seq_len(n)) {
            for (j in seq_len(n)) {
                if (i != j) {
                    D[i, j] <- .df(
                        model_list[[i]], 
                        model_list[[j]])
                }
            }
        }
        Dsym <- (D + t(D)) / 2
    }
    
    if (impute_undefined) {
        # First fill Dsym on mirror
        fill <- is.nan(Dsym) & !is.nan(t(Dsym)) 
        Dsym[fill] <- t(Dsym)[fill]
        if (any(is.nan(D))) {
            message("Imputing missing values")
            fit <- softImpute::softImpute(D, rank.max = 4, lambda = 0)
        D <- complete(D, fit)
        }
        if (any(is.nan()))
        fitSym <- softImpute::softImpute(Dsym, rank.max = 4, lambda = 0)
        Dsym <- complete(Dsym, fit)
    } 
    
    return(list(D = D, Dsym = Dsym))
}

```

#### A note on clustering methods

A note on clustering: Some methods of `hclust` have underlying geometric assumptions (assume distances are Euclidean, which they are not), and these are `ward.D`, `ward.D2`, `centroid` (UPGMC) and `median` (WPGMC). Of the remainder `average` and `complete` are the best candidates. There is another challenge when using these - the distance matrix must be symmetric and in the case of the L2 distances it is not. So a decision also needs to be made about how to do this along with the algorithm for the hierarchical clustering. 

- `average` (UPGMA) clustering method uses the mean pairwise distance between clusters, and the default for a non-Euclidean distance matrix. It is common when using Jaccard / dissimilarity. This works naturally with a matrix symmetrisation process of taking the average of the cross diagonals.  
- `complete` is a stronger method that looks for clusters where every pair within it is consistant, no exceptions. This is a better 'worst case' scenario, and works more naturally with taking the maximum value of the cross-diagonals when forcing the distance matrix to be symmetric. 

As such two methods of visualising the between model distances are used: **average** and **worst-case**.

``` {r between-distances-plotting}
library(ggdendro)

fmt_lab <- function(x) {
    # split leading letters from trailing digits: "M1" -> "M", "1"
    highlight_col = rgb(87, 180, 66, maxColorValue = 255) # mygum
    # letter <- sub("([A-Za-z]+).*", "\\1", x)
    # num    <- sub("[A-Za-z]+", "", x)
    # ifelse(
    #     startsWith(x, "D"),
    #     sprintf("<span style='background-color:%s'>%s<sub>%s</sub></span>", highlight_col, letter, num),
    #     sprintf("%s<sub>%s</sub>", letter, num)   # non-M: subscript only, no highlight
    # )
    return(x)
        # str_c(
        #     str_extract(x, "^.{1}"),
        #     "<sub>",
        #     str_extract(x, ".{1}$"),
        #     "</sub>"
        # )
}

plotting_Ln_between_distances <- function(
        D, level_label = "L", leg_pos = "none",
        clustering_worst_case = FALSE
        ) {
    nm <- D |> colnames()
    
    if (clustering_worst_case) {
        D_sym <- pmax(D, t(D))
        hc  <- hclust(as.dist(D_sym), method = "complete")
    } else {
        D_sym <- (D + t(D)) / 2
        hc  <- hclust(as.dist(D_sym), method = "average")
    } 
    
    ord <- hc$order
    lev <- nm[ord]  

    # --- dendrogram panel ---
    dend <- dendro_data(hc)
    
    p_dend <- ggplot() +
    geom_segment(
        data = segment(dend),
        aes(x = x, y = y, xend = xend, yend = yend),
        linewidth = 0.4, colour = "grey30"
    ) +
    scale_x_continuous(
        breaks = seq_along(lev),
        labels = lev,
        limits = c(0.5, length(lev) + 0.5),
        expand = expansion(add = 0),
        name = "Model"
    ) +
    scale_y_continuous(
        expand = expansion(mult = c(0, 0.02))
    ) +
    theme_minimal() +
    coord_flip() +
    theme(
        axis.title  = element_blank(),
        axis.text.x = element_blank(),
        axis.ticks  = element_blank(),
        panel.grid  = element_blank()
    )
    
    D_long <- as_tibble(D_sym, rownames = "from") |>
        pivot_longer(-from, names_to = "to", values_to = "dist") |>
        mutate(
            from = factor(from, levels = lev),
            to   = factor(to,   levels = lev)
        )
    
    p_heat <- ggplot(D_long, aes(to, from, fill = dist)) +
        geom_tile(colour = "white", linewidth = 0.5) +
        geom_text(
            aes(label = round(dist, 2)), 
            size = 3, colour = "grey15") +
        scale_fill_viridis_c(
            name = str_c(level_label, "-consistency\ndiscrepency"), 
            direction = -1) +
        scale_x_discrete(position = "top", labels = fmt_lab) +
        labs(x = NULL, y = NULL) +
        coord_fixed() +
        theme_minimal(base_size = 12) +
        theme(
            # axis.text.x = element_text(angle = 45, hjust = 0),
            panel.grid  = element_blank(),
            legend.position = leg_pos
        )
    
    # combined 
    combined <- p_heat + p_dend

    return(list(
        heatmap = p_heat,
        dendrogram = p_dend,
        combined.plot = combined,
        hclust = hc
    ))

}
```

### L1 consistency between models

``` {r}
# scalar L1 distance between two models
L1_dist <- function(m_i, m_j) {
  fuzzyL1 <- gcm_fuzzy_L1_consistency(
      gX = m_i, gY = m_j)
  fuzzyL1$distance
}

L1_between_model_metrics <- 
    compute_dist_matrix(tcdags_node_group, L1_dist, symmetric = TRUE)

plots_L1_similarities <- plotting_Ln_between_distances(
    L1_between_model_metrics$D, level_label = "Average\nL1", leg_pos = "left")
plots_L1_similarities_worst_case <- plotting_Ln_between_distances(
    L1_between_model_metrics$D, level_label = "Maximal\nL1", clustering_worst_case = TRUE,leg_pos = "left")

```

``` {r}
plots_L1_similarities$combined.plot
```

``` {r}
plots_L1_similarities_worst_case$combined.plot
```

### L2 consistency between models

``` {r}
# L2 distance between two models
L2_dist <- function(m_i, m_j) {
  fuzzyL2 <- gcm_fuzzy_L2_consistency(
      gX = m_i, gY = m_j)
  return(fuzzyL2$dXY)
}

L2_between_model_metrics <- 
    compute_dist_matrix(
        tcdags_node_group, L2_dist, symmetric = FALSE)

```

``` {r}
plots_L2_similarities <- plotting_Ln_between_distances(
    L2_between_model_metrics$D, 
    level_label = "Average\nL2",
    leg_pos = "left"
)
plots_L2_similarities_worst_case <- plotting_Ln_between_distances(
    L2_between_model_metrics$D, 
    clustering_worst_case = TRUE,
    level_label = "Maximal\nL2",
    leg_pos = "left"
)
```

``` {r}
plots_L2_similarities$combined.plot
```


``` {r}
plots_L2_similarities_worst_case$combined.plot
```


#### Combined L1, L2 between models

``` {r}
plots_L12_similarities_avg_maximal <- 
    plotting_Ln_between_distances(
        (L1_between_model_metrics$D + L2_between_model_metrics$D) / 2,
        clustering_worst_case = TRUE,
        level_label = "Maximal\nL",
        leg_pos = "none"
    )
```

``` {r}
plots_L12_similarities_avg_maximal$combined.plot
```

#### Clusters of cluster DAGs

It seems that models M1, M2 (and maybe M4) are close to each other (GC1), as well as M3, M6 and M7 (GC2) and M5 and M8 (GC3).

``` {r}
# thought - merge on latent projection of shared nodes only?
ng_counts <- 
    all_node_groups_aggregated_metrics |> 
    count(node_group)

gcm_latent_projection_merge <- function(model_list, shared_nodes = NULL) {
    # merges the latent projection of a list of models
    # each model is projected onto the shared set of nodes, 
    # with the other nodes being treated as latent
    if (is.null(shared_nodes)) {
        shared_nodes <- 
            model_list |> 
            map(
                \(m)
                gcm_nodelist(m) |> 
                    pull(name) 
            ) |> 
            reduce(intersect)
    }
    
    merged_projections <- 
        model_list |> 
        map(
            \(m)
            gcm_latent_projection(m, latent_nodes = 
                                      setdiff(gcm_nodelist(m) |> pull(name),
                                              shared_nodes)) |> 
                activate(edges) |> 
                mutate(edge_w = replace_na(edge_w, 1)) |> 
                activate(nodes)
        ) |> 
        reduce(gcm_merge)
    
    return(merged_projections)
}

tcdags_node_group |> 
    gcm_latent_projection_merge(
        shared_nodes = ng_counts |> 
            filter(n > 3) |> 
            pull(node_group)) |>
    activate(edges) |> 
    filter(edge_w > 2) |> 
    g_plot_coarse()

cdagGC1 <- tcdags_node_group[c(1,2,4)] |> gcm_latent_projection_merge()
cdagGC2 <- tcdags_node_group[c(3,7,6)] |> gcm_latent_projection_merge()
cdagGC3 <- tcdags_node_group[c(5,8)] |> gcm_latent_projection_merge()

g_plot_coarse(cdagGC1)
# g_plot_coarse(cdagGC2) |> activate(edges) |> filter(edge_w > 0)
# g_plot_coarse(cdagGC3) |> activate(edges) |> filter(edge_w > 1))
```


# Simulation


Now we examine the feasibility of learning these models from data. A typical counter to going to the effort of eliciting these models from domain experts is that they can be learned from the data alone. However, we have seen here that much of the data is not present in the institutions ecosystem, with the nodes being deemed inherently latent (`r round(100*(sum(all_nodes_with_metrics$measurability == "L") / nrow(all_nodes_with_metrics)), 1)`%), potentially observable (`r round(100*(sum(all_nodes_with_metrics$measurability == "PO") / nrow(all_nodes_with_metrics)), 1)`%), observable but not currently in the data (`r round(100*(sum(all_nodes_with_metrics$measurability == "O") / nrow(all_nodes_with_metrics)), 1)`%), or already measured and stored in the current data ecosystem (`r round(100*(sum(all_nodes_with_metrics$measurability == "DW") / nrow(all_nodes_with_metrics)), 1)`%).

Examining at a coarser grained level by looking at the node group cluster graphs, we still see that a majority of node groups fall into difficult to measure categories:

``` {r}
node_groups_table |> 
    janitor::tabyl(measurability) |> 
    janitor::adorn_pct_formatting() |> 
    arrange(measurability)
```

## Creating a simulated data set

We design this experiment as follows. We ask the question can a causal graph discovered from available data generate similar answers to the problem of assuring assessment? This translates to getting an unbiased estimate of Student Knowledge on Grade. As such 3 of the models are excluded, lacking sufficient adjustment sets (2 models) or having broken student knowledge into multiple nodes (1 model). 
For the "potentially observable" nodes a single descendant node is added, so that $N \rightarrow N_P$, a proxy measurement node that is not the exact node measured but whose value depends only on the potentially observable node and some additional noise. The set of proxy nodes for the potentially observable ($PO$) nodes we denote ($\hat{PO}$). 
Each of the 5 remaining GCMs, with their additional proxy measurement nodes, is used to generate a synthetic data set. We assuming each node has an equal linear effect on its descendants ($\beta=0.4$), generated using `simulateSEM` of the `dagitty` package in R. This treats all variables as continuous with mean 0 and variance 1. 


``` {r}
sim_signal_strength <- list(
    "idyllic" = 0.9, 
    "typical" = 0.4, 
    "noisy" = 0.1)

tibble(Scenario = names(sim_signal_strength),
       Beta = as.numeric(sim_signal_strength)) |>
    mutate(
       `Variance due to each parent node` = Beta^2 / (1 + Beta^2)
       ) |> 
    knitr::kable(digits = 3)
```

We then apply a structure learning algorithm to each models simulated data in three phases: learning from the currently measured data only (only DW nodes), learning from all observable nodes (DW and O) nodes, and finally learning from all potentially available data (DW, O and $\hat{PO}$). This represents increasing levels of resourcing applied to learning the underlying causal structure, given the model. 

Note we are working under a local assumption that each model is the 'ground truth', but fully acknowledge that we do not know for this system what the true model is. However, if the structure learning claims of application to discover latent nodes are to be believed (CITE) then we should see some reasonable convergence towards each locally true model. 


``` {r simulate-data-from-dags}
# maybe use dagitty::simulateSEM() or dagitty::simulateLogistic()
# dagitty suggests using lavaan instead for SEM

# should add layout!!
# use same x, y from main graph

m_build_node_group_model_with_proxy_nodes <- function(
        m = "M1", 
        x_p_nudge = 0, y_p_nudge = -2) {
    model <- m

    nodes_in_graphs <- 
        all_node_groups_aggregated_metrics |>
        mutate(measurability = case_when(
            # Simplifying for sim
            node_group == "S.Kn" ~ "O", # For simming only
            str_detect(measurability_node_group, "atent") ~ "L",
            str_detect(measurability_node_group, "Potent") ~ "PO",
            TRUE ~ "O",
        )) |> 
        mutate(name = node_group) |> 
        inner_join(merged_coarse_layout |> 
                       select(name, x, y),
                   by = "name") |> 
        filter(Model == toupper(m)) |> 
        select(name, x, y, node_group, node_type, 
               measurability)
    
    proxy_nodes <- 
        nodes_in_graphs|> 
                filter(measurability == "PO") |> 
                select(name, node_type, x, y) |> 
                mutate(name = str_c(name, "_p"),
                       # Need to place proxy nodes close - might cause issues
                       x = x + x_p_nudge,
                       y = y + y_p_nudge  
                       ) |> 
                mutate(node_group = "Proxy") |> 
                mutate(measurability = "O") # the assumption for the proxy
    
    nodes <- nodes_in_graphs |> 
        bind_rows( # adding proxy nodes for PO nodes
            proxy_nodes
        ) |> 
        mutate(measurability = factor(
            measurability, 
            levels = c("L", "PO", "O")))
    
    edges <- tcdags_node_group[[m]] |>
        gcm_edgelist() |> 
        bind_rows(
            tibble(
                from = nodes |> 
                    filter(measurability == "PO") |> 
                    pull(name)) |> 
                mutate(to = str_c(from, "_p"))
        )
    
    new_tidydag <- tidygraph::tbl_graph(
        nodes = nodes, edges = edges, directed = TRUE
    )
    new_dag <- dagitty_from_edge_df(edges)
    
    # setting coordinates of dagitty object - needs to be a named list
    node_names <- nodes$name
    node_coord_list <- list(x = nodes$x, y = nodes$y)
    names(node_coord_list$x) <- node_names
    names(node_coord_list$y) <- node_names
    coordinates(new_dag) <- node_coord_list
    
    node_list_o <- 
        nodes |> 
        filter(measurability == "O") |> 
        filter(!str_detect(name, "_p")) |> 
        pull(name)
    node_list_opo <- nodes |>  
        filter(measurability %in% c("DW", "O")) |> 
        pull(name)
    
    return(list(
        dag = new_dag, # new dag in dagitty format
        tdag = new_tidydag, # new dag as tidydag
        node_list_o = node_list_o, # list of observable nodes (excluding proxies)
        node_list_opo = node_list_opo # list of all non-latent nodes
        ))
}



simSEM <- partial(simulateSEM, eps = 1, standardized = F, N = 1000)
simSEM <- partial(simulateLogistic, eps = 1, N = 1000)


bn_fit_node_group_model_list <- function(
        m_proxy_list, Beta = 0.6,
        blacklist = tibble(
            from = "Grade", 
            to = "S.Kn")) {
    
    # getting simulated data, and subsets
    d.sim.sem <- simSEM(m_proxy_list$dag, b.default = Beta)
    d.o <- as.data.frame(d.sim.sem[,m_proxy_list$node_list_o])
    d.opo <- as.data.frame(d.sim.sem[,m_proxy_list$node_list_opo])
    
    # fitting BNets
    bn.full =  bnlearn::bn.fit(bnlearn::hc(d.sim.sem,
                                           blacklist = blacklist), d.sim.sem)
    bn.o =    bnlearn::bn.fit(bnlearn::hc(d.o,
                                           blacklist = blacklist), d.o)
    bn.opo = bnlearn::bn.fit(bnlearn::hc(d.opo,
                                           blacklist = blacklist), d.opo)
    
    return( # returns a named list of simmed data and models
        list(
            # The full data set
            d.full   = d.sim.sem,
            bn.full  = bn.full,
            # Only observable nodes
            d.o     = d.o,
            bn.o    = bn.o,
            # Including Potentially Observable (PO), so ALL non-latent nodes
            d.opo  = d.opo,
            bn.opo = bn.opo
    ))
}


bn_to_dagitty <- function(bn) {
  edges <- bnlearn::arcs(bn) |> 
    as_tibble() |> 
    mutate(edge_str = str_c(from, " -> ", to))
  
  dagitty(str_c(
    "dag{",
    str_c(edges$edge_str, collapse = "; "),
    "}"
  ))
}

# Thinking of a simple dag X -> Y, we have 
# Var(Y) = b^2 * Var(X) + Var(eps) -- eps is the noise, and here it is set to 1 by default
# then Var(Y) = b^2 + 1
# Since the R^2 is the ratio of the variance explained by X:
# R^2 = Var(explained) / Var(total)
# R^2 = b^2 / (b^2 + 1)

dagitty_to_bn <- function(dag) {
  
  edges <- dagitty::edges(dag) |> 
    as_tibble() |> 
    filter(e == "->") |>  # directed edges only, excludes bidirected
    select(v, w) |> 
    rename(from = v, to = w)
  
  nodes <- unique(c(edges$from, edges$to))
  
  # Build empty network then set arcs
  net <- bnlearn::empty.graph(nodes)
  bnlearn::arcs(net) <- as.matrix(edges)
  net
}


causal_estimate_over_all_adj_sets <- function(
        dag, data, latent_nodes = NULL,
        exposure = "S.Kn", outcome = "Grade",
        x1_val = 1, x0_val = -1) {
    # This estimates P(Grade|S.Kn, Adjustment Set) over all adjustment sets
    # This will give a range of that the model suggests is the true effect
    adj_sets <- adjustmentSets(dag, exposure = exposure, outcome = outcome,
                               type = "all")
    if (!is.null(latent_nodes)) {
        # remove any with latent nodes 
        adj_sets <- keep(adj_sets, \(x) !any(x %in% latent_nodes))
    }
    # Throw error if none!
    estimate_per_set <- map(adj_sets, function(adj_set) {
        if (length(adj_set) == 0) {
            message("No adjustment sets found")
            formula_str <- str_c(outcome, " ~ ", exposure)
        } else {
            formula_str <- str_c(outcome, " ~ ", exposure, " + ",
                                 str_c(adj_set, collapse = " + "))
        }
        if (class(data[[1]]) == "factor") {
            fit <- glm(as.formula(formula_str), data = data, family = binomial)
            x1_data <- data |> 
                mutate({{exposure}} := factor(x1_val, levels = c(-1,1)))
            x0_data <- data |> 
                mutate({{exposure}} := factor(x0_val, levels = c(-1,1)))
        } else {
            fit <- lm(as.formula(formula_str), data = data)
            x1_data <- data |> 
                mutate({{exposure}} := x1_val)
            x0_data <- data |> 
                mutate({{exposure}} := x0_val)
        }
        # Simulating for potential outcomes
        y1_x1 <- predict(fit,
                     newdata = x1_data,
                     se.fit = TRUE)
        y1_x0 <- predict(fit,
                     newdata = x0_data,
                     se.fit = TRUE)
        
        preds.fit <- y1_x1$fit - y1_x0$fit
        se1 <- sqrt(mean(y1_x1$se.fit^2))
        se0 <- sqrt(mean(y1_x0$se.fit^2))
        se <- max(se1, se0)
        
    list(
        y1_x0 = y1_x0,
        y1_x1 = y1_x1,
        eff_mean = mean(preds.fit),
        eff_se   = se
    )
    }
    )
    estimates <- map_dbl(estimate_per_set, "mean")
    std_errors <- map_dbl(estimate_per_set, "se")
    return(list(adjustment_set = adj_sets,
                estimate = estimates,
                std_error = std_errors))
}

generate_node_group_causal_estimate_comparisons <- function(
        Model = "M1",
        Beta = 0.6, # strength of signal along paths
        # might need to change for different models depending on present nodes
        exposure = "S.Kn", outcome = "Grade",
        x1_val = 1, x0_val = -1, # should be 1 and 0 for continuous
        # for placiing proxy nodes
        x_p_nudge = 0, y_p_nudge = -2
) {
    ground_truth_dag <- 
        tcdags_node_group[[Model]] |> 
        tidy_graph_to_dag()
    
    model_list <- 
        m_build_node_group_model_with_proxy_nodes(
        m = Model, 
        x_p_nudge = x_p_nudge, 
        y_p_nudge = y_p_nudge)
    
    bn_list <- bn_fit_node_group_model_list(model_list, Beta = Beta)
    # Ground truth uses full simmed data
    est <- partial(causal_estimate_over_all_adj_sets,
                   exposure = exposure, outcome = outcome,
                   x1_val = x1_val, x0_val = x0_val)
    
    ground_truth <- est(dag = ground_truth_dag, data = bn_list$d.full)
    expert_dag <- est(
        dag = ground_truth_dag, 
        data = bn_list$d.full,
        latent_nodes = 
            all_node_groups_aggregated_metrics |> 
            filter(Model == Model, 
                   str_detect(measurability_node_group, "atent")) |> 
            pull(node_group))
    can_estimate_o <- 
        (exposure %in% model_list$node_list_o) & (outcome %in% model_list$node_list_o)
    can_estimate_opo <- 
        (exposure %in% model_list$node_list_opo) & (outcome %in% model_list$node_list_opo)

    if (can_estimate_o) {
        est.o <- est(   
            dag = bn_to_dagitty(bn_list$bn.o) , 
            data = bn_list$d.o)
    } else {
        est.o <- list("adjustment_set" = NA, "estimate" = NA)
        message("Nodes not found in Observable node list")
    }
    
    if (can_estimate_opo) {
        est.opo <- est(  
            dag = bn_to_dagitty(bn_list$bn.opo) , 
            data = bn_list$d.opo)
    } else {
        est.opo <- list("adjustment_set" = NA, "estimate" = NA)
        message("Nodes not found in Potentially Observable node list")
    }
    
    estimates <- tibble(
        Model = Model,
        DAG = c(
            "Ground Truth", 
            "Expert DAG",
            "Observable", 
            "Potentially Observable") |> 
            factor(levels = c(
                "Ground Truth", 
                "Expert DAG", 
                "Observable", 
                "Potentially Observable")),
        exposure = exposure, 
        outcome = outcome,
        adjustment_sets = list(
            ground_truth$adjustment_set,
            expert_dag$adjustment_set,
            est.o$adjustment_set,
            est.opo$adjustment_set
        ),
        estimate_mean = c(
            mean(ground_truth$estimate),
            mean(expert_dag$estimate),
            mean(est.o$estimate),
            mean(est.opo$estimate)
        ),
        estimate_std_errors = c(
            mean(ground_truth$std_error),
            mean(expert_dag$std_error),
            mean(est.o$std_error),
            mean(est.opo$std_error)
        ),
        estimate_sd = c(
            sd(ground_truth$estimate),
            sd(expert_dag$estimate),
            sd(est.o$estimate),
            sd(est.opo$estimate)
        )
    )
    
    return( # Monster list of everything we need
        list(
            Model = Model,
            DAG = ground_truth_dag,
            augmented_DAGs = model_list,
            fitted_BN = bn_list,
            estimates = estimates
        )
    )
}
```

``` {r simulating-cluster-models}
simulations_ng_cluster_models <- 
    c("M1", "M2", "M3", "M4", "M5", "M6", "M7", "M8") |> 
    set_names() |> 
    imap(
        \(M, n)
        generate_node_group_causal_estimate_comparisons(
            Model = M,
            Beta = sim_signal_strength[["typical"]]
        )
    )
```

``` {r}
all_ng_estimates <-
    map(
        simulations_ng_cluster_models,
        \(s)
           s$estimates
        ) |> 
    list_rbind() |> 
    group_by(Model) |> 
    mutate(
        gt_est_diff = estimate_mean - estimate_mean[DAG == "Ground Truth"] )|> 
    ungroup() |> 
    mutate(error = estimate_std_errors + replace_na(estimate_sd, 0)) |> 
    select(Model, DAG, estimate_mean, error, gt_est_diff, everything())
```

## Simulation comparisons

Between Ground Truth, Expert DAGs, and Learned DAGs from observable data. 

``` {r plotting-cluster-model-simulation-results}
dodge_amt <- 0.1

ground_truth_estimate_mean_estimate <- 
    all_ng_estimates |> 
    filter(DAG == "Ground Truth") |> 
    filter(!is.nan(estimate_mean)) |> 
    pull(estimate_mean) |> 
    mean()

ground_truth_error_estimate <- 
    ( # the largest base error
        all_ng_estimates |> 
    filter(DAG == "Ground Truth") |> 
    filter(!is.nan(error)) |> 
    pull(error) |> 
    max())
    # ) + ( # as well as the difference in ground truth estimates
    #     all_ng_estimates |> 
    # filter(DAG == "Ground Truth") |> 
    # filter(!is.nan(estimate_mean)) |> 
    # pull(estimate_mean) |> 
    # range() |> 
    #     diff()
    # )


est_to_plot <- 
    all_ng_estimates |>
    mutate(estimated_ground_truth = is.nan(estimate_mean)) |> 
    mutate(
        estimate_mean = if_else(is.nan(estimate_mean) & DAG == "Ground Truth",
                                ground_truth_estimate_mean_estimate,
                                estimate_mean),
        error = if_else(is.nan(error) & DAG == "Ground Truth",
                                ground_truth_error_estimate,
                                error)
    ) |> 
    group_by(Model) |> 
    mutate(
        gt_est_diff = if_else(
            is.nan(gt_est_diff),
            estimate_mean - estimate_mean[DAG == "Ground Truth"],
            gt_est_diff
        )
    ) |> 
    ungroup() |> 
    mutate(ogDAG = DAG) |> 
    mutate(DAG = if_else(estimated_ground_truth & DAG == "Ground Truth",
                         "Ground Truth (est)",
                         DAG) |> 
               factor(levels = c(
                   "Ground Truth",
                   "Ground Truth (est)",
                   "Expert DAG",
                   "Observable",
                   "Potentially Observable"
               )))
    
pal_cols <- palette.colors()
pal_cols <- c(pal_cols[1], "#777777", pal_cols[2], pal_cols[3])



# Would be nice to replicate this, but group 
# models into those identifiable and those not
est_to_plot |> 
    filter(DAG != "Observable") |> 
    mutate(DAG = if_else(DAG == "Potentially Observable", "Learned DAG", 
                         DAG) |> 
               factor(levels = c(
                   "Ground Truth",
                   "Ground Truth (est)",
                   "Expert DAG",
                   "Learned DAG"
               ))) |> 
    mutate(
        y_pos = as.numeric(factor(Model)) + 
            (as.numeric(factor(ogDAG)) - mean(as.numeric(factor(ogDAG)))) * dodge_amt
    ) |> 
    ggplot(aes(color = DAG, y = y_pos)) + 
    # geom_pointinterval(
    #     aes(x = gt_est_diff, 
    #         xmin = gt_est_diff - 2 * error, 
    #         xmax = gt_est_diff + 2 * error),
    #        alpha = 0.3, size = 1
    # ) +
    geom_pointinterval(
        aes(x = gt_est_diff,
            xmin = gt_est_diff - 1 *error,
            xmax = gt_est_diff + 1 *error),
           alpha = 0.8, size = 3
    ) +
    scale_y_continuous(
        breaks = seq_along(levels(factor(est_to_plot$Model))),
        labels = levels(factor(est_to_plot$Model)),
        name = "Model"
    ) +
    annotate("label", x = -0.0, y = c(2.2,3.2,5.2,6.2,8.2), label = "?",
             color = pal_cols[3], alpha = 0.4) +
    scale_colour_discrete(palette = pal_cols) +
    xlab("Deviation from ground truth estimate") +
    theme_minimal()
```

``` {r}
est_to_plot_grouped <- est_to_plot |> 
    filter(DAG %in% c("Potentially Observable", "Ground Truth", "Expert DAG")) |> 
    mutate(DAG = if_else(DAG == "Potentially Observable", "Learned DAG", 
                         DAG) |> 
               factor(levels = c(
                   "Ground Truth",
                   "Expert DAG",
                   "Learned DAG"
               ))) |> 
    group_by(Model) |> 
    mutate(model_group = if_else(
        is.nan(estimate_mean[DAG == "Expert DAG"]),
        "Non-identifiable", 
        "Identifiable"
        )) |> 
    ungroup() |> 
    group_by(model_group) |> 
    mutate(model_group_list = str_c(unique(model_group), " models\n(from observable data):\n",
        str_c(unique(sort(Model)) , collapse = ", "))) |> 
    ungroup() |> 
    mutate(
        y_pos = as.numeric(factor(model_group_list)) + 
            (as.numeric(factor(ogDAG)) - mean(as.numeric(factor(ogDAG)))) * dodge_amt
    )

est_to_plot_grouped |> 
    ggplot(aes(color = DAG, y = y_pos)) + 
    # geom_pointinterval(
    #     aes(x = gt_est_diff, 
    #         xmin = gt_est_diff - 2 * error, 
    #         xmax = gt_est_diff + 2 * error),
    #        alpha = 0.3, size = 1
    # ) +
    geom_pointinterval(
        aes(x = gt_est_diff,
            xmin = gt_est_diff - 1 *error,
            xmax = gt_est_diff + 1 *error),
           alpha = 0.6, size = 5
    ) +
    scale_y_continuous(
        breaks = seq_along(levels(factor(est_to_plot_grouped$model_group_list))),
        labels = levels(factor(est_to_plot_grouped$model_group_list)),
        name = "Model estimates by identifiability"
    ) +
    annotate("label", x = -0.0, y = c(2), label = "?",
             color = pal_cols[3], alpha = 0.4) +
    scale_colour_discrete(palette = palette.colors()) +
    xlab("Deviation from ground truth estimate") +
    theme_minimal() +
    theme(legend.position = "bottom")
```



``` {r}
plot_bnlearn_comparisons <- function(s_estimates) {
    p_orig <- s_estimates$DAG |> 
        dag_to_tidy_graph() %N>%
        mutate(name = str_remove(name, "_p$")) |> 
        g_plot_fine(weight_nodes = F)
    p_dwo <- s_estimates$fitted_BN$bn.dwo |>
        bn_to_dagitty() |> 
        dag_to_tidy_graph() %N>%
        mutate(name = str_remove(name, "_p$")) |> 
        g_plot_fine(weight_nodes = F)
    p_dwoop <- s_estimates$fitted_BN$bn.dwoop |> 
        bn_to_dagitty() |> 
        dag_to_tidy_graph() %N>%
        mutate(name = str_remove(name, "_p$")) |> 
        g_plot_fine(weight_nodes = F)
    return(p_orig + p_dwo + p_dwoop)
}

plot_bnlearn_coarse_comparison <- function(s_estimates) {
    p_orig <- s_estimates$DAG |> 
        dag_to_tidy_graph() %N>%
        mutate(name = str_remove(name, "_p$")) |>
        gcm_cluster_graph(code_node_group) |> 
        g_plot_coarse()
    p_opo <- s_estimates$fitted_BN$bn.opo |>
        bn_to_dagitty() |> 
        dag_to_tidy_graph() %N>%
        mutate(name = str_remove(name, "_p$")) |> 
        gcm_cluster_graph(code_node_group) |> 
        g_plot_coarse()
    return(p_orig  / p_opo )
    
}

plot_bnlearn_coarse_comparison(s_estimates = simulations_ng_cluster_models$M1) + 
    ggtitle("Model 1: Actual coarsened DAG (top) vs coarsened learned DAG from observable nodes")
plot_bnlearn_coarse_comparison(s_estimates = simulations_ng_cluster_models$M2) + 
    ggtitle("Model 2: Actual coarsened DAG (top) vs coarsened learned DAG from observable nodes")
plot_bnlearn_coarse_comparison(s_estimates = simulations_ng_cluster_models$M3) + 
    ggtitle("Model 3: Actual coarsened DAG (top) vs coarsened learned DAG from observable nodes")
plot_bnlearn_coarse_comparison(s_estimates = simulations_ng_cluster_models$M4) + 
    ggtitle("Model 4: Actual coarsened DAG (top) vs coarsened learned DAG from observable nodes")
plot_bnlearn_coarse_comparison(s_estimates = simulations_ng_cluster_models$M5) + 
    ggtitle("Model 5: Actual coarsened DAG (top) vs coarsened learned DAG from observable nodes")
plot_bnlearn_coarse_comparison(s_estimates = simulations_ng_cluster_models$M6) + 
    ggtitle("Model 6: Actual coarsened DAG (top) vs coarsened learned DAG from observable nodes")
plot_bnlearn_coarse_comparison(s_estimates = simulations_ng_cluster_models$M7) + 
    ggtitle("Model 7: Actual coarsened DAG (top) vs coarsened learned DAG from observable nodes")
plot_bnlearn_coarse_comparison(s_estimates = simulations_ng_cluster_models$M8) + 
    ggtitle("Model 8: Actual coarsened DAG (top) vs coarsened learned DAG from observable nodes")



```

## Distances between the simulations and the originals

``` {r}
lcdags <- 
    simulations_ng_cluster_models |> 
    set_names(nm = str_replace(names(simulations_ng_cluster_models), "M", "D")) |> 
    imap(
        \(est,m)
        est$fitted_BN$bn.opo |> 
            bn_to_dagitty() |> 
            dagitty_to_tidygraph()
    )

all.cluster.dags <- 
    c(tcdags_node_group, lcdags)

L1_between_model_and_simulation_metrics <- 
    compute_dist_matrix(all.cluster.dags, L1_dist, symmetric = TRUE)

plots_L1_similarities_full <- plotting_Ln_between_distances(
    L1_between_model_and_simulation_metrics$D, level_label = "L1")

```

### L1 consistency between models

Includes the models, $M_k$ and their corresponding DAG learned from data, $D_k$

``` {r}
plots_L1_similarities_full$p_combined
```

### L2 consistency between models

``` {r}
# L2 distance between two models


L2_between_model_metrics_full <- 
    compute_dist_matrix(
        all.cluster.dags, 
        L2_dist, 
        symmetric = FALSE)
```

``` {r}
plots_L2_similarities_full <- plotting_Ln_between_distances(
    L2_between_model_metrics_full$D, 
    level_label = "L2"
)

plots_L2_similarities_full$p_combined
```

### Averaged L1, L2 similarity between all models

``` {r}
library(ggtext)
plots_L12_mean_consistency_all <- 
    plotting_Ln_between_distances(
        (L1_between_model_and_simulation_metrics$D + L2_between_model_metrics_full$D) / 2,
        level_label = "L1-2"
    )


plots_L12_mean_consistency_all$p_combined
```

### L2 consistency between models??

Could base on the Jenson-Shannon distance between distributions - https://arxiv.org/pdf/2305.04357 - see definition 4. 