---
title: "Eliciting causal models -- Analysis"
execute:
  dev: svglite
format:
    html:
        code-fold: true
        warning: false
        message: false
        fig-format: svg
editor: 
  markdown: 
    wrap: 72
---

# Overview

### Render settings

```{r rendering-preferences}
# Set to true if L metrics
recalculate_distance_matrices <- FALSE
rerun_simulations <- FALSE
export_and_save_every_model <- FALSE
```

# TODO

## Considering adding to L1, L2 diss:

- Might already be there, but Independence / Dependence for L1
- Similarly, X->Y or not (perhaps as pa() style instead, so squiggle arrow) for L2.

## Check new definitions throughout:


## Change L2 cluster degradation to use optimal adjustment set as well

- Also, generally check the coarsening degradation aligns well enough
  with the consistency

## Vignette on causal degradation?

```{r}
#| include: false
# Old yaml - put back for de-anonymised repo
# execute:
#   dev: svglite
# format: 
#     html:
#         code-fold: true
#         warning: false
#         message: false
#         fig-format: svg

# Anonymised repo yaml
# format: 
#    pdf:
#        fig-format: png
```

```{r library-load}
#options(device = svglite::svglite)

library(svglite)
library(tidyverse)
library(dagitty)
library(tidygraph)
library(scales)
library(patchwork)
library(ggdist)
library(ggraph)
```

```{r helper-function}
source("gcm_helpers.R")
source("dagitty_tidygraph_translators.R")
```

# Notation

We will use curvy uppercase ($\mathcal{M,G,V,E}$) script for sets across
all the models, $\mathbf{BOLD}$ uppercase for sets, and $U,P,P,E,R$ case
standard for instances in the models. Lowercase like $v$ will be
reserved for imagining actual values (so a node taking on a particular
value).

The models are indexed as $\mathcal{M}=\{M_1,M_2,\dots,M_8 \}$ with
their corresponding graphical form as
$\mathcal{G}=\{G_1,G_2,\dots,G_8 \}$, and each graphical model $G_k$
comprises of a set of vertices $\mathbf{V}_{k}$ and edges
$\mathbf{E}_{k}$, so $G_k=\langle\mathbf{V}_{k}, \mathbf{E}_{k}\rangle$,
with the set of all nodes being $\mathcal{V}$ edges being $\mathcal{E}$.
$E_{ij}=V_i \rightarrow V_j$ indicates an edge between two nodes, and we
may use $X=V_i,Y=V_j,Z_k \in \mathbf{V}$ to streamline notation when
subscripts are getting messy. If two nodes are connected through a
mediated path, such as $V_i \rightarrow \dots \rightarrow V_j$ then we
write $\mathrm{De}_{ij}=V_i \rightsquigarrow V_j$ or
$V_i \in \mathrm{an}(V_j)$. To specify model context for ancestors,
descendants and mediated parts we subscript the model:
$V_i \in \mathrm{an}_{G_k}(V_j)$ or $V_i \rightsquigarrow _{G_k} V_j$.
Nodes $V_i$ are indexed over all models (so across $\mathcal{V}$), but
if we want to talk about $V_i$ as it stands in $G_k$ we would use
$V_{i:G_k}$, and likewise for an edge $E_{ij:G_k}$ residing in the
context of model $G_k$.

Alternate notation for parents: $\{\mathrm{pa}_X \}$

To describe some of the features of DAGs we will examine the set of
independence relationships in the models as well.

We will also be applying coarsening functions to create cluster DAGs. A
typical coarsening functions $\varphi$ is applied to the graph, and
within it all the nodes and edges. So
$\varphi (G_{k})=\varphi G_{k}=\langle\varphi\mathbf{V}_{k},\varphi\mathbf{E}_{k}\rangle=\langle[\mathbf{V}_k]_{\varphi},[\mathbf{E_k}]_{\varphi}\rangle$,
which will also refer to as a new cluster model
$G^C_{k}=\varphi G_k=\langle\mathbf{V}^{C}_{k},\mathbf{E}^{C}_{k}\rangle$
for slightly cleaner notation, and including the $C$ for cluster. The
coarsening function $\varphi$ induces an equivalence relation
$\sim _\varphi$ on $\mathbf{V}_{k}$ where $V_i \sim_\varphi V_j$ iff
$\varphi (V_i) = \varphi (V_j)$, i.e. both nodes map to the same
cluster. In this cluster model we would find $\varphi\mathbf{V}_{k}$.
The coarsened / cluster nodes would be
$[V_{i}]_{\varphi}=\{V_j \in \mathcal{V}:V_j \sim_{\varphi} V_i\}$, the
equivalence class of nodes $[V_i]_{\varphi}$, or simply $[V_i]$ if
$\varphi$ is clear from the context. These clusters can be relabled and
reindexed to $C_a = [V_i]_{\varphi}$ for legibility. The edges in
$\varphi G_{k}=G_{k}^{C}$ are then
$E^{\varphi}_{[i][j]}=[V_i]\rightarrow [V_j]=C_a \rightarrow C_b=E^{C}_{ab}$.
Choosing to use the $C$ or $\varphi$ syntax depends on the context - are
we mainly discussing the cluster model $G_k^C$ or the processes of
coarsening $\varphi G_{k}$? There is also the *causal degradation*
produced by $\varphi$, which we denote as $\mathrm{deg}_\varphi$
generally, $\mathrm{deg}_\varphi({G_k})$ as applied to a particular
model, and ${\mathcal{L}_i}\text{-}\mathrm{deg}_\varphi(G_k)$ to specify
a particular level of the causal hierarchy,
$\mathcal{L}_i:i\in \{1,2,3\}$.

A measure of *causal consistency* is also computed, but this is between
two models, rather than under the process of a clustering. This will be
denoted as ${\mathcal{L}_{i}}\text{-}\mathrm{d}_{xy}$ or
${\mathcal{L}_{i}}\text{-}\mathrm{d}$ if symmetric, so where
${\mathcal{L}_{i}}\text{-}\mathrm{d}_{xy}(G_a,G_b)={\mathcal{L}_{i}}\text{-}\operatorname{d}_{xy}(G_b,G_a)$.

Thought: There is also the graph surgery notation $G_{x\rightarrow y}$
for the adjusted graph for estimating $P(y|\operatorname{do}(x))$, might
be nicer even as $G_{y \leftarrow x}$. This works nicer having lowercase
$x,y$ as the nodes, rather than the $V_i$ notation...but it can still be
$G_{X \rightarrow Y}$ or $G_{v_{i}\rightarrow v_{j}}$.

Notation for adjustment sets? Things like that?

The paper above with the snip from uses a hard (fix a value)
intervention and a soft (fix a set of values) / stochastic intervention
with $\mathcal{I}_X$.


# Theory: Causal consistency between models

Good paper on projections and preservation of causal quantities under
them:
https://www-sciencedirect-com.ezproxy.lib.uts.edu.au/science/article/pii/S0888613X24000951

## Metrics, semimetrics, pseudometrics, dissimilarities

Supposing we have some function $\operatorname{d}(x,y) \rightarrow R$

- **Dissimilarities** are non-negative, symmetric, and zero on the
  diagonal. This is good enough for things like `hclust`.
- If we add the triangle inequality, we have a **pseudometric**.
- If we add the iden...we have a **semimetric** (although this language depends on the field.)

## Levels

+---------------------------------+---------------------------------+
| Causal level                    | Example Statements; $\gamma$    |
+=================================+=================================+
| $\mathcal{L}_0$ : Description / | List of variables (the          |
|    bounding                     | vocabulary)                     |
+---------------------------------+---------------------------------+
| $\mathcal{L}_1$ : Association / | Implied conditional             |
| Seeing                          | independencies;                 |
|                                 |                                 |
|                                 | d-separation                    |
+---------------------------------+---------------------------------+
| $\mathcal{L}_2$ : Intervention  | Adjustment sets; graph          |
| / Doing                         | operations under interventions  |
+---------------------------------+---------------------------------+
| $\mathcal{L}_3$ :               | Joint distribution under        |
| Counterfactual / Imagining      | manipulation of variables       |
+---------------------------------+---------------------------------+

## Causal consistency

There is a notion of causal consistency based on Pearl's causal
hierarchy. Two models, $\mathcal{M}_a$ and $\mathcal{M}_b$ with a shared
vocabulary $\mathcal{S}=\mathbf{V}_a\cap \mathbf{V}_b$, are said to be
$\mathcal{L}_i\text{-consistent}$ if they agree on all statements for
that given level (Schooltink?). The challenge then is when a range of
models share a different number of variables. Two models can be
vacuously consistent by $\mathcal{S}=\varnothing$. As such we can define
an $\mathcal{L}_0$ level with an abuse of the notation to be simply the
collection of variables.

As consistency is described as agreement between sets of statements we
can use the Jaccard index between two sets of an agreed collection of
$\mathcal{L}_i$ statemnts to measure the consistency, and the
Jaccard distance to create a distance function between the two models.
These are built from the principles articulated in Schooltink et. al,
were we are looking at some measure of the preservation of all the
$\mathcal{L}_i$ algebraic statements about the GCM. These measures are
not exhaustive, but rather choose the implied conditional independence
relationships for $\mathcal{L}_1$ and the sets of adjustment sets for
$\mathcal{L}_2$.

Let $\gamma_{i}$ be a claim at $\mathcal{L}_i$ level, and
$\Gamma_{i}(G)=\{\gamma_i | \gamma_i \text{ is true within }G\}$, so the
set of all true claims (of a given type) for the model $G$. Those
statements may be about any nodes in $G$, so if we want to talk only
about the statements belonging to the shared set of nodes $\mathcal{S}$
then we write $\Gamma_{i}(G){[\mathcal{S}]}=\hat{\Gamma}_{i}(G)$.
Depending on the level of causality the statement could be more or less
strict about what is included. For instance we could look at
d-separation on $\mathcal{L}_1$ statements of the form
$\gamma=\langle x \perp_{d} y \ | \ z \rangle _{G}$ . There is a hard
form of inclusion in $\hat{\Gamma_{i}}$ where
$x,y,\mathbf{z} \in \mathcal{S}$ and a softer form of inclusion where
$x,y \in \mathcal{S}; \exists \mathbf{z} \in \mathbf{V}_G$. The former
asks if $x,y \in \mathcal{S}$ are d-separated by some set of variables
also in the shared vocabulary $\mathcal{S}$, where as the latter asks if
$x,y \in \mathcal{S}$ are able to be d-separated by some set of
variables in $G$.

### Causal dissimilarity

Then we define:

$$\operatorname{d}_{\mathcal{L}_i}(G,H)=1 - J_{\mathcal{L}_i}; J_{\mathcal{L}_i}=\frac{|\hat{\Gamma}_i(G) \ \cap \ \hat{\Gamma}_{i}(H)|}{|\hat{\Gamma}_i(G) \ \cup \ \hat{\Gamma}_{i}(H)|}$$

Or, using the symmetric difference operator, and using restriction
notation, so $\Gamma\!\upharpoonright\!S$ are the statements restricted
to a set S.

$$\operatorname{d}_{\mathcal{L}_i}(G,H)=\frac{|\ (\Gamma_{i}(G) \ \Delta \ {\Gamma}_{i}(H))\!\upharpoonright\!\mathcal{S}_{GH}\ | }{|\ ({\Gamma}_i(G) \ \cup \ {\Gamma}_{i}(H))\!\upharpoonright\!\mathcal{S}_{GH}\ |}$$

Or, we could look at all the statements, without the $\mathcal{S}$, and
let $\mathcal{L}_0$ carry that load:

$$\operatorname{d}_{\mathcal{L}_i}(G,H)=\frac{|{\Gamma}_i(G) \ \Delta \ {\Gamma}_{i}(H)|}{|{\Gamma}_i(G) \ \cup \ {\Gamma}_{i}(H)|}$$

It might be easier to use the latent projection notation. A latent
projection of a graph $G$ onto a set of nodes $\mathcal{S}$ is
$\pi_{\mathcal{S}}(G)$ and preserves $\mathcal{L}_{1,2}$ statements. L1
is Verma and Pearls theorem, for disjoint X, Y, Z, X *\|* Y \| Z is
d-separated in G iff m-separated in $\pi_{\mathcal{S}}(G)$. So
$\Gamma_{1}(G)\upharpoonright \mathcal{S}$, the statements restricted to
S, is $\pi_{\mathcal{S}}(G)$ exactly (get ref) - well so some claim. However, that has not proven to be the case with the implementation here. Using the restrction to a set of statements instead of the projection method. 

Key to the implementation is the choice of what statement to include in
$\Gamma_i$, and how then to restrict this to the possible shared
statements between the models $\hat{\Gamma}_{i}$.

### Causal degradation

A similar use of $\Gamma_{i}$ can be used to examine the possible loss
of $\mathcal{L}_i$ claims under some coarsening of graph through the
clustering of nodes. One way this can happen when the coarsening divides
the adjustment sets of some pair of variables into different clusters.
Informally this degradation measure counts the proportion of
$\mathcal{L}_i$ claims in the CDAG that have no potential surrogate
claim in the fine grained DAG. More formally, we define:

$$\varphi\text{-}\operatorname{deg}_{\mathcal{L}_i}(G)=\frac{|\ \{ [\gamma]_{\varphi} \in \Gamma_{i}(\varphi G): \nexists \ \eta \in \Gamma_i(G) \text{ s.t. } \varphi(\eta) = [\gamma]_{\varphi} \}\ |}{|\ \varphi\Gamma_{i}(G) \cup  \Gamma_{i}(\varphi G)  |}$$



# Defining dissimilarity functions

```{r defining-distance-functions, echo = T}
L0_d <- L0_dissimilarity_vocab

L1_d <- partial(
    L1_dissimilarity_ci, 
    max_cond = 5)
L2_d <- partial(
    L2_dissimilarity_adj_set,
    adj_type = "all")

```

# Vignettes

To articulate some of the theory and the new metrics.

```{r vignette-dags}
dagc1 <- dagitty::dagitty(
    "dag{
    X -> Y;
    X <- Z -> Y
    }"
)
dagc1w <- dagitty::dagitty(
    "dag{
    X -> Y;
    X <- Z -> Y;
    W <- Z;
    X <- W
    }"
)

dagc2 <- dagitty::dagitty(
    "dag{
    X <- Y;
    X <- Z -> Y
    }"
)

dagf1a <- dagitty::dagitty(
    "dag{
    X1 -> X2 -> Y;
    X1 <- Z1 -> Y;
    X2 <- Z2 -> Y;
    Z1 -> Z2
    }"
)
dagf1b <- dagitty::dagitty(
    "dag{
    X1 <- Z -> Y;
    X1 <- X2 -> Y;
    X1 -> Y
    }"
)

vnt.layout <- tribble(
    ~name, ~x, ~y,
    "W", 0,1,
    "X", 0, 0,
    "X1", 0, 0,
    "X2", 0.6, 0.2,
    "Y", 2, 0,
    "Z", 1, 1, 
    "Z1", 0.7, 1, 
    "Z2", 1.3, 1 
)

first_letter_pal <- list(
    "W" = palette.colors()[4],
    "X" = palette.colors()[1],
    "Y" = palette.colors()[2],
    "Z" = palette.colors()[3]
    )

first_letter_coarsen <- function(V){
    str_extract(V, "^.{1}")
}

vnt.tdags <- map(
    list(dagc1, dagc2, dagc1w, dagf1a, dagf1b), 
    \(g) 
    dagitty_to_tidygraph(g) |> 
        activate(nodes) |> 
        inner_join(vnt.layout, by = "name") |> 
        mutate(group = first_letter_coarsen(name))) |> 
    set_names(c("Cluster1", "Cluster2", "Cluster1W", "Fine1a","Fine1b"))
    
# gcm_cluster_L2_degradation(vnt.tdags$Fine1a, first_letter_coarsen)
# gcm_cluster_L2_degradation(vnt.tdags$Fine1b, first_letter_coarsen) # want to simultaneously control for X2, but not for X1

plot_vnt_dag <- function(g) {
    ggraph(g, layout = "manual", x = x, y = y)+
        geom_edge_fan(
            arrow = arrow(length = unit(4, 'mm'), type = "closed"),
            start_cap = circle(6, 'mm'),
            end_cap = circle(6, 'mm')) +
        geom_node_point(aes(colour = group), size = 6) +
        geom_node_text(aes(label = name), colour = "white", alpha = 0.9) +
        scale_colour_manual(
            values = first_letter_pal,
            guide = "none") +
        xlim(c(-0.1,2.1)) +
        ylim(c(-0.1,1.1)) +
        theme_graph()
}

vntPlotC1 <- plot_vnt_dag(vnt.tdags$Cluster1)
vntPlotC2 <- plot_vnt_dag(vnt.tdags$Cluster2)
vntPlotC1W <- plot_vnt_dag(vnt.tdags$Cluster1W)

vntPlotF1A <- 
    plot_vnt_dag(vnt.tdags$Fine1a) + 
    annotate(geom = "rect", 
             xmin = -0.1, xmax = 0.7, ymin = -0.1, ymax = 0.3, 
             fill = "grey", linewidth = 1, 
             colour = "black", linetype = 5, alpha = 0.2) +
    annotate(geom = "rect", 
             xmin = 0.6, xmax = 1.4, ymin = 0.9, ymax = 1.1, 
             fill = "grey", linewidth = 1, colour = palette.colors()[[3]], linetype = 5, alpha = 0.2)

vntPlotF1B <- plot_vnt_dag(vnt.tdags$Fine1b) +
    annotate(
        geom = "rect", 
        xmin = -0.1, xmax = 0.7, ymin = -0.1, ymax = 0.3, 
        fill = "grey", linewidth = 1, colour = "black", linetype = 5,
        alpha = 0.2) 

# Causal consistency between C1, C2
d.L0.C1_C2 <-    L0_d(
    vnt.tdags$Cluster1, 
    vnt.tdags$Cluster2)
d.L1.C1_C2 <-    L1_d(
    vnt.tdags$Cluster1, 
    vnt.tdags$Cluster2)
d.L2.C1_C2 <- L2_d(
    vnt.tdags$Cluster1, 
    vnt.tdags$Cluster2)

# Causal consistency between C1, C1W
d.L0.C1_C1W <-    L0_d(
    vnt.tdags$Cluster1, 
    vnt.tdags$Cluster1W)
d.L1.C1_C1W <-     L1_d(
    vnt.tdags$Cluster1, 
    vnt.tdags$Cluster1W)
d.L2.C1_C1W <-  L2_d(
    vnt.tdags$Cluster1, 
    vnt.tdags$Cluster1W)

# Causal consistency between C2, C1W for completeness - probably don't need to include
d.L0.C2_C1W <-    L0_d(
    vnt.tdags$Cluster2, 
    vnt.tdags$Cluster1W)
d.L1.C2_C1W <-    L1_d(
    vnt.tdags$Cluster2, 
    vnt.tdags$Cluster1W)
d.L2.C2_C1W <- L2_d(
    vnt.tdags$Cluster2, 
    vnt.tdags$Cluster1W)

# Causal degradation under coarsening map
deg.L1.FA <- gcm_cluster_L1_degradation(
    g = vnt.tdags$Fine1a, .f = first_letter_coarsen)$degradation
deg.L1.FB <- gcm_cluster_L1_degradation(vnt.tdags$Fine1b, first_letter_coarsen)$degradation

deg.L2.FA <- gcm_cluster_L2_degradation(vnt.tdags$Fine1a, first_letter_coarsen)$degradation
deg.L2.FB <- gcm_cluster_L2_degradation(vnt.tdags$Fine1b, first_letter_coarsen)$degradation
```

## $\mathcal{L}$-dissimilarities

Given a base DAG of
$\mathcal{G}=\{X \rightarrow Y\} \cup \{X \leftarrow Z \rightarrow Y\}$
we examine the ${^{\mathcal{L}_i}\operatorname{d}}$ distance for two
similar graphs, one where we switch the direction of an arrow,
$\mathcal{G}^{\overleftarrow{xy}}$ and one where we add a node,
$\mathcal{G}^{+w}$.

```{r figure-Ldist-vignette, fig.height=3}
library(ggtext)

dLstringC1_C1W <- str_c(
    "Node W added",
    "<br>L<sub>0</sub>-d = ", 
    round(d.L0.C1_C1W,2), 
    "<br>L<sub>1</sub>-d = ", 
    round(d.L1.C1_C1W, 2),
    "<br>L<sub>2</sub>-d = ", 
    round(d.L2.C1_C1W, 2),
    "<br>L-d = ", 
    round((d.L0.C1_C1W + 
               d.L1.C1_C1W + 
               d.L2.C1_C1W) /3, 2))

dLstringC1_C2 <- str_c(
    "Edge reversed",
    "<br>L<sub>0</sub>-d = ", 
    round(d.L0.C1_C2,2), 
    "<br>L<sub>1</sub>-d = ", 
    round(d.L1.C1_C2, 2),
    "<br>L<sub>2</sub>-d = ", 
    round(d.L2.C1_C2, 2),
    "<br>L-d = ", 
    round((d.L0.C1_C2 + 
               d.L1.C1_C2 + 
               d.L2.C1_C2) /3, 2)
)

p.vignette.diss <- (
    vntPlotC1W +
                        geom_richtext(
        data = data.frame(x = 1, y = 0.3),
        aes(x = x, y = y, label = dLstringC1_C1W),
        fill = NA, label.color = "lightgrey",   # no box/border around the text
        inherit.aes = FALSE)
    ) +
    (vntPlotC1 + 
        annotate("text", x = 1, y = 0.3, label = "Initial DAG")) + 
    (vntPlotC2 +                                                                                      geom_richtext(
        data = data.frame(x = 1, y = 0.3),
        aes(x = x, y = y, label = dLstringC1_C2),
        fill = NA, label.color = "lightgrey",   # no box/border around the text
        inherit.aes = FALSE))

ggsave("figures/fig--vignette-causal-dissimilarity.png", 
       height = 4, width = 12,
       plot = p.vignette.diss)

p.vignette.diss

```

We can see that the mean distance is similar between $\mathcal{G}$ and
the two variations. However the edge reversal is penalising on the
$\mathcal{L}_2$ distance, whilst the addition of a (causally) irrelevant
node only penalises on the $\mathcal{L}_0$ distance. This highlights
that $\operatorname{d}^{\mathcal{L}_1}$.

> It is important to note that $^\mathcal{L}_{1,2}\operatorname{d}$ is
> sensitive to the proportion of nodes shared between the graphs. Two
> large graphs that only overlap on a small number of nodes and still be
> largely $\mathcal{L}_{1,2}$-consistent if the shared nodes have
> similar causal implications.

## $\mathcal{L}$-degradation

Here we examine the same baseline DAG, but instead look at two
'finer-grained' DAGs that have more detail in the causal structure, but
we can recover the original DAG by clustering some nodes and coarsening
the finer grained DAGs. We consider two finer grained DAGs,
$\mathcal{G}^{[x,z]}=\{X_1 \rightarrow X_2 \rightarrow Y\} \cup \{X_1 \leftarrow Z_1 \rightarrow Y\} \cup \{X_2 \leftarrow Z_2 \rightarrow Y\}$
that clusters on both $X$ and $Z$, and
$\mathcal{G}^{[x]}=\{X_2 \rightarrow X_1 \rightarrow Y\} \cup \{X_2 \rightarrow Y\} \cup \{X_1 \leftarrow Z \rightarrow Y\}$
that only clusters on $X$ but has a more problematic causal structure
for this coarsening. The coarsening map, $\phi$ is simply grouping the
variables by their first letter.

```{r figure-L-degradation-vignette, fig.height=3}
p.vignette.deg <- 
    (vntPlotF1A ) + 
    (vntPlotC1 + 
        annotate("text", x = 1, y = 0.3, label = "Cluster DAG")) + 
    (vntPlotF1B)

ggsave("figures/fig--vignette-causal-degradation.png", 
       height = 4, width = 12,
       plot = p.vignette.deg)

p.vignette.deg
```

Under this coarsening we see:

- $\operatorname{deg}_{\mathcal{L_1}}^{\phi}(\mathcal{G}_{[x,z]})$ =
  `r deg.L1.FA`
- $\operatorname{deg}_{\mathcal{L_2}}^{\phi}(\mathcal{G}_{[x,z]})$ =
  `r deg.L2.FA`
- $\operatorname{deg}_{\mathcal{L_1}}^{\phi}(\mathcal{G}_{[x]})$ =
  `r deg.L1.FB`
- $\operatorname{deg}_{\mathcal{L_2}}^{\phi}(\mathcal{G}_{[x]})$ =
  `r deg.L2.FB`

Here we see that the only causal degradation is when we coarsen
$\mathcal{G}^{[x]}$. We see the $\mathcal{L}_2$ degradation because the
adjustment sets for $P(Y |\operatorname{do}(Z))$ disagree on what to do
with $X$. In the fine graph $X_1$ should be excluded from the adjustment
set, wheras $X_2$ should be included as the open path
$Z \rightarrow X_1 \leftarrow X_2 \rightarrow Y$ is open but non-causal.
This disagreement cannot be reconciled when we form the cluster
$[X]=\{X_1, X_2\}$ in the cluster DAG.
We see the $\mathcal{L}_1$ degradation because the cluster DAG is fully connected, but the fine DAG has $X_2 \perp Z$, which is lost moving to the cluster DAG.

> Thought: Could the causal degradation work equally well for projections, and embeddings?

### Relationship with causal consistency of 


## Diabolic trio breaks the triangle inequality

The following three DAGs have been constructed to push the triangle
inequality: $d(x,z) \le d(x,y)+d(y,z)$. They are chosen so that model
$G_Y$ is very similar in its overlapping nodes to $G_X$ and $G_Z$, but
that $G_X$ and $G_Z$ strongly disagree where they overlap.

```{r}
diabolic.X <- tbl_graph(
    edges = tribble(
        ~from, ~to,
        # Has strong connections between A, B, C
        "A", "B",
        "A", "C",
        "B", "C",
        "A", "Lx",
        # Identical structure in Y
        "Lx", "Mx",
        "Lx", "Kx",
        "Mx", "Kx",
    )
)

diabolic.Y <- tbl_graph(
    edges = tribble(
        ~from, ~to,
        # Identical structure 
        "Lx", "Mx",
        "Lx", "Kx",
        "Mx", "Kx",
        "Lz", "Mz",
        "Lz", "Kz",
        "Mz", "Kz",
        # joining them together
        "Y", "Kz",
        "Y", "Kx"
    )
)

diabolic.Z <- tbl_graph(
    edges = tribble(
        ~from, ~to,
        # Has strong connections between A, B, C
        "A", "Mz",
        "B", "Lz",
        "C", "Kz",
        # Identical structure in Y
        "Lz", "Mz",
        "Lz", "Kz",
        "Mz", "Kz"
    )
)
```

```{r}
gcm_qplot(diabolic.X)
gcm_qplot(diabolic.Y)
gcm_qplot(diabolic.Z)

```

```{r}
diabolic.dissimilarities <- 
    tibble(pair = c("X, Z", "X, Y", "Y, Z"),
       L0 = c(
           L0_d(diabolic.X, diabolic.Z),
           L0_d(diabolic.X, diabolic.Y),
           L0_d(diabolic.X, diabolic.Y)
           ),
       L1 = c(
           L1_d(diabolic.X, diabolic.Z),
           L1_d(diabolic.X, diabolic.Y),
           L1_d(diabolic.X, diabolic.Y)
           ),
       L2 = c(
           L2_d(diabolic.X, diabolic.Z),
           L2_d(diabolic.X, diabolic.Y),
           L2_d(diabolic.X, diabolic.Y)
           )
       ) 

```

For these difficult models, we first look at the distance metric, which checks all $\mathcal{L}_i$ statements, regardless of a common vocabulary, and finds the Jaccard distance of that:

The triangle equality holds for this set, but it also describes X,Y and Y,Z as nearly as far apart on $\mathcal{L}_1$ and $\mathcal{L}_2$ as possible, despite containing no contradictions of statements within either of these shared vocabularies. These seems like an unhealthy way to examine causal consistency between models. The dissimilarity measure completely violates the triangle equality however, but returns seemingly more intuitive results: 

``` {r}
diabolic.dissimilarities |> 
    knitr::kable(digits = 2)
```

Note here that $(\mathcal{L}_0 + \mathcal{L}_1 + \mathcal{L}_2)/3$ does not satisfy the triangle inequality in this case. This, however, is not required for clustering anyway, and it is better that the measures make sense. 

# Data import

```{r loading-models}
# imports all causal models as DAGitty objects
# They are names dag_s1, ..., dag_s8 corresponding to the eight sessions
# A table of session data is also included
source("load-GCMs.R")
```

```{r colour-settings}
# Some colours
# \definecolor{myochre}{RGB}{204,119,34}
# \definecolor{mygum}{RGB}{87,180,66}
myochre <- rgb(204, 119, 34, maxColorValue = 255)
mygum <- rgb(87, 180, 66, maxColorValue = 255)
node_type_colours <- c(
  "Assessment data"    = "#000000",
  "Learning data"      = "#999999",
  "Help"               = "#CC79A7",
  "Student"            = "#009E73",
  "Teacher"            = "#E69F00",
  "Learning design"    = "#56B4E9"
)

wrap_width <- 10

hue1 <- palette.colors(3, palette = "Okabe-Ito")[[2]]
hue2 <- palette.colors(3, palette = "Okabe-Ito")[[3]]

```

Modelling sessions were originally run on Loopy, then transcribed into
DAGitty (online) to generate the following `dot` language syntax for
importing into `R`. A better way of doing this is under development.

Data imported as `DAGitty` objects, but need to be in `tidygraph` format
for many calculations and visualisations.

Creating a list of all the DAGs in tidy format:

```{r}
# making list of dagitty dags (used later)
ddags <- list(
    "M1" = dag_m1, 
    "M2" = dag_m2,
    "M3" = dag_m3,
    "M4" = dag_m4,
    "M5" = dag_m5,
    "M6" = dag_m6,
    "M7" = dag_m7,
    "M8" = dag_m8
    )

tdags <- list(
    "M1" = dagitty_to_tidygraph(dag_m1), 
    "M2" = dagitty_to_tidygraph(dag_m2), 
    "M3" = dagitty_to_tidygraph(dag_m3), 
    "M4" = dagitty_to_tidygraph(dag_m4),
    "M5" = dagitty_to_tidygraph(dag_m5), 
    "M6" = dagitty_to_tidygraph(dag_m6) |> 
    filter( # this removes feedback link to next learning cycle
        !str_detect(name, "Post")), 
    "M7" = dagitty_to_tidygraph(dag_m7), 
    "M8" = dagitty_to_tidygraph(dag_m8))

ddags.raw <- ddags

ddags <- 
    imap(tdags, \(g, n)
         tidygraph_to_dagitty(g)) |> 
    map(\(d) { 
        dagitty::outcomes(d) <- "Grade"
        d}) |> 
    imap(\(d, n) {
        if (n != "M7") {
            dagitty::exposures(d) <- 
                "S.Kn"
        } else {
            dagitty::exposures(d) <- 
                c("S.Kn.Coms", "S.Kn.Interp", "S.Kn.Tech")
        }
        d})

quick_list_of_all_nodes <- 
    map(tdags, \(g) pull(g, name)) |> 
    unlist() |> 
    unique() |> 
    sort()

rm(dag_m1, dag_m2, dag_m3, dag_m4, dag_m5, dag_m6, dag_m7, dag_m8)

```

# Coding and Clustering

Different people used different language to describe their understanding
of the system in question. In order to make comparisons we must work out
when they are talking about the same thing (in graph terms, the same
node) or when they are talking about different things. To complicate
things there is also the challenge of hierarchical modelling choices -
sometimes people would describe one thing in more detail than others. In
graph terms this would mean one model might have "student subject
knowledge" (see model 8) whereas another might break this into
components (see model 7).

To make meaningful comparisons we must:

(1) Form a common dictionary of nodes across all models, identifying
    those that should be treated as the the same across different
    models.
(2) Construct a map from the finer-grained nodes to higher level nodes.
    This will facilitate comparisons between models that discuss

## Node Classifiers

Coding the nodes into rough groups. There are two levels:

- `node_group` is the first aggregation into slightly coarser groups of
  nodes.
- `node_type` is an even more coarse grouping into a handful of types.

Also some nicer names of the fields: - `node_nice_name` is a longer more
explanatory name for plotting and putting in tables. -
`node_group_nice_name` a longer more explanatory name for node groups,
as above.

```{r nicer-names}
name_node_nice_name <- tribble(
    ~name, ~nice_name,
    "CD", "Any system data",
    "CD.Att", "Attendance",
    "CD.LMS", "LMS activity",
    "CD.LMS.code", "Code artifacts",
    "CD.LMS.write", "Writing artifacts", 
    "CD.Str", "Course Data",
    "Grade", "Assessment grade",
    "Grade.Coms", "Mark - Communication",
    "Grade.Exam", "Mark - Exam",
    "Grade.Interp", "Mark - Interpretation",
    "Grade.Proc", "Mark - Learning process",
    "Grade.Quiz", "Mark - Quiz",
    "Grade.Ref", "Mark - Learning reflection",
    "Grade.Tech", "Mark - Technical",
    "H", "Help (general)",
    "H.AI", "Help from AI (any)",
    "H.AI.Agent", "Help from AI (agent)",
    "H.AI.evil", "Misuse of AI",
    "H.AI.evil.code", "Misuse of AI (code)",
    "H.AI.evil.write", "Misuse from AI (write)",
    "H.AI.Fb", "Help from AI (feedback)",
    "H.AI.good", "Help from AI (to learn)",
    "H.AI.Res", "Help from AI (research)",
    "H.AI.Study", "Help from AI (revision)",
    "H.Ment", "Help from mentor",
    "H.other.evil", "Cheating - non-AI",
    "H.Peer", "Help from peers",
    "H.Tut", "Help from teacher / tutor",
    "H.Tut.Fb", "Feedback from teacher / tutor",
    "LD", "Learning design",
    "LD.Other", "Other learning design",
    "LD.AIguide", "Guidance on AI use",
    "LD.Task", "Task design",
    "LD.Task.Code", "Coding task design",
    "LD.Task.Grp", "Groupwork task design",
    "LD.Task.Other", "Other task design",
    "LD.Task.Prac", "Practical task design",
    "LD.Task.Write", "Writing task design",
    "S.Act", "Student decisions / actions",
    "S.Act.AIstrat", "Student AI strategy",
    "S.Act.Fb", "Student acts on feedback",
    "S.Act.Fb.Post", "Student acts on feedback between subjects",
    "S.Act.Frm", "Student does formative tasks",
    "S.Act.Perf", "Student performance on task",
    "S.Act.Time", "Student commits time",
    "S.Aff.Integ", "Student acting with integrity",
    "S.Aff.Joy", "Student enjoyment",
    "S.Aff.MindSet", "Student mindset",
    "S.Aff.Safe", "Student feels safe",
    "S.Aff.WrkLd", "Student workload pressure",
    "S.Cp.WrkLd", "Student workload pressure",
    "S.Aff.Hsk", "Student help seeker",
    "S.Att.Eq", "Student equity group",
    "S.Cp.Hsk", "Student help seeker",
    "S.Att.Nro", "Student neuro-divergent",
    "S.Att.SocCap", "Student social capital",
    "S.Kn.Aca", "Student academic skill",
    "S.Cp.Aca", "Student academic skill",
    "S.Cp.Aca.Post", "Student academic skill: Post grade",
    "S.Cp.EvJdg", "Student evaluative judgement",
    "S.Cp.Fb", "Student feedback literacy",
    "S.Cp.Grp", "Student groupwork capability",
    "S.Cp.Ind", "Student individual work capability",
    "S.Cp.Lrn", "Student metacognative learning capabilities",
    "S.Cp.SRL", "Student SRL",
    "S.Kn", "Student subject knowledge",
    "S.Kn.Post", "Student subject knowledge: Post grade",
    "S.Kn.Coms", "Student communications knowledge",
    "S.Kn.Interp", "Student interpretation knowledge",
    "S.Kn.Prior", "Student prior knowledge",
    "S.Kn.Tech", "Student technical knowledge",
    "T.Act.UseAI", "Teacher use of AI",
    "T.Aff.Integ", "Teacher acts with integrity",
    "T.Aff.WrkLd", "Teacher workload pressure",
    "T.Cp.WrkLd", "Teacher workload pressure",
    "T.Cp.Inst", "Teacher instruction quality",
    "T.Cp.KnStu", "Teacher knows students",
    "T.Cp.Mark", "Teacher reliably marks",
    "T.Kn.AI", "Teacher AI literacy",
    "T.Cp.AI", "Teacher AI expertise",
    "T.Kn.Ass", "Teacher assessment literacy",
    "T.Kn.Sub", "Teacher subject knowledge",
    "T.Cp.Sub", "Teacher subject expertise"
)

node_group_nice_name <- tribble(
    ~node_group, ~nice_name,
    "Assessments", "Preliminary assessments", 
    "CD.Att", "Student attendance",
    "CD.LMS", "LMS activity",
    "CD.Str", "Course Data",
    "Grade", "Assessment grade",
    "H.AI.evil", "Misuse of AI",
    "H.AI.good", "Help from AI",
    "H.Cheat", "Cheating - non-AI",
    "H.Peer", "Help from peers",
    "H.Gen", "Help (General)",
    "H.Tut", "Help from teacher / tutor",
    "LD.Gen", "Learning design",
    "LD.Task", "Task design",
    "S.Act", "Student decisions / actions",
    "S.Att", "Student attributes",
    "S.Cp", "Student metacognitive skill",
    "S.EM", "Student emotion / motivation",
    "S.Kn", "Student knowledge",
    "S.Kn.Prior", "Student prior knowledge",
    # "T.Act", "Teacher use of AI",
    "T.Cp", "Teacher capabilities", 
    "T.EM", "Teacher emotion / motivation",
    "T.Kn", "Teacher knowledge"
)
```

This process involved iterating between:

1.  Examining the transcript, the node description (label in the
    diagram), and forming an initial classification of the node. A
    natural place for a node-group to form was where in one model a node
    was broken into sub-components, where in others it was kept whole.
    `Student knowledge` was a key example of this, where several models
    split this into finer grained nodes, where others did not.
2.  After this the graph coarsening according to the map
    $\varphi(V_i)=\mathrm{nodegroup}(V_i)$, calculating the
    $\mathcal{L}_1$ and $\mathcal{L}_2$ cost of clustering the nodes
    into their node groups. From this two operations were performed on
    the cluster graph: (a) identifying which node clusters had
    contributed most to the degradation of the causal consistency,
    and (b) parts of the cluster graph that could potentially be
    coarsened further.
3.  Nodes and node groups were then then re-evaluated, based on which
    clusters of nodes caused the most degradation of causal consistency
    when coarsening the graphs. Often this amounted to reasonable
    re-interpretations of the transcript and the graph, such as
    distinguishing between aspects of student knowledge that were
    present prior to the period of study, and those aspects of student
    knowledge that were changing during the study. Step 2 was then
    repeated and the re-evaluation continued.

```{r node-group}
code_node_group <- function(v) {
    case_when(
        # Course data - Str, LMS --------------
        v == "CD" ~ "CD.Str", # was blending LMS and Str, but original graph was more about structure
        str_detect(v, "CD\\.LMS") ~ "CD.LMS",
        v == "CD.Att" ~ "CD.Att", # Attendance data
        v == "CD.Str" ~ "CD.Str",
        # Grade data --------------------------
        v == "Grade" ~ "Grade",
        str_detect(v, "Grade\\.") ~ "Grade", # Assessments
        # Kinds of help / assistance -----------------------
        # - Help from teacher or tutor
        # - - H.Gen(eral) for the broad H category
        v == "H" ~ "H.Gen", # This should actually be the node_type...somehow
        v == "H.Tut" ~ "H.Tut",
        v == "H.Ment" ~ "H.Tut",
        v == "H.Tut.Fb" ~ "H.Tut",
        # - Help from peers - collapsing with Tut for "help from people" - change from first coding
        v == "H.Peer" ~ "H.Tut",
        # - Help from AI (good or bad usage)
        str_detect(v, "H\\.AI\\.evil") ~ "H.AI.evil",
        str_detect(v, "H\\.AI") ~ "H.AI.good",
        # - Other nefarious help (i.e. cheating)
        v == "H.other.evil" ~ "H.Cheat",
        # Learning Design --------------------
        str_detect(v, "^LD\\.Task") ~ "LD.Task",
        str_detect(v, "^LD") ~ "LD.Gen",
        # Student nodes ---------------------------------
        # - Student activity / decisions 
        # - - Action post grade a seperate node
        v == "S.Act.Fb.Post" ~ "S.Act.Post",
        v == "S.Cp.Aca.Post" ~ "S.Cp.Post",
        str_detect(v, "^S\\.Act") ~ "S.Act",
        # - Student attributes (generally unchangable) 
        str_detect(v, "^S\\.Att") ~ "S.Att",
        # - Student metacognitive capability (metacognition) / capacity / workload
        str_detect(v, "^S\\.Cp") ~ "S.Cp",
        # - Student emotion / motivation
        str_detect(v, "^S\\.Aff") ~ "S.EM",
        # - Student knowledge (cognition)
        v == "S.Kn.Prior" ~ "S.Kn.Prior", 
        v == "S.Kn.Post" ~ "S.Kn.Post", 
        str_detect(v, "^S\\.Kn") ~ "S.Kn",
        # Teacher nodes ---------------------------------
        # - Teacher knowledge / meta-cognition: What they bring to the table
        str_detect(v, "^T\\.Kn") ~ "T.Kn",
        # - Teacher capabilities - usually executed during learning cycle
        str_detect(v, "^T\\.Cp") ~ "T.Cp",
        # - Teacher emotion / motivation
        str_detect(v, "^T\\.Aff") ~ "T.EM",
        TRUE ~ v
        
    )
}
```

```{r node-type}
code_node_type <- function(n) {
    fl <- str_extract(n, "^.{1}")
    case_when(
        fl == "S" ~ "Student",
        fl == "T" ~ "Teacher",
        fl == "C" ~ "Learning data",
        fl == "H" ~ "Help",
        fl == "L" ~ "Learning design",
        fl == "G" ~ "Assessment data"
    ) |> 
        factor(levels = rev(c(
            "Learning design", 
            "Teacher", 
            "Student",
            "Help", 
            "Learning data",
            "Assessment data"
            )))
}

```

From the node groups, describe how easy (or not) they are to measure.
This is completed by someone familiar to the data warehouse at the
institution:

- **DW** - *Data Warehouse*: Data is already collected and stored in the
  data warehouse of the institution.
- **O** - *Observable*: Node is not currently collected data, but would
  be possible to measure and collect with resourcing.
- **PO** - *Potentially Observable*: Node is not measured, and would be
  difficult to measure directly. It is possible to find a direct proxy
  and measure that instead, such as a survey instrument.
- **L** - *Latent*: Node is not measured and would be difficult to even
  find a direct proxy for the latent construct.

```{r node-measurability}
code_node_measurability <- function(node) {
    case_when(
        # Student
        node == "S.Att.SocCap" ~ "PO", # check
        node == "S.Att.Nro" ~ "O", 
        node == "S.Att.Eq" ~ "DW", 
        node == "S.Att.HSk" ~ "PO", 
        node == "S.Act.AIstrat" ~ "L",
        node == "S.Act.Fb" ~ "L",
        node == "S.Act.Fb.Post" ~ "L",
        str_detect(node, "^S\\.Act") ~ "PO", 
        str_detect(node, "^S\\.Kn") ~ "PO", 
        str_detect(node, "^S") ~ "L", 
        # Help
        node == "H.AI.Agent" ~ "O", # within system tool 
        node == "H.other.evil" ~ "L",
        node == "H.Peer" ~ "L",
        node == "H.AI.evil" ~ "PO", # Potentially observable
        node == "H.AI.evil.code" ~ "PO", # Potentially observable
        node == "H.AI.evil.write" ~ "PO",
        str_detect(node, "^H\\.AI") ~ "PO", # not trying to hide
        str_detect(node, "^H\\.Tu") ~ "PO",
        str_detect(node, "^H\\.Ment") ~ "PO",
        str_detect(node, "^H") ~ "L",
        # Assessment
        node == "Grade" ~ "DW",
        node == "Grade.Exam" ~ "DW",
        str_detect(node, "^Grade") ~ "O",
        # Learning data
        node == "CD.LMS" ~ "DW",
        str_detect(node, "^CD\\.LMS") ~ "O",
        node == "CD.Att" ~ "O",
        node == "CD.Str" ~ "DW", # check
        node == "CD" ~ "O", # check
        # Learning design
        str_detect(node, "^LD\\.Task") ~ "O",
        str_detect(node, "^LD") ~ "PO",
        # Teacher
        node == "T.Cp.UseAI" ~ "PO",
        node == "T.Aff.WrkLd" ~ "PO", 
        node == "T.Cp.Mark" ~ "PO", 
        str_detect(node, "^T\\.K") ~ "PO",
        str_detect(node, "^T") ~ "L",
        node == "Assessments" ~ "DW"
    ) |> 
        factor(levels = c("DW", "O", "PO", "L"))
}

coarsen_node_measurability <- function(ng_list) {

  x <- str_squish(as.character(ng_list))

  n_tot <- str_count(x, "[^;]+")
  n_obs <- str_count(x, "(?<![A-Z])(O|DW)(?![A-Z])")
  n_po  <- str_count(x, "(?<![A-Z])PO(?![A-Z])")
  n_lat <- str_count(x, "(?<![A-Z])L(?![A-Z])")

  out <- case_when(
    is.na(x) | n_tot == 0        ~ NA_character_,
    n_obs == n_tot               ~ "Observable",
    n_po  == n_tot               ~ "Potentially observable",
    n_lat == n_tot               ~ "Latent",
    n_po + n_lat == n_tot        ~ "Likely latent",
    n_obs + n_po == n_tot        ~ "Likely observable",
    TRUE ~ str_c("Some observable (", n_obs, " of ", n_tot, ")")
  )

  dyn <- sort(unique(out[str_starts(out, "Some observable") & !is.na(out)]))

  factor(out, levels = c("Latent", "Likely latent", dyn,
                         "Potentially observable", "Likely observable",
                         "Observable"))
}



```

``` {r attaching-node-measurability-to-dags}
for (i in 1:length(tdags)) {
    dag <- tdags[[i]] |> 
        activate(nodes) |> 
        mutate(measurable = code_node_measurability(name))
    
    tdags[[i]] <- dag
}
```

``` {r}
node_group_measurability <- 
    tdags |> 
    map(
        \(d)
        d |> 
            activate(nodes) |> 
            as_tibble() |> 
            select(name, measurable)) |>
    list_rbind(names_to = "Model") |> 
    mutate(node_group = code_node_group(name)) |> 
    group_by(node_group) |> 
    summarise(m = str_c(measurable, collapse = ";")) |> 
    mutate(measurability_node_group = coarsen_node_measurability(m)) |> 
    distinct(node_group, measurability_node_group) |> 
    arrange(measurability_node_group)
```

## Graph surgery

Functions outlined in the `gcm_helpers.R` script. Functions starting
with `gcm_` should operate on either a `dagitty` or `tidygraph` object.

Generating a range of edge metrics - comparisons across models. Want to
label edges in the unified DAG as 'contested' if they switch direction
between some models. Also want to examine the same idea for a list of
descendants. This is handled as a new table of edges - including all
possible edges.

```{r}
fill_matrix_NAs_with_colmean <- function(m) {
    stopifnot(is.matrix(m), is.numeric(m))
    
    cm <- colMeans(m, na.rm = TRUE)
    idx <- which(is.na(m), arr.ind = TRUE)
    m[idx] <- cm[idx[, "col"]]
    
    m
}
```

```{r graph-metric-functions}
########################################################
# graph metric functions ------------
# Not included in graph_helpers as they are less generic
########################################################

gcm_node_metrics <- function(g) {
    if (class(g)[[1]] == "dagitty") g <- dagitty_to_tidygraph(g)
    pagerank_no_outcome <- 
        g %N>%
        filter(name != "Grade") |> 
        mutate(pagerank_no_outcome = centrality_pagerank(damping = 0.85)) |> 
        as_tibble() |> 
        select(name, pagerank_no_outcome)
    
    g %N>%
        mutate(
            node_w = 1,
            node_group = code_node_group(name),
            node_type = code_node_type(name),
            measurability = code_node_measurability(name),
            degree = centrality_degree(),
            btw = centrality_betweenness(),
            btw_rank = rank(-btw, ties.method = "min"),   # rank (highest = 1)
            btw_rel_rank = (btw_rank - 1) / (n() - 1),
            btw_rel = 1 - btw_rel_rank,
            eigen = centrality_eigen(),
            pagerank = centrality_pagerank(damping = 0.5) # lowers the influence of sinks
        ) |> 
        as_tibble() |> 
        left_join(pagerank_no_outcome, by = "name") |> 
        mutate(pagerank_no_outcome = replace_na(pagerank_no_outcome, 0))
}

gcm_summary_metrics <- function(g){
    if (class(g)[[1]] == "dagitty") g <- dag_to_tg(g)
    pagerank_no_outcome <- 
        g %N>%
        filter(name != "Grade") |> 
        mutate(pagerank_no_outcome = centrality_pagerank(damping = 0.5)) |> 
        as_tibble() |> 
        select(name, pagerank_no_outcome)
    
    n.df <- g %N>%
        mutate(
            pgrnk = centrality_pagerank(damping = 0.5),
            ng = code_node_group(name),
            m = code_node_measurability(name)) |> 
        as_tibble() |> 
        left_join(pagerank_no_outcome, by = "name")
    
    # p_traced <- mean(n.df$m == "Traced")
    # p_untraced <- mean(n.df$m == "Untraced")
    # pgrnk_traced <- sum(n.df |> filter(m == "Traced") |> pull(pgrnk))
    # pgrnk_untraced <- sum(n.df |> filter(m == "Untraced") |> pull(pgrnk))
    # pgrnk_no_outcome_traced <- sum(n.df |> filter(m == "Traced") |> pull(pagerank_no_outcome))
    # pgrnk_no_outcome_untraced <- sum(n.df |> filter(m == "Untraced") |> pull(pagerank_no_outcome))
    
    p_dw <- mean(n.df$m == "DW")
    p_o <- mean(n.df$m == "O")
    p_po <- mean(n.df$m == "PO")
    p_l <- mean(n.df$m == "L")
    measure_metric <- (3 * p_dw + 2 * p_o + 1 * p_po ) / 3
    
    graph_metrics <- list(
        order     = gorder(g), # number of nodes
        size      = gsize(g), # number of edges
        density   = edge_density(g),
        diameter  = diameter(g), # longest shortest path
        mean_path = mean_distance(g),
        clustering= transitivity(g, type = "global"),
        acyclic = igraph::is_acyclic(as.igraph(g)),
        measurability = measure_metric,
        p_dw = p_dw,
        p_o = p_o,
        p_po = p_po,
        p_l = p_l
    )
    graph_metrics
}

```

```{r fancy-edge-weight}
all_nodes <- 
    imap(tdags, \(g,m) gcm_nodelist(g)) |> 
    list_rbind(names_to = "Model")

all_edges <- 
    imap(tdags, \(g, m) gcm_edgelist(g) |> mutate(Model = m)) |> 
    list_rbind() |> 
    adorn_switches_exist(from, to) |> 
    mutate(edge_alignment = if_else(switched, "contested", "aligned"))

all_possible_edges <- 
    imap(tdags, \(g, m) gcm_possible_edges(g)) |> 
    list_rbind(names_to = "Model")

max_possible_edges <- 
    all_possible_edges |> 
    filter(from != to) |> 
    count(from, to) |> 
    rename(max_n = n)

all_descendants <- 
    imap(tdags, \(g,m) gcm_all_descendants_edge_list(g)) |> 
    list_rbind(names_to = "Model") |> 
    adorn_switches_exist(from, to) |> 
    mutate(descendant_alignment = if_else(switched, "descendant:contested", "descendant:aligned")) |> 
    select(-switched)

all_possible_edges_with_metrics <- 
    all_possible_edges |> 
    filter(from != to) |> 
    left_join(
        all_edges |> 
            mutate(in_graph = TRUE),
        by = c("Model", "from", "to")) |> 
    mutate(in_graph = replace_na(in_graph, FALSE)) |> 
    left_join(
        all_descendants |> 
            mutate(descendant_in_graph = TRUE),
        by = c("Model", "from", "to")) |> 
    mutate(descendant_in_graph = replace_na(
        descendant_in_graph, FALSE))

all_edges_sum <- 
    all_possible_edges_with_metrics |>
    group_by(from, to) |> 
    summarise(
        n = sum(in_graph),
        n_descendant = sum(descendant_in_graph),
        N = n(),
        direct_alignment = mean(edge_alignment == "aligned", na.rm = T),
        descendant_alignment = mean(
            descendant_alignment == "descendant:aligned", na.rm = T),
        .groups = "drop") |> 
    mutate(
        p_direct = n / N, 
        p_descendant = n_descendant / N) |> 
    arrange(desc(n)) 


all_edges_in_models_with_metrics <- 
    all_edges_sum |> 
    filter(n > 0) |> 
    inner_join(max_possible_edges,
               by = c("from", "to")) |> 
    mutate(
        edge_p_w = n / max_n,
        edge_dp_w = n_descendant / max_n) |> 
    mutate(edge_w = n) |> 
    adorn_switches_exist(from, to)

```

## Creating merged graph

```{r making-merged-fine-graph}
merged_edges <- 
    all_edges_in_models_with_metrics |> 
    mutate(path_type = "causal")

merged_nodes <- 
    all_nodes |> 
    count(name) |> 
    rename(node_w = n) |> 
    mutate(node_group = code_node_group(name)) |> 
    mutate(node_type = code_node_type(name)) |> 
    mutate(measurable = code_node_measurability(name)) |> 
    left_join(name_node_nice_name, by = "name")

merged_all_graph <- tbl_graph(nodes = merged_nodes, edges = merged_edges)

```

## Coarsen graphs into CDAGs

This is based on the coding of the nodes into node groups.

Here we merge nodes into node groups before generating plot later. This
has the potential to introduce loops depending on how the constructs
were built, so we need to check the causal consistency between the
fine-grained graph and the clustered / merged / higher-level graph.

```{r creating-CDAGs}
clean_node_group_cluster_graph <- function(cdag) {
    # After coarsening with code_node_group cleans up some of the 
    # names and nodes
    dag_out <- 
        cdag |> 
        activate(edges) |> 
        mutate(path_type = "causal") |> 
        activate(nodes) |> 
        mutate(node_type = code_node_type(name))  |> 
        left_join(
            node_group_measurability, 
            by = c("name" = "node_group")
        ) |> 
        select(-any_of(c("nice_name", "node_group"))) |> 
        left_join(
            node_group_nice_name,
            by = c("name" = "node_group")) 
    
    # Attaching node lists
    obs_nodes <- dag_out |> 
        activate(nodes) |> 
        filter(
            measurability_node_group %in% c(
                "Observable", 
                "Likely observable")) |> 
        pull(name)
    dag_out$observable <- obs_nodes
    
    pot_obs_nodes <- dag_out |> 
        activate(nodes) |> 
        filter(
            measurability_node_group %in% c(
                "Observable", 
                "Likely observable",
                "Potentially observable")) |> 
        pull(name)
    dag_out$potentially_observable <- pot_obs_nodes
    
    return(dag_out)
}

node_group_cluster_graph_from_merged <- 
    gcm_cluster_graph(merged_all_graph, code_node_group) |> 
    clean_node_group_cluster_graph()
    

CDAGs <- 
    map(tdags, 
        \(g)
        gcm_cluster_graph(g, code_node_group) |> 
            clean_node_group_cluster_graph()) 

node_group_cluster_graph_from_clusters <- 
    gcm_merge(
        gcm_merge(
            gcm_merge(CDAGs[[1]], CDAGs[[2]]),
            gcm_merge(CDAGs[[3]], CDAGs[[4]])
        ),
        gcm_merge(
            gcm_merge(CDAGs[[5]], CDAGs[[6]]),
            gcm_merge(CDAGs[[7]], CDAGs[[8]])
        )
    )

```

## Graph metrics

```{r path-metrics}
all_paths_summary <-
    imap(ddags, \(d, n) gcm_count_paths(d)) |>
    list_rbind(names_to = "Model")

all_paths <- 
    imap(
        ddags,
        \(d, n) bind_cols(
            tibble(Model = n),
            dagitty::paths(d, limit = 5000)) |> 
            group_by(Model) |> 
            summarise(
                paths = list(paths),
                n_paths = n(),
                n_open = sum(open),
                p_open = mean(open)
            ))|> 
    list_rbind()


dag_complexity_metrics <- 
    imap(tdags, \(g, n)  as_tibble(gcm_summary_metrics(g))
    ) |> 
    list_rbind(names_to = "Model") |> 
    inner_join(
        imap(
            ddags, 
            \(d, n) 
            tibble(Model = n) |> 
                mutate(
                    n_adj_sets = length(dagitty::adjustmentSets(d)))) |> 
            list_rbind(), 
        by = "Model"
    ) |> 
    inner_join(all_paths_summary, by = "Model")
```

```{r building-all-nodes-with-metrics}
all_nodes_with_metrics <- 
    imap(tdags, \(g,n) gcm_node_metrics(g)) |> 
    list_rbind(names_to = "Model")
```

```{r joining-session-data-to-gcm-metrics}
gcm_metrics <- 
    inner_join(session_data, dag_complexity_metrics, by = "Model") 
```

# Visualisations

```{r table-of-node-type-counts}
all_nodes_with_metrics |> 
    count(node_type) |> 
    arrange(node_type) |> 
    knitr::kable(caption = "Counts of the nodes according to type")
```

## Modelling sessions, time and perceptions

```{r figure-session-completions-and-gcm-representation}
# denser option - radial arc bars for observable
library(scales)
library(ggforce)
library(patchwork)
library(cowplot)


# Main plot - hide all legends
p_validity_and_time <- gcm_metrics |> 
    mutate(
        x_duration = as.numeric(Modelling.duration, units = "minutes"),
        y_validity = GCM.validity
    ) |> 
    mutate(acyclic = if_else(acyclic, "Yes", "No")) |> 
    ggplot(aes(
        color = acyclic,
        size = n_adj_sets
    )) +
    geom_point(aes(
        x = x_duration,
        y = y_validity)) +
    scale_y_continuous(
        limits = c(1, 4.05),
        breaks = c(1, 2, 3, 4),
        labels = c("Trivially", "Simply", "Adequately", "Well"),
        name = "Representation of GCM") +
    scale_x_continuous(
        limits = c(33, 52),  # was 55 to  Keep original limits for the data
        breaks = c(35, 40, 45,50),
        name = "Time spent constructing model (minutes)") +
    scale_size(range = c(2, 7),
               breaks = range(gcm_metrics$n_adj_sets),
               name = "Adjustment\nsets (count)") +
    scale_color_manual(values = c(hue1, hue2), name = "Acyclic?") +
    scale_fill_manual(values = c(hue1, hue2), guide = "none") +
    theme_minimal() +
    # coord_fixed(ratio = 2
    #             # , clip = "off"
    #             ) +
    theme(
        legend.position = "left",  # Hide legends from main plot
        # panel.grid.major.x = element_blank(),  # Remove vertical gridlines
        panel.grid.minor = element_blank(),
        plot.margin = margin(10, 10, 10, 10)
    )


p_visibility <- gcm_metrics |>
    mutate(
        cp_dw = p_dw,
        cp_o = p_dw + p_o,
        cp_po = p_dw + p_o + p_po,
        cp_l = p_dw + p_o + p_po + p_l # should equal 1
    ) |> 
    mutate(GCM.validity = case_when(
        GCM.validity == 3 & acyclic ~ GCM.validity - 0.03,
        GCM.validity == 3 & !acyclic ~ GCM.validity + 0.03,
        GCM.validity == 2.5 & acyclic ~ GCM.validity - 0.03,
        GCM.validity == 2.5 & !acyclic ~ GCM.validity + 0.03,
        TRUE ~ GCM.validity
    )) |> 
    pivot_longer(c(starts_with("cp_"), starts_with("p_")), names_to = "node_vis", values_to = "proportion") |> 
    mutate(node_visibility = case_when(
        str_detect(node_vis, "p_dw") ~ "Data Warehouse",
        str_detect(node_vis, "p_o") ~ "Observable",
        str_detect(node_vis, "p_po") ~ "Potentially observable",
        str_detect(node_vis, "p_l") ~ "Latent"
    ) |> 
        ordered(levels = c(
            "Latent",
            "Potentially observable",
            "Observable",
            "Data Warehouse"
        ))) |> 
    mutate(prop_type = if_else(str_detect(node_vis, "^cp_"), "cp", "p")) |> 
    pivot_wider(names_from = prop_type, values_from = proportion, id_cols = c(Model,acyclic, GCM.validity, node_visibility)) |> 
    mutate(acyclic = if_else(acyclic, "Yes", "No")) |>  
    ggplot(aes(
        color = acyclic
    )) +
    geom_rect(aes(
        xmin = cp - p,
        xmax = cp,
        ymin = GCM.validity - 0.03,
        ymax = GCM.validity + 0.03,
        fill = acyclic,
        alpha = node_visibility),
        color = "white") +
    scale_y_continuous(
        limits = c(1, 4.05),
        breaks = c(1, 2, 3, 4),
        labels = c("Trivially", "Simply", "Adequately", "Well"),
        name = "Representation of GCM",
        position = "right") +
    scale_alpha_discrete(name = "Data visibility") +
    scale_fill_manual(values = c(hue1, hue2), guide = "none") +
    theme_minimal() +
    # coord_fixed(ratio = 0.15
    #             # , clip = "off"
    #             ) +
    xlab("Proportion of nodes") +
    theme(
        legend.position = "right",  # Hide legends from main plot
        # panel.grid.major.x = element_blank(),  # Remove vertical gridlines
        panel.grid.minor = element_blank(),
        plot.margin = margin(10, 10, 10, 10),
        axis.title.y = element_blank(),
        axis.text.y = element_blank()
    )


p_time_and_representation_1 <- p_validity_and_time + p_visibility
p_time_and_representation_1

ggsave("figures/FIG--time-modelling-and-representation.jpg", 
       plot = p_time_and_representation_1,
       width = 10, height = 5, 
       dpi = 300)

```

As a table:

```{r table-of-gcm-metrics}
gcm_metrics |> 
    mutate(acyclic = if_else(acyclic, "Yes, DAG", "No, contained cycles")) |> 
    arrange(desc(Modelling.duration)) |> 
    select(Modelling.duration, GCM.validity, acyclic, n_adj_sets, p_dw, p_o, p_po, p_l) |> 
    knitr::kable(caption = "GCM session metrics")
```

## Merged graphs

```{r plotting-merged-graph}

merged_all_graph_layout <- create_layout(
    merged_all_graph %E>%
        mutate(switched = if_else(
            switched, "contested", "alligned")), # needs switched boolean
    layout = "sugiyama")

merged_all_graph_layout_adjusted <- 
    merged_all_graph_layout |> 
    mutate(
        # x = if_else(name == "S.Post.Fb",x + 20,x),
        y = if_else(name == "S.Act.Fb.Post",y + 4,y)
           ) |> 
    mutate(
        x = if_else(name == "S.Act",x + 20,x),
        y = if_else(name == "S.Act",y - 4,y)
           ) 


plot_fine_graph_from_layout <- function(lyt, weight_nodes = TRUE) {
    ppp <- 
        lyt |> 
        ggraph() 
    
    if (weight_nodes) {
        pp <- ppp +
            geom_node_point(aes(colour = node_type,
                                size = node_w)) +
            scale_size(name = "point-size") +
            geom_edge_fan(aes(
                alpha = edge_p_w * edge_w + edge_w, 
                edge_linetype = factor(switched)), 
                strength = 0.2,
                arrow = arrow(length = unit(1, 'mm'),
                              type = "closed"),
                start_cap = circle(3, 'mm'),
                end_cap = circle(3, 'mm'))
    } else {
        pp <- ppp +
            geom_node_point(aes(colour = node_type)) +
            geom_edge_fan(strength = 0.2, alpha = 0.5,
                          arrow = arrow(length = unit(1, 'mm'),
                                        type = "closed"),
                          start_cap = circle(1, 'mm'),
                          end_cap = circle(1, 'mm'))
            
    }
    pp +
        scale_colour_manual(
            values = node_type_colours,
            limits = names(node_type_colours),
            drop = FALSE
        ) +
        theme_graph() +
        scale_edge_alpha_continuous(range = c(0.15,0.8)) +
        scale_alpha_continuous(range = c(0.3,1)) +
        theme(legend.position = "none",
              plot.margin = margin(t = 0, r = 0, b = 0, l = 0, unit = "pt"))  +
        scale_x_continuous(expand = expansion(mult = 0.1)) 
        
}

g_plot_fine <- function(g, 
                        lyt = merged_all_graph_layout_adjusted, 
                        weight_nodes = TRUE) {
    create_layout(g, layout = "sugiyama") |> 
        left_join(
            lyt |> 
                select(x,y,name,node_group,node_type,node_w), 
            by = "name",
            suffix = c(".default", "")) |> 
        plot_fine_graph_from_layout(weight_nodes = weight_nodes)
}


pMergedFineLegendTall <- merged_all_graph_layout_adjusted |> 
    g_plot_fine() + 
    theme(legend.position = "right") +
    guides(size = "none", alpha = "none", edge_alpha = "none", edge_width = "none",
           color = guide_legend(ncol = 2)) +
    theme(legend.title = element_blank())

leg_tall <- cowplot::get_legend(pMergedFineLegendTall)
leg_tall_plot <- cowplot::ggdraw(leg_tall)

```

```{r plotting-coarse-graph}
set.seed(1)
merged_coarse_layout <- 
    merged_all_graph |> 
    gcm_cluster_graph(code_node_group) |> 
    clean_node_group_cluster_graph() |> 
    activate(edges) |> 
    mutate(path_type = "causal") |> 
    create_layout(layout = "sugiyama") |>
    mutate(
        x = case_when(
            name == "Grade" ~ 20,
            name == "CD.Str" ~ 18.6,
            name == "CD.Att" ~ 21.5,
            name == "CD.LMS" ~ 23,
            name == "S.Kn" ~ 20,
            name == "S.Kn.Prior" ~ 20,
            name == "S.Act" ~ 21,
            name == "S.EM" ~ 23,
            name == "S.Cp" ~ 21.6,
            name == "S.Att" ~ 22,
            # name == "T.EM" ~ 30,
            # name == "T.Kn" ~ 31,
            name == "H.Gen" ~ 19,
            name == "H.AI.evil" ~ 22.4,
            name == "H.Cheat" ~ 22.5,
            name == "H.AI.good" ~ 19.4,
            name == "H.Tut" ~ 19,
            name == "LD.Task" ~ 18,
            name == "LD.Gen" ~ 18,
            name == "T.Cp" ~ 17,
            name == "T.Kn" ~ 17.3,
            name == "T.EM" ~ 18,
            # name == "H.AI.good" ~ 27,
            TRUE ~ x),
        y = case_when(
            name == "Grade" ~ 1,
            name == "CD.Str" ~ 2,
            name == "CD.Att" ~ 2.5,
            name == "CD.LMS" ~ 3,
            name == "S.Kn" ~ 8,
            name == "S.Kn.Prior" ~ 12,
            name == "S.Act" ~ 6,
            name == "S.EM" ~ 10,
            name == "S.Cp" ~ 8,
            name == "S.Att" ~ 12,
            # name == "Grade" ~ y - 2.5,
            # name == "CD.Att" ~ 7,
            name == "LD.Task" ~ 6,
            # name == "CD.LMS" ~ 4,
            # name == "LD.Task" ~ y +2.5,
            name == "H.Gen" ~ 4,
            name == "H.Tut" ~ 6.5,
            name == "H.AI.good" ~ 5,
            name == "H.AI.evil" ~ 5,
            name == "H.Cheat" ~ 7,
            # name == "S.Cp" ~ y + 0.5,
            # name == "T.Cp" ~ 4.5,
            # name == "T.Kn" ~ y - 2,
            name == "T.EM" ~ 10,
            name == "T.Cp" ~ 9,
            name == "T.Kn" ~ 11,
            # name == "H.Tut" ~ y - 1,
            # name == "H.AI.good" ~ y - 0.8,
            # name == "H.AI.evil" ~ y - 2.5,
            # name == "H.Cheat" ~ y,
            # name == "S.EM" ~ 5,
            TRUE ~ y)
           )

plot_coarse_graph_from_layout <- function(lyt) {
    nodes <- get_nodes()(lyt)
    
    causal_edges <- filter(get_edges()(lyt), 
                           path_type == "causal") |> 
        select(from, to, edge_w, path_type) |> 
        mutate(edge_w = 2*edge_w)
    
    confound_edges <- filter(get_edges()(lyt), 
                             path_type == "confound") |> 
        mutate(edge = if_else(from > to, 
                              str_c(to, "--", from), 
                              str_c(from, "--", to))) |>
        group_by(edge, path_type) |> 
        summarise(
            from = min(from),
            to = max(to),
            edge_w = sum(edge_w),
            .groups = "drop"
        ) |> 
        select(from, to, edge_w, path_type)
    
    tbl_graph(
        nodes = nodes,
        edges = bind_rows(causal_edges, confound_edges),
        directed = TRUE) |> 
        ggraph(layout = "manual", x = x, y = y) +
        geom_edge_fan(
            aes(alpha = edge_w, filter = path_type == "causal"), 
            strength = 0.8, linetype = 1,
            color = "grey20",
            arrow = arrow(length = unit(2.5, 'mm')),
            start_cap = circle(3, 'mm'),
            end_cap = circle(3, 'mm')) +
        geom_edge_arc(
            aes(alpha = edge_w, filter = path_type == "confound"),
            linetype = 2,
            color = "grey30",
            strength = -0.1,
            start_cap = circle(3, 'mm'),
            end_cap = circle(3, 'mm')) +
        geom_node_point(aes(colour = node_type
                            # , size = node_w
                            )) +
        scale_colour_manual(
            values = node_type_colours,
            limits = names(node_type_colours),
            drop = FALSE
        ) +
        theme_graph() +
        scale_edge_alpha_continuous(range = c(0.4,0.8)) +
        scale_x_continuous(expand = expansion(mult = 0.1)) +
        # scale_alpha_continuous(range = c(0.2,0.6)) +
        theme(legend.position = "none") 
}

g_plot_coarse <- function(g, lyt = merged_coarse_layout,
                          label = FALSE,
                          label_size = 5) {
    edge_attr <- g |> 
        activate(edges) |> 
        as_tibble() |> 
        names()
    
    if (!("path_type" %in% edge_attr)) {
        g <- g |> 
            activate(edges) |> 
            mutate(path_type = "causal")
    }
    if (!("edge_w" %in% edge_attr)) {
        g <- g |> 
            activate(edges) |> 
            mutate(edge_w = 1)
    }
    if (is.null(lyt)) {
        p <- create_layout(g, layout = "sugiyama") |> 
            plot_coarse_graph_from_layout()
    } else {
        p <- create_layout(g, layout = "sugiyama") |> 
            left_join(
                lyt |> 
                    select(x,y,
                           name, nice_name, 
                           node_type,node_w), 
                by = "name",
                suffix = c(".default", "")) |> 
            plot_coarse_graph_from_layout()    
    }
    
    if (label) {
        p <- p +
            geom_node_text(aes(
                colour = node_type,
                label = str_wrap(
                    nice_name, width = wrap_width)), 
                nudge_y = -0.2 * label_size, 
                alpha = 0.7,
                size = label_size)
    }
    
    return(p)
}

pCoarseMerged <- merged_coarse_layout |> 
    plot_coarse_graph_from_layout() +
    geom_node_text(aes(label = name))
pCoarseMerged
# ggsave("dag_merged_course.svg", pCoarseMerged, , width = 6, height = 6)
```

```{r plotting-a-highlighted-graph-functions}
plot_highlighted_fine_graph <- function(model = 1, LEGEND_POSITION = "none") {
    MODEL <- paste0("M", model)
    nodes_in_graph <- all_nodes |> 
        filter(Model == MODEL) |> 
        pull(name)
    edges_in_graph <- all_edges |> 
        filter(Model == MODEL) |> 
        select(from, to)
    
    edges_in_graph_indexed <- 
        edges_in_graph |> 
        mutate(
            from = match(from, merged_all_graph %N>% as_tibble() |> pull(name)),
            to = match(to, merged_all_graph %N>% as_tibble() |> pull(name))
        )
    
    g_highlight <- merged_all_graph %N>%
        mutate(
            node_highlight = if_else(
                name %in% nodes_in_graph, 2, 1
            )) %E>%
        left_join(
            edges_in_graph_indexed |> 
                mutate(edge_highlight = 2),
            by = c("from", "to")
        ) |> 
        mutate(edge_highlight = replace_na(edge_highlight, 1)) |> 
        create_layout("sugiyama")
    
    g_highlight |>
        ggraph() +
        geom_edge_fan(aes(
            # edge_linetype = factor(switched),
            alpha = edge_highlight), 
            strength = 0.2,
            arrow = arrow(length = unit(1, 'mm'),
                          type = "closed"),
            start_cap = circle(3, 'mm'),
            end_cap = circle(3, 'mm')) +
        geom_node_label(
            aes(label = str_wrap(nice_name, width = wrap_width), 
                color = node_type,
                size = node_w),
            alpha = 0.7,
            data = g_highlight |> 
                filter(node_highlight == 2),
            repel = T, max.overlaps = 30,
            show.legend = FALSE) +
        geom_node_point(aes(colour = node_type,
                            size = node_w,
                            alpha = node_highlight)) +
        scale_size(name = "point-size", range = c(2,3)) +
        scale_colour_manual(
            values = node_type_colours,
            limits = names(node_type_colours),
            drop = FALSE
        ) +
        theme_graph() +
        scale_edge_alpha_continuous(range = c(0.08,0.8)) +
        scale_alpha_continuous(range = c(0.1,1.0)) +
        # theme(legend.position = "none",
        #       plot.margin = margin(t = 0, r = 0, b = 0, l = 0, unit = "pt")) +
        theme(legend.position = LEGEND_POSITION) +
        guides(size = "none", edge_width = "none", alpha = "none", edge_alpha = "none",
               color = guide_legend(ncol = 1)) +
        theme(legend.title = element_blank())
}

merged_cluster_graph_ng_no_post_nodes <- 
    merged_all_graph |> 
    activate(nodes) |> 
    filter(!str_detect(name, "Post")) |> 
    gcm_cluster_graph(code_node_group) |> 
    clean_node_group_cluster_graph() |> 
    mutate(node_group = name) |> 
    activate(nodes) |> 
    left_join(merged_coarse_layout |> select(name, x, y), by = "name")



plot_highlighted_coarse_graph <- function(model = 1, LEGEND_POSITION = "none",
                                          repel_labels = TRUE) {
    merged_coarse_models <- 
        merged_cluster_graph_ng_no_post_nodes
    
    MODEL <- paste0("M", model)
    node_groups_in_graph <- all_nodes |> 
        filter(Model == MODEL) |> 
        mutate(name = code_node_group(name)) |> 
        distinct(name) |> 
        pull(name) 
    edges_in_coarse_graph <- all_edges |> 
        filter(Model == MODEL) |> 
        select(from, to) |> 
        mutate(across(from:to, code_node_group))
    
    edges_in_graph_indexed <- 
        edges_in_coarse_graph |> 
        mutate(
            from = match(from, 
                         merged_coarse_models |> 
                             as_tibble() |> 
                             pull(name)),
            to = match(to, 
                       merged_coarse_models %N>% 
                           as_tibble() |> 
                           pull(name))
        )
    
    edge_integers <-
        edges_in_graph_indexed |>
        mutate(p2p3 = (2 ** from) * (3 ** to) ) |>
        pull(p2p3)
    
    
    g_highlight <- 
        merged_coarse_models %N>%
        mutate(
            node_highlight = if_else(
                name %in% node_groups_in_graph, 1.0, 0.1
            )) %E>%
        mutate(
            edge_highlight = if_else(
                ((2 ** from) * (3 ** to)) %in% edge_integers,
                1.0, 0.1
            )
        )
    
    g_highlight |> 
        ggraph(layout = "manual",
               x = g_highlight %N>% pull(x),
               y = g_highlight %N>% pull(y)
        ) +
        geom_edge_fan(aes(
            # edge_linetype = factor(switched),
            alpha = edge_highlight), 
            strength = 0.2,
            arrow = arrow(length = unit(1, 'mm'),
                          type = "closed"),
            start_cap = circle(3, 'mm'),
            end_cap = circle(3, 'mm')) +
        geom_node_point(aes(colour = node_type,
                            size = node_w,
                            alpha = node_highlight)) +
        geom_node_text(
            aes(label = str_wrap(nice_name, width = wrap_width),
                color = node_type,
                alpha = node_highlight,
                size = node_w
                ),
            repel = repel_labels, 
            max.overlaps = 30,
            show.legend = FALSE) +
        scale_size(name = "point-size", range = c(2,3)) +
        # scale_color_npg() +
        scale_colour_manual(
            values = node_type_colours,
            limits = names(node_type_colours),
            drop = FALSE
        )+
        theme_graph() +
        scale_edge_alpha_identity() +
        scale_alpha_identity() +
        # theme(legend.position = "none",
        #       plot.margin = margin(t = 0, r = 0, b = 0, l = 0, unit = "pt")) +
        theme(legend.position = LEGEND_POSITION) +
        guides(size = "none", edge_width = "none", alpha = "none", edge_alpha = "none",
               label = "none",
               color = guide_legend(ncol = 1)) +
        theme(legend.title = element_blank())
}
```

```{r figure--highlighted-coarse-model}
# Model used in paper
md_for_paper <- 5
p_merged_highlighted_fine <- plot_highlighted_fine_graph(
    md_for_paper, 
    LEGEND_POSITION = "right")
p_merged_highlighted_coarse <- plot_highlighted_coarse_graph(
    md_for_paper, LEGEND_POSITION = "right")
p_merged_highlighted_fine + ggtitle(str_c("Model  ",md_for_paper , ", original"))
p_merged_highlighted_coarse + ggtitle(str_c("Model  ",md_for_paper , ", coarsened")) 
# ggsave(plot = p_merged_highlighted_fine, 
#        filename = "figures/FIG--merged-fine-with-1-highlighted.jpg",
#        width = 8, height = 5, dpi = 600)
# 
# ggsave(plot = p_merged_highlighted_coarse, 
#        filename = "figures/FIG--merged-coarse-with-1-highlighted.jpg",
#        width = 8, height = 5, dpi = 600)
```

```{r show-and-save-highlighted-plots-function}
# Generating all
p_fine_show_and_save <- function(n){
    p_temp <- plot_highlighted_fine_graph(n, LEGEND_POSITION = "right") + ggtitle(paste0("DAG ", n, " highlighted"))
    ggsave(plot = p_temp, 
           filename = paste0("figures/plt--merged-fine-with-", n, "-highlighted.png"),
           width = 8, height = 5, dpi = 600)
    return(p_temp)
}
p_coarse_show_and_save <- function(n, title = T){
    p_temp <- plot_highlighted_coarse_graph(
        n, LEGEND_POSITION = "right") 
    
    if (title) {
        p_temp <- p_temp +
        ggtitle(paste0("CDAG ", n, " highlighted"))
    }
    
    
    ggsave(plot = p_temp, 
           filename = paste0("figures/plt--merged-coarse-with-", n, "-highlighted.png"),
           width = 8, height = 5, dpi = 600)
    return(p_temp)
}
```

``` {r}
p_coarse_show_and_save(2, title = F)
```

## Visualising cluster DAGs

Note the appearance of feedback loops - these are between S.Cp and S.Kn,
but also S.Cp and H.AI.evil. These metacognitive skills are clearly seen
as important feedback processes in this system.

Most common node group patterns - nodes in at least half the models,
then looking at the resulting graph of node-groups.

```{r building-consensus-dag}
node_list_majority_of_models <- 
    all_nodes |> 
    count(name) |> 
    filter(n >= 4) |> 
    pull(name)

coarse_filtered_graph_temp <- 
    merged_all_graph |> 
    activate("nodes") |> 
    filter(name %in% node_list_majority_of_models)

coarse_filtered_graph <- 
    coarse_filtered_graph_temp |> 
    gcm_cluster_graph(code_node_group)

```

Weighting by num of models instead This one uses only nodes that are in
the majority of models, then coarsens the graph, then aggregates edges
based on the number of models that they appear in. The edges must also
appear in at least 2 models

```{r building-consensus-coarse-dag}
node_list_majority_of_models <- 
    all_nodes |> 
    count(name) |> 
    filter(n >= 4) |> 
    pull(name)

coarse_filtered_graph_ng_filled <- 
    tbl_graph( # only nodes that are in most models (4 or more)
        nodes = all_nodes |> 
            filter(name %in% node_list_majority_of_models) |> 
            mutate( # but then coarsen into node groups
                node_group = code_node_group(name),
                   node_type = code_node_type(name)) |> 
            count(node_group, node_type) |> 
            rename(node_w = n),
        edges = all_edges |> # Only edges between node groups...
            mutate(across(from:to, code_node_group)) |>
            distinct(Model, from, to) |> 
            count(from, to) |>
            # filter(n > 1) |> # ...that occur in at least two models
            filter(from %in% code_node_group(node_list_majority_of_models)) |>
            filter(to %in% code_node_group(node_list_majority_of_models)) |>
            rename(edge_w = n)) |> 
    activate("edges") |> 
    adorn_switches_exist(from, to) |> 
    activate("nodes") |> 
    inner_join(node_group_nice_name)
```

All edges between node groups from the majority nodes

```{r plot-coarse-consensus-fuller}
lyt_coarse_filtered_ng_full <- create_layout(
    coarse_filtered_graph_ng_filled, 
    layout = "sugiyama")

p_top_node_groups_full <- ggraph(lyt_coarse_filtered_ng_full) +
    geom_edge_fan(aes(edge_linetype = factor(switched), alpha = edge_w),
                  strength = 0.8,
                  arrow = arrow(length = unit(2, 'mm'), type = "closed"),
                  start_cap = circle(3, 'mm'),
                  end_cap = circle(3, 'mm')) +
    geom_node_point(aes(colour = node_type, size = node_w)) +
    scale_colour_manual(
            values = node_type_colours,
            limits = names(node_type_colours),
            drop = FALSE
        ) +
        theme_graph() +
        scale_edge_alpha_continuous(range = c(0.15,0.8)) +
        scale_alpha_continuous(range = c(0.2,0.9)) +
        theme(legend.position = "none") +
    scale_size(range = c(3,6)) +
    geom_node_text(aes(label = str_wrap(nice_name, width = wrap_width), colour = node_type), 
                   alpha = 0.9, 
                   nudge_y = -.3,
                   repel = F) +
    xlim(c(
        min(lyt_coarse_filtered_ng_full$x) - 0.15, 
        max(lyt_coarse_filtered_ng_full$x) + 0.15))

p_top_node_groups_full

ggsave("figures/FIG--consensus-sub-model-full.jpg",
       p_top_node_groups_full,
       dpi = 900,
       height = 4, width = 6)
```

Filtered - edges must be in at least 2 models

```{r plot-coarse-stronger-consensus}
lyt_coarse_filtered_ng_filtered <- 
    create_layout(
    coarse_filtered_graph_ng_filled |> 
        activate("edges") |> 
        filter(edge_w > 1) |> 
        mutate(
            switched = map_lgl(row_number(), function(i) {
                from_i <- .E()$from[i]
                to_i   <- .E()$to[i]
                any(from == to_i & to == from_i)
            })
        ) , 
    layout = "sugiyama") |> 
    select(-x, -y) |> 
    inner_join(lyt_coarse_filtered_ng_full |> select(x, y, node_group, node_type))

p_top_node_groups_filtered <- 
    ggraph(lyt_coarse_filtered_ng_filtered) +
    geom_edge_fan(aes(edge_linetype = factor(switched), alpha = edge_w, ), # alpha = edge_w, 
                  strength = 0.8,
                  arrow = arrow(length = unit(2, 'mm'), type = "closed"),
                  start_cap = circle(3, 'mm'),
                  end_cap = circle(3, 'mm')) +
    geom_node_point(aes(colour = node_type, size = node_w)) +
    scale_colour_manual(
            values = node_type_colours,
            limits = names(node_type_colours),
            drop = FALSE
        ) +
    theme_graph() +
        scale_edge_alpha_continuous(range = c(0.15,0.8)) +
        scale_alpha_continuous(range = c(0.2,0.9)) +
        theme(legend.position = "none") +
    scale_size(range = c(3,6)) +
    geom_node_text(aes(label = str_wrap(nice_name, width = wrap_width), colour = node_type), 
                   alpha = 0.9, nudge_y = -.3, repel = F) +
    xlim(c(min(lyt_coarse_filtered_ng_filtered$x) - 0.15, max(lyt_coarse_filtered_ng_filtered$x) + 0.15))

p_top_node_groups_filtered

ggsave("figures/FIG--consensus-sub-model-filtered.jpg", 
       p_top_node_groups_filtered, 
       dpi = 900, 
       height = 4, width = 6)
```

```{r table-of-nodes-with-metrics}
# all nodes
all_nodes_with_metrics |> 
    inner_join(name_node_nice_name) |> 
    inner_join(node_group_nice_name |> rename(g_nice_name = nice_name)) |> 
    select(name, nice_name, g_nice_name, node_group, Model, pagerank_no_outcome, pagerank, btw_rel_rank) |> 
    arrange(name, Model) |> 
    knitr::kable()

```

## Node importance and measurability

```{r building-node-groups-aggregated-metrics}
all_node_groups_aggregated_metrics <- 
    all_nodes_with_metrics |> 
    group_by(node_group) |> 
    mutate(
        measurability_node_group = coarsen_node_measurability(measurability)
    ) |> 
    group_by(Model, node_group, node_type, measurability_node_group) |> 
    summarise(
        p_traced = mean(measurability == "DW"),
        measurability_metric = (3 * mean(measurability == "DW") +
                                    2 * mean(measurability == "O") +
                                    mean(measurability == "PO")) / 3,
        pagerank = sum(pagerank),
        pagerank_no_outcome = sum(pagerank_no_outcome),
        btw_rel = sum(btw_rel),
        .groups = "drop_last") |> 
    group_by(node_group, node_type) |> 
    mutate(
        n = n(),
        overall_pagerank = sum(pagerank / 8),
           overall_pagerank_no_outcome = sum(pagerank_no_outcome / 8),
           overall_btw = sum(btw_rel)) |> 
    ungroup()


```

#### Figure -- Node importance and data availability

```{r figure-node-importance-and-measurability}
m_pal <- c("darkslateblue",
    "dodgerblue","orange","deeppink")
    
d_ng_measurability_importance <- 
    all_node_groups_aggregated_metrics |> 
    filter(node_group != "Grade") |>
    inner_join(node_group_nice_name, by = "node_group") |>
    mutate(node_group = nice_name) |>
    group_by(node_group) |>
    mutate(m_group = mean(measurability_metric)) |>
    ungroup() |> 
    mutate(node_group = fct_reorder(node_group, pagerank_no_outcome)) |> 
    arrange(desc(node_group)) |> 
    filter(n > 2) # need at least half the models with node group

p_top_node_pagerank_no_outcome_measurability_discrete <- 
    d_ng_measurability_importance |>
    mutate(
        measurability = case_when(
            measurability_metric < (1/6) ~ "Latent", 
            measurability_metric < 0.5 ~ "Potentially observable",
            measurability_metric < (5/6) ~ "Observable",
            TRUE ~ "Data warehouse") |> 
            factor(levels = c(
                "Latent", 
                "Potentially observable", 
                "Observable", 
                "Data warehouse"
                ))) |> 
    arrange(Model, node_group) |>
    ggplot(aes(x = pagerank_no_outcome, y = node_group)) +
    geom_boxplot(aes(fill = m_group), colour = "lightgray", size = 0.4, alpha = 0.7) +
    # geom_line(aes(x = pagerank_no_outcome, y = node_group, group = Model), alpha = 0.6) +
    geom_point(aes(colour = measurability), alpha = 0.9) +
    scale_color_manual(values = m_pal, # or retro four
                       labels = c("Latent","Potentially\nobservable", "Observable", "Data\nWarehouse"),
                       name = ""
    ) +
    scale_fill_gradientn(
        colours = m_pal,
        breaks = c(0, 1/3, 2/3, 5/6)
        ) +
        theme_minimal() +
            ylab("Node group") +
            xlab("Node groups combined Centrality (Pagerank)") +
            # geom_rug(aes(colour = m_dist_to_trace), alpha = 0.6, sides = "b") +
            # coord_cartesian(clip = "off") +
            guides(fill = "none",
                   color = guide_legend(nrow = 1, byrow = FALSE)) +
            theme(legend.position = "bottom",
          legend.direction = "vertical",
                  legend.justification = "left",
                  legend.text  = element_text(size = 8),
                  legend.key.width = unit(1, "cm"),
                  axis.ticks.y = element_blank(),
                  axis.line.y = element_blank(),
                  panel.grid.major.y = element_blank(),
          panel.grid.minor.y = element_blank())
    
p_top_node_pagerank_no_outcome_measurability_discrete

# 
# # extract legend
# leg <- get_legend(p_top_node_pagerank_no_outcome_measurability_discrete)
# 
# # arrange with relative heights (legend gets a small slice)
# p_node_importance_with_measurability <- (wrap_elements(leg) /
#     (p_top_node_pagerank_no_outcome_measurability_discrete + theme(legend.position = "none"))) +
#   plot_layout(heights = c(0.2, 1))


ggsave("figures/FIG--node_pagerank_no_outcome_measurability.jpg",
       p_top_node_pagerank_no_outcome_measurability_discrete, 
       width = 5, 
       height = 4, 
       dpi = 900)
```

As table

```{r table-node-measurability}
all_nodes_with_metrics |> 
    distinct(name, measurability) |> 
    janitor::tabyl(measurability) |> 
    janitor::adorn_totals("row") |> 
    knitr::kable(caption = "Proportions of nodes by measurability")
```

## Tables of nodes

Node group measurability was coded according to:

- Observed: All nodes were either in the data warehouse or observable.
- Observable: No nodes were latent.
- Likely latent: Nodes were a mix of latent or potentially observable.
- Latent: All nodes were latent.

```{r appendix-node-groups-table}
# could add distance to traced
node_groups_table <- 
    all_nodes_with_metrics |>
    inner_join(name_node_nice_name, by = "name") |> 
    mutate(name =  nice_name) |> 
    left_join(
        node_group_nice_name |> 
            rename(ng_nice_name = nice_name), by = "node_group") |> 
    mutate(node_group = ng_nice_name) |> 
    group_by(node_type, node_group) |>
    mutate(measurability = coarsen_node_measurability(measurability)) |> 
    group_by(node_type, node_group, measurability) |>
    summarise(
        nodes = str_c(unique(name), collapse = "; "),
        n = sum(node_w),
        pagerank = round(sum(pagerank_no_outcome) / 8, 2)
        # ,btw = round(sum(btw_rel), 3)
    ) |> 
    mutate(pagerank = if_else(node_group == "Assessment grade", NA, pagerank)) |> 
    # mutate(node_type = case_when(
    #     node_type == "L" ~ "Learning Design",
    #     node_type == "S" ~ "Student",
    #     node_type == "T" ~ "Teacher",
    #     node_type == "C" ~ "Course Data",
    #     node_type == "G" ~ "Grade Data",
    #     node_type == "H" ~ "Help"
    # )) |> 
    arrange(desc(n), desc(pagerank)) |> 
    select(node_type, node_group, nodes, n, measurability, pagerank) |> 
    ungroup()

node_groups_table |> 
    arrange(node_type, node_group) |> 
    knitr::kable()
```

```{r results = 'asis'}
all_nodes_with_metrics |> 
    distinct(name) |> 
    nrow() |> 
    paste("distinct nodes\n") |> 
    cat()

all_nodes_with_metrics |> 
    distinct(node_group) |> 
    nrow() |> 
    paste("node groups.\n") |> 
    cat()
```

``` {r}
all_nodes_with_metrics |> 
    distinct(name, measurability) |> count(measurability) |> 
    knitr::kable()
```

What about grouping by the node category first, so it's hierarchical?

```{r}
node_types_table <- 
    node_groups_table |> 
    group_by(node_type) |> 
    mutate(nt = sum(n)) |> 
    ungroup() |> 
    arrange(desc(nt), desc(n)) |> 
    select(node_type, nt, node_group, nodes, n, measurability, pagerank) 
# |> 
    # mutate(node_type = str_c(node_type, " (", nt, ")")) |> 
    # select(-nt)

node_types_table |> 
    knitr::kable()
```

## Informal consistency: Edge coherence

How often do causal paths point in the same direction?

> Note on Model 8. This is a structurally different system, in which any
> misuse of AI effects are deliberately mediated through the task
> design. This is a case where the causal structure *should* be
> different to the others - we are describing a different system due to
> deliberate design choices.

```{r getting-fine-graph-edge-coherence}
# Nodes in fine graph
n_possible_edges <- nrow(all_possible_edges)
n_edges <- nrow(all_edges_in_models_with_metrics)

prob_edge_inclusion_given_existence <- 
    mean(all_edges_in_models_with_metrics$p_direct)

prob_universal_inclusion_given_existence <- 
    mean(all_edges_in_models_with_metrics$p_direct == 1)

prob_descendant_inclusion_given_existence <- 
    mean(all_edges_in_models_with_metrics$p_descendant)

prob_universal_descendant_inclusion_given_existence <- 
    mean(all_edges_in_models_with_metrics$p_descendant == 1)

prob_universal_descendant_inclusion_given_descendant_existance <- 
    all_edges_sum |> 
    filter(p_descendant > 0) |> 
    mutate(universal_descendant = p_descendant == 1) |> 
    pull(universal_descendant) |> 
    mean()
```

```{r eval = FALSE}
# Checking the above, using the method implemented with the cluster graphs - checks out!

all_nodes_fine_graphs <- 
    tdags |> 
    imap(
        \(g, m)
        gcm_nodelist(g)
    ) |> 
    list_rbind(names_to = "Model")

all_possible_edges_fine_graphs <- 
    tdags |> 
    imap(
        \(g, m)
        gcm_possible_edges(g)
    ) |> 
    list_rbind(names_to = "Model") |> 
    filter(from != to)

all_edges_fine_graphs <- 
    tdags |> 
    imap(
        \(g, m)
        gcm_edgelist(g)
    ) |> 
    list_rbind(names_to = "Model")

all_descendants_fine_graphs <- 
    tdags |> 
    imap(
        \(g, m)
        gcm_all_descendants_edge_list(g)
    ) |> 
    list_rbind(names_to = "Model")

all_edges_summary_fine_graphs <- 
    all_possible_edges_fine_graphs |> 
    left_join(
        all_edges_fine_graphs |> 
            mutate(is_edge = TRUE),
        by = c("Model", "from", "to")) |> 
    mutate(is_edge = replace_na(is_edge, FALSE)) |> 
    left_join(
        all_descendants_fine_graphs |> 
            mutate(is_descendant = TRUE),
        by = c("Model", "from", "to")) |> 
    mutate(is_descendant = replace_na(is_descendant, FALSE)) |> 
    group_by(from, to) |> 
    summarise(
        n_possible = n_distinct(Model),
        n_edge = sum(is_edge),
        n_descendant = sum(is_descendant),
        .groups = "drop"
    )

n_possible_edges_fine <- nrow(all_possible_edges_fine_graphs)
n_edges_fine <- nrow(all_edges_fine_graphs)

prob_edge_inclusion_given_existence_fine_graphs <- 
    all_edges_summary_fine_graphs |> 
    filter(n_edge > 0) |> 
    mutate(p_inc = n_edge / n_possible) |> 
    pull(p_inc) |> 
    mean()

prob_descendant_inclusion_given_existence_fine_graphs <- 
    all_edges_summary_fine_graphs |> 
    filter(n_edge > 0) |> 
    mutate(p_inc = n_descendant / n_possible) |> 
    pull(p_inc) |> 
    mean()



```

We examine the `r n_edges` edges from `r n_possible_edges` possible
edges across the 8 models, and in addition include the ancestor -
descendant pairs. From this we define a few events:

- $E_{ij}\in G_k$ for and edge belonging to a model
- $P(E_{ij} \in \mathbf{E}_{G_k} |V_i,V_j \in \mathbf{V}_{G_k} , E_{ij} \in \mathcal{E} )$ -
  the probability of inclusion of an edge in models it could be included
  in, given it appears in at least one model. This is
  `r prob_edge_inclusion_given_existence`
- $P(E_{ij} \in \mathbf{E}_{G_k}  \forall  G_k |V_i,V_j \in \mathbf{V}_{G_k} , E_{ij} \in \mathcal{E} )$ -
  the probability of universal inclusion of an edge in all models it
  could be included in, given it appears in at least one model. This is
  `r prob_universal_inclusion_given_existence`
- $P(V_j \in \mathrm{an}_{G_k}(V_i) | V_i,V_j \in \mathbf{V}_{G_k} , E_{ij} \in \mathcal{E} )$,
  or
  $P(V_i \rightsquigarrow_{G_k} V_j | V_i,V_j \in \mathbf{V}_{G_k} , E_{ij} \in \mathcal{E} )$ -
  the probability of universal inclusion of a descendant in all models
  it could be included in, given the edge appears in at least one model.
  This is `r prob_descendant_inclusion_given_existence`.
- $P(V_j \in \mathrm{an}_{G_k}(V_i)  \forall  G_k | V_i,V_j \in \mathbf{V}_{G_k} , E_{ij} \in \mathcal{E} )$ -
  the probability of universal inclusion of a descendant in all models
  it could be included in, given the edge appears in at least one model.
  This is `r prob_universal_descendant_inclusion_given_existence`. Could
  maybe use the squiggle notation here of mediated path?

We now perform the same analysis, but for the node-group cluster DAGs.
It is worth noting that these numbers are *lower* due to the node
clustering - it is much more likely that common nodes are found across
graphs.

```{r coarse-graph-edge-coherence}
# Nodes in cluster graph
all_nodes_cluster_graphs <- 
    CDAGs |> 
    imap(
        \(g, m)
        gcm_nodelist(g)
    ) |> 
    list_rbind(names_to = "Model")

all_possible_edges_cluster_graphs <- 
    CDAGs |> 
    imap(
        \(g, m)
        gcm_possible_edges(g)
    ) |> 
    list_rbind(names_to = "Model") |> 
    filter(from != to)

all_edges_cluster_graphs <- 
    CDAGs |> 
    imap(
        \(g, m)
        gcm_edgelist(g)
    ) |> 
    list_rbind(names_to = "Model")

all_descendants_cluster_graphs <- 
    CDAGs |> 
    imap(
        \(g, m)
        gcm_all_descendants_edge_list(g)
    ) |> 
    list_rbind(names_to = "Model")

all_edges_summary_cluster_graphs <- 
    all_possible_edges_cluster_graphs |> 
    left_join(
        all_edges_cluster_graphs |> 
            mutate(is_edge = TRUE),
        by = c("Model", "from", "to")) |> 
    mutate(is_edge = replace_na(is_edge, FALSE)) |> 
    left_join(
        all_descendants_cluster_graphs |> 
            mutate(is_descendant = TRUE),
        by = c("Model", "from", "to")) |> 
    mutate(is_descendant = replace_na(is_descendant, FALSE)) |> 
    group_by(from, to) |> 
    summarise(
        n_possible = n_distinct(Model),
        n_edge = sum(is_edge),
        n_descendant = sum(is_descendant),
        .groups = "drop"
    )

n_possible_edges_cluster <- nrow(all_possible_edges_cluster_graphs)
n_edges_cluster <- nrow(all_edges_cluster_graphs)

prob_edge_inclusion_given_existence_cluster_graphs <- 
    all_edges_summary_cluster_graphs |> 
    filter(n_edge > 0) |> 
    mutate(p_inc = n_edge / n_possible) |> 
    pull(p_inc) |> 
    mean()

prob_descendant_inclusion_given_existence_cluster_graphs <- 
    all_edges_summary_cluster_graphs |> 
    filter(n_edge > 0) |> 
    mutate(p_inc = n_descendant / n_possible) |> 
    pull(p_inc) |> 
    mean()

```

- $E_{[i][j]}\in \mathbf{E}_{{G_k^C}}=\mathbf{E}_{\varphi G_{k}}$ for
  and edge belonging to a clluster model
- $P(E_{[i][j]} \in \mathbf{E}_{\varphi G_k} |V_{[i]},V_{[j]} \in \mathbf{V}_{\varphi G_k} , E_{[i][j]} \in \mathcal{E}_\varphi )$ -
  the probability of inclusion of an edge in models it could be included
  in, given it appears in at least one model. This is
  `r prob_edge_inclusion_given_existence_cluster_graphs`
- $P(V_{[j]} \in \mathrm{an}_{\varphi G_k}(V_{[i]}) | V_{[i]},V_{[j]} \in \mathbf{V}_{\varphi G_k} , E_{[i][j]} \in \mathcal{E}_{\varphi} )$,
  or
  $P(V_{[i]} \rightsquigarrow_{\varphi G_k} V_{[j]} | V_{[i]},V_{[j]} \in \mathbf{V}_{\varphi G_{k}} , E_{[i][j]} \in \mathcal{E}_{\varphi} )$ -
  the probability of universal inclusion of a descendant in all models
  it could be included in, given the edge appears in at least one model.
  This is `r prob_descendant_inclusion_given_existence`.

# Causal degradation when clustering

This computes two metrics, the loss ratio of conditional independence
relationships (ci_loss) and the rate of introduction of spurious
ancestors when coarsening. CI loss indicates how often a conditional
independence relationship in the fine graph did not have its analogue in
the coarsened graph. Spurious ancestors indicates the rate at which
merged nodes introduced new ancestors in the cluster DAG - violating L1
consistency (CITE - Annan???)

```{r L1-degradation-of-clustering}
# slow
if (FALSE) {
    clustering_L1_degradation <- 
        tdags |> 
        imap(
            \(g, m)
            {
            message(str_c(m, ":"))
            tibble(L1_degradation = gcm_cluster_L1_degradation(
                g, code_node_group,
                max_cond = 4)$degradation
            )}) |> 
        list_rbind(names_to = "Model")
    saveRDS(clustering_L1_degradation, "cache/L1-degradation.RDS")
} else {
    clustering_L1_degradation <-readRDS("cache/L1-degradation.RDS")
}
```

```{r L1-degradation-table}
clustering_L1_degradation |> 
    knitr::kable(
        caption = "L1-consistency metrics when coarsening GCMs",
        digits = 4)
```

This computes a single error metric - the % of lost identifiable pairs
when coarsening. So this is where the $P(Y|do(X))$ is identifiable in
the fine graph but $P(\varphi(Y)|do(\varphi(X)))$ is not identifiable in
the coarsened cluster graph.

```{r L2-degradation-of-clustering}
if (FALSE) {
    # adj_type = "all" for coverage, but slow
    # adj_type = "optimal" for best efficient estimation
    clustering_L2_degradation <- 
        tdags |> 
        imap(
            \(g,m)
            {
            message(str_c(m, ":"))
            tibble(
                L2_degradation = gcm_cluster_L2_degradation(
                    g, code_node_group,
                    adj_type = "all")$degradation
            )}) |> 
        list_rbind(names_to = "Model")
    saveRDS(clustering_L2_degradation, "cache/L2-degradation.RDS")
} else {
    clustering_L2_degradation <- readRDS("cache/L2-degradation.RDS")
}
```

```{r L2-degradation-table}
clustering_L2_degradation |> 
    knitr::kable(
        caption = "L2-consistency metrics when coarsening GCMs",
        digits = 4)
```

# Causal dissimilarity and joining similar models

## Computing L-dissimilarity

```{r dist-matrix-helper-funcion}
compute_dist_matrix <- function(
        model_list, # list of models
        .df, # function that computes distance between two models
        symmetric = TRUE,
        impute_undefined = FALSE) {
    n  <- length(model_list)
    nm <- names(model_list) 
    D <- matrix(0, n, n, dimnames = list(nm, nm))
    
    if (symmetric) {
        # fill upper triangle, mirror, zero diagonal
        for (i in seq_len(n - 1)) {
            for (j in (i + 1):n) {
                d <- .df(model_list[[i]], model_list[[j]])
                D[i, j] <- d
                D[j, i] <- d
                message(str_c("Computed distance between ",
                              nm[i], " and ", nm[j],
                              " as: ", d))
            }
        }
        Dsym <- D
    } else {
        for (i in seq_len(n)) {
            for (j in seq_len(n)) {
                if (i != j) {
                    D[i, j] <- .df(
                        model_list[[i]], 
                        model_list[[j]])
                }
            }
        }
        Dsym <- (D + t(D)) / 2
    }
    
    if (impute_undefined) {
        # First fill Dsym on mirror
        fill <- is.nan(Dsym) & !is.nan(t(Dsym)) 
        Dsym[fill] <- t(Dsym)[fill]
        if (any(is.nan(D))) {
            message("Imputing missing values")
            fit <- softImpute::softImpute(D, rank.max = 4, lambda = 0)
        D <- complete(D, fit)
        }
        if (any(is.nan()))
        fitSym <- softImpute::softImpute(Dsym, rank.max = 4, lambda = 0)
        Dsym <- complete(Dsym, fit)
    } 
    
    return(D)
}

compute_dist_matrix_from_fingerprints <- function(
        fp_list, # list of precomputed fingerprints
        vocab_list, # list of nodes in each model
        .df, # function that computes distance between two models from fps
        impute_undefined = FALSE) {
    n  <- length(fp_list)
    nm <- names(fp_list) 
    D <- matrix(0, n, n, dimnames = list(nm, nm))
    
    # fill upper triangle, mirror, zero diagonal
    for (i in seq_len(n - 1)) {
        for (j in (i + 1):n) {
            
            d <- .df(
                fp_list[[i]], 
                fp_list[[j]], 
                intersect(
                    vocab_list[[i]] |> pull(name),
                    vocab_list[[j]] |> pull(name)
                ))
            D[i, j] <- d
            D[j, i] <- d
        }
    }
    
    if (impute_undefined) {
        # First fill Dsym on mirror
        if (any(is.nan(D))) {
            message("Imputing missing values")
            fit <- softImpute::softImpute(
                D, 
                rank.max = 4, 
                lambda = 0)
            D <- complete(D, fit)
        }
    } 
    
    return(D)
}



matrix_fill_NAs_with_colmean <- function(M, include_ids = FALSE) {
    Mfilled <- M
    idNAs <- which(is.na(Mfilled), arr.ind = TRUE)
    Mfilled[idNAs] <- colMeans(Mfilled, na.rm = TRUE)[idNAs[, "col"]]
    if (include_ids) {
        return(list(
            M = Mfilled,
            NAs = idNAs
        ))
    } else {
        return(Mfilled)
    }
}

```

A note on clustering: Some methods of `hclust` have underlying geometric
assumptions (assume distances are Euclidean, which they are not), and
these are `ward.D`, `ward.D2`, `centroid` (UPGMC) and `median` (WPGMC).
Of the remainder `average` and `complete` are the best candidates. There
is another challenge when using these - the distance matrix must be
symmetric and in the case of the L2 distances it is not. So a decision
also needs to be made about how to do this along with the algorithm for
the hierarchical clustering.

- `average` (UPGMA) clustering method uses the mean pairwise distance
  between clusters, and the default for a non-Euclidean distance matrix.
  It is common when using Jaccard / dissimilarity. This works naturally
  with a matrix symmetrisation process of taking the average of the
  cross diagonals.
- `complete` is a stronger method that looks for clusters where every
  pair within it is consistent, no exceptions. This is a better 'worst
  case' scenario, and works more naturally with taking the maximum value
  of the cross-diagonals when forcing the distance matrix to be
  symmetric.

As such two methods of visualising the between model distances are used:
**average** and **worst-case**. Currently defaulting to the average
scenario.

```{r L0-between-models-distances-as-matrix}

if (recalculate_distance_matrices) {
    L0_between_model_dist_matrix <- 
        compute_dist_matrix(CDAGs, 
                            L0_d, 
                            symmetric = TRUE)
    saveRDS(L0_between_model_dist_matrix, 
            "cache/L0_models_dist_matrix.RDS")
} else {
    L0_between_model_dist_matrix <- 
        readRDS("cache/L0_models_dist_matrix.RDS")
}
```

```{r L1-distance-between-two-models}

if (recalculate_distance_matrices) {
    L1_between_model_dist_matrix <- 
        compute_dist_matrix(
            CDAGs,
            .df = L1_d)
    saveRDS(L1_between_model_dist_matrix, "cache/L1_models_dist_matrix.RDS")
} else {
    L1_between_model_dist_matrix <- readRDS("cache/L1_models_dist_matrix.RDS")
}
```

```{r L2-distance-between-two-models}

if (recalculate_distance_matrices) {
    L2_between_model_dist_matrix <- 
        compute_dist_matrix(
            CDAGs,
            .df = L2_d
        )
    
    saveRDS(L2_between_model_dist_matrix, "cache/L2_models_dist_matrix.RDS")
} else {
    L2_between_model_dist_matrix <- readRDS("cache/L2_models_dist_matrix.RDS")
}


```


```{r}
library(dendextend)
library(viridis)
dendo_pal <- rev(viridis(100))[1:100]
L1_hclust_models <- hclust(as.dist(L1_between_model_dist_matrix),
                           method = "average")
L2_hclust_models <- hclust(as.dist(L2_between_model_dist_matrix),
                           method = "average")

dend_L12 <- dendlist(
    L1_hclust_models |> as.dendrogram() |> highlight_branches_col(dendo_pal),
    L2_hclust_models |> as.dendrogram() |> highlight_branches_col(dendo_pal)
)

tanglegram(dend_L12, main_left = "L1-dissimilarity", main_right = "L2-dissimilarity",
           sort = TRUE, edge.lwd = 3, lwd = 2,
           common_subtrees_color_lines = TRUE, 
           highlight_distinct_edges  = FALSE)


```

```{r L12-clustering-CDAGs}
L12_dist_matrix <- 
    ((  L1_between_model_dist_matrix + 
         L2_between_model_dist_matrix) / 2)

L12_hclust_models <- 
    hclust(
    as.dist(L12_dist_matrix),
    method = "average") 

```

## Grouping CDAGs by L-similarity


We now combine all 3 levels of dissimilarity into an averaged measure, then perform a hierarchical clustering on the resultant distance matrix. 

```{r L012-clustering-CDAGs}
L012_dist_matrix <- 
    ((L0_between_model_dist_matrix + 
         L1_between_model_dist_matrix + 
         L2_between_model_dist_matrix) / 3)

L012_hclust_models <- 
    hclust(
    as.dist(L012_dist_matrix),
    method = "average") 

```

```{r figure--cdag-cluster-dendro-portrait}
png("figures/FIG--CDAG-dendrogram.png", 
    width = 10, height = 3, units = "in", res = 300)
L012_hclust_models |> 
    as.dendrogram() |> 
    set("labels_col", 
        value = c("#009E73", 
                  "#0072B2",
                  # "#56B4E9", 
                  "#D55E00", "#CC79A7"),
        k = 4) |> 
    set("branches_k_color", 
        value = c("#009E73", 
                  "#0072B2",
                  # "#56B4E9", 
                  "#D55E00", "#CC79A7"),
        k = 4) |> 
    plot(axes = F, horiz = F) 

dev.off()
par(mar = c(3, 3, 0.5, 0.5),   # default c(5.1, 4.1, 4.1, 2.1)
    xaxs = "i", yaxs = "i")    # drop the 4% range padding
png("figures/FIG--CDAG-dendrogram--3.png", 
    width = 10, height = 3, units = "in", res = 300)
L012_hclust_models |> 
    as.dendrogram() |> 
    set("labels_col", 
        value = c("#008595",
                  "#D55E00", 
                  "#CC79A7"),
        k = 3) |> 
    set("branches_k_color", 
        value = c("#008595",
                  "#D55E00",
                  "#CC79A7"),
        k = 3) |> 
    hang.dendrogram(hang = 0.1) |> 
    plot(axes = F, horiz = F) 

dev.off()

# Displaying in R Markdown as well. 
L012_hclust_models |> 
    as.dendrogram() |> 
    set("labels_col", 
        value = c("#008595",
                  "#D55E00", 
                  "#CC79A7"),
        k = 3) |> 
    set("branches_k_color", 
        value = c("#008595",
                  "#D55E00",
                  "#CC79A7"),
        k = 3) |> 
    hang.dendrogram(hang = 0.1) |> 
    plot(axes = F, horiz = F) 

```


```{r latent-projection-merge-graphs}
# thought - merge on latent projection of shared nodes only?
ng_counts <- 
    all_node_groups_aggregated_metrics |> 
    count(node_group)

gcm_projection_merge <- function(model_list, shared_nodes = NULL,
                                        additional_nodes = NULL) {
    # merges the latent projection of a list of models
    # each model is projected onto the shared set of nodes, 
    # with the other nodes being treated as latent
    if (is.null(shared_nodes)) {
        shared_nodes <- 
            model_list |> 
            map(
                \(m)
                gcm_nodelist(m) |> 
                    pull(name) 
            ) |> 
            reduce(intersect)
    }
    if (!is.null(additional_nodes)) {
        shared_nodes <- union(shared_nodes, additional_nodes)
    }
    
    merged_projections <- 
        model_list |> 
        map(
            \(m)
            gcm_projection(
                m, 
                latent_nodes = 
                    setdiff(gcm_nodelist(m, as_tibble = FALSE),
                            shared_nodes)) |> 
                activate(edges) 
            # |> 
            #     mutate(edge_w = replace_na(edge_w, 1)) |> 
            #     activate(nodes)
        ) |> 
        reduce(gcm_merge) |> 
        gcm_split_types() |> 
        activate(edges) |> 
        mutate(path_type = case_when(
            str_detect(type, "<->") ~ "confound",
            type =="->" ~ "causal",
            type =="<-" ~ "causal"
        ))
    
    return(merged_projections)
}
```

Looking at 3 clusters of causally similar models, and plotting a merging
of their latent projections using their shared node group clusters.

```{r plot-merged-latent-projections-of-similar-cluster-dags}
plot_latent_clusters_merged <- function(indexes, 
                                        nl = NULL,
                                        additional_nodes = NULL,
                                        ...) {
    CDAG.projection <- 
        CDAGs[indexes] |> 
        gcm_projection_merge(
            shared_nodes = nl,
            additional_nodes = additional_nodes) 
    
    CDAG.projection |> 
        g_plot_coarse(...) +
        ggtitle(str_c(
            "Merged latent projection of: ",  
            str_c("M", indexes, collapse = ", "))) 
}

most_common_ngs <- all_node_groups_aggregated_metrics |> 
    count(node_group) |> 
    filter(n > 2) |> 
    pull(node_group)
```

## Latent projections of CDAG groups

The latent projection of all models is relatively uninformative:

```{r}
plot_latent_clusters_merged(1:8, label = T)
```



Assisted by the clustering there is a natural way to partition the
models:

``` {r}
grouped_cdags <- list(
    C1plus2 = list(
        name = "metacognition-and-misuse",
        models = c(5,8, 2,4,7),
        description = "Exploring student metacognition skil and misuse of AI",
        colour = "#008595"
    ),
    C1 = list(
        name = "help-confounds--metacog-mediates",
        models = c(5,8),
        description = "Each help (tutor, AI, AI misuse) similarily influences student knowledge, and grade, mediated through metacognitive skill.",
        colour = "#009E73"
    ),
    C2 = list(
        name = "actions-leave-traces--metacog-skill-defends-AI-misuse",
        models = c(2, 4, 7),
        description = "Student actions leacing traces in LMS, lower metacognitive skill  drives Misuse of AI.",
        colour = "#0072B2"
    ),
    C2a = list(
        name = "actions-leave-traces--metacog-skill-defends-AI-misuse",
        models = c(2, 4),
        description = "Student actions leacing traces in LMS, lower metacognitive skill  drives Misuse of AI.",
        colour = "#5572B2"
    ),
    C3 = list(
        name = "course-design-and-teacher-capabilities",
        models = c(1),
        description = "Focused on course design, teacher capabilities. Did not identify knowledge to grade confounding.",
        colour = "#D55E00"
    ),
    C4 = list(
        name = "knowledge-action-feedback--hidden-misuse-confounding",
        models = c(3, 6),
        description = "Ongoing feedback between student actions and knowledge, hidden confounding between AI use and grades.",
        colour = "#CC79A7"
    )
)

grouped_cdags |> 
    map(as_tibble) |> 
    list_rbind() |> 
    select(-colour) |> 
    group_by(name, description) |> 
    summarise(models = str_c(models, collapse = "; ")) |> 
    mutate(name = name |> 
               str_replace("--", ";\n") |> 
               str_replace_all("-", " ")) |> 
    mutate(description = str_wrap(description, width = 50)) |> 
    knitr::kable()
```



```{r plotting-latent-projections-of-cdag-groups, eval = T}
grouped_cdags |> 
    map(
        \(l)
        plot_latent_clusters_merged(l$models, label = TRUE)
    )
```

``` {r}
save_latent_merged_cdag_rect_plot <- function(l, ...) {
    p <- plot_latent_clusters_merged(l$models, ...) + 
        ggtitle("") +
        coord_equal()+
        theme(
            panel.border = element_rect(
                colour = l$colour, fill = NA, linewidth = 1))
    ggsave(
        str_c("figures/projCDAG-", l$name,".png"), 
        plot = p,
        width = 4, height = 4,
        units = "in")
    
}


grouped_cdags |> 
    map(save_latent_merged_cdag_rect_plot)
```

``` {r}
pg1 <- plot_latent_clusters_merged(
    grouped_cdags$C1plus2$models, label = T) +
    scale_y_continuous(limits = c(-1, 10))+
    # scale_y_continuous(limits = c(-0.5, 10))+
    theme(
        panel.border = element_rect(
            colour = grouped_cdags$C1plus2$colour, 
            fill = NA, 
            linewidth = 2)) +
    ggtitle("")
pg1

ggsave(
    str_c("figures/projCDAG-labelled-", grouped_cdags$C1plus2$name,".png"), 
        plot = pg1,
        width = 6, height = 4,
        units = "in")
```


``` {r}
pg2 <- plot_latent_clusters_merged(
    grouped_cdags$C4$models, label = T) +
    scale_y_continuous(limits = c(-1, 10)) +
    theme(
        panel.border = element_rect(
            colour = grouped_cdags$C4$colour, 
            fill = NA, 
            linewidth = 2)) +
    ggtitle("")

pg2

ggsave(
    str_c("figures/projCDAG-labelled-", grouped_cdags$C4$name,".png"), 
        plot = pg2,
        width = 6, height = 4,
        units = "in")
```


``` {r}
plot_latent_clusters_merged(
    grouped_cdags$C3$models, label = T) +
    scale_y_continuous(limits = c(-0.5, 12)) +
    theme(
            panel.border = element_rect(
                colour = grouped_cdags$C3$colour, fill = NA, linewidth = 2))
```




# Simulation

Now we examine the feasibility of learning these models from data. A
typical counter to going to the effort of eliciting these models from
domain experts is that they can be learned from the data alone. However,
we have seen here that much of the data is not present in the
institutions ecosystem, with the nodes being deemed inherently latent
(`r round(100*(sum(all_nodes_with_metrics$measurability == "L") / nrow(all_nodes_with_metrics)), 1)`%),
potentially observable
(`r round(100*(sum(all_nodes_with_metrics$measurability == "PO") / nrow(all_nodes_with_metrics)), 1)`%),
observable but not currently in the data
(`r round(100*(sum(all_nodes_with_metrics$measurability == "O") / nrow(all_nodes_with_metrics)), 1)`%),
or already measured and stored in the current data ecosystem
(`r round(100*(sum(all_nodes_with_metrics$measurability == "DW") / nrow(all_nodes_with_metrics)), 1)`%).

Examining at a coarser grained level by looking at the node group
cluster graphs, we still see that a majority of node groups fall into
difficult to measure categories:

```{r}
node_groups_table |> 
    janitor::tabyl(measurability) |> 
    janitor::adorn_pct_formatting() |> 
    arrange(measurability)
```

## Creating a simulated data set

We design this experiment as follows. We ask the question can a causal
graph discovered from available data generate similar answers to the
problem of assuring assessment? This translates to getting an unbiased
estimate of Student Knowledge on Grade. As such 3 of the models are
excluded, lacking sufficient adjustment sets (2 models) or having broken
student knowledge into multiple nodes (1 model). For the "potentially
observable" nodes a single descendant node is added, so that
$N \rightarrow N_P$, a proxy measurement node that is not the exact node
measured but whose value depends only on the potentially observable node
and some additional noise. The set of proxy nodes for the potentially
observable ($PO$) nodes we denote ($\hat{PO}$). Each of the 5 remaining
GCMs, with their additional proxy measurement nodes, is used to generate
a synthetic data set. We assuming each node has an equal linear effect
on its descendants ($\beta=0.4$), generated using `simulateSEM` of the
`dagitty` package in R. This treats all variables as continuous with
mean 0 and variance 1.

```{r sim-parameter-choices}
sim_signal_strength <- list(
    "idyllic" = 0.9, 
    "typical" = 0.4, 
    "noisy" = 0.1)

tibble(Scenario = names(sim_signal_strength),
       Beta = as.numeric(sim_signal_strength)) |>
    mutate(
       `Variance due to each parent node` = Beta^2 / (1 + Beta^2)
       ) |> 
    knitr::kable(digits = 3)
```

We then apply a structure learning algorithm to each models simulated
data in three phases: learning from the currently measured data only
(only DW nodes), learning from all observable nodes (DW and O) nodes,
and finally learning from all potentially available data (DW, O and
$\hat{PO}$). This represents increasing levels of resourcing applied to
learning the underlying causal structure, given the model.

Note we are working under a local assumption that each model is the
'ground truth', but fully acknowledge that we do not know for this
system what the true model is. However, if the structure learning claims
of application to discover latent nodes are to be believed (CITE) then
we should see some reasonable convergence towards each locally true
model.



```{r simulate-data-from-dags}
# maybe use dagitty::simulateSEM() or dagitty::simulateLogistic()
# dagitty suggests using lavaan instead for SEM

# should add layout!!
# use same x, y from main graph


simSEM <- partial(simulateSEM, eps = 1, standardized = F, N = 1000)
# simSEM <- partial(simulateLogistic, eps = 1, N = 1000)

bn_fit_simulations <- function(
        dag, observable_nodes,
        force_observable = c("Grade", "S.Kn"),
        Beta = 0.4,
        proxy_nodes = NULL, # not implemented. Add to observable but also add noise
        proxy_noise_sd = NULL, # will use Beta if not specified
        blacklist = tibble(
            from = "Grade", 
            to = "S.Kn")) {
    
    observable_nodes <- union(observable_nodes, force_observable)
    # getting simulated data, and subsets
    d.sim.sem <- simSEM(dag, b.default = Beta)
    d.o <- as.data.frame(d.sim.sem |> select(all_of(observable_nodes)))
    if (!is.null(proxy_nodes)) {
        proxy_nodes <- setdiff(proxy_nodes, force_observable)
        # adding in proxy nodes, with additional noise due to using proxy measurement
        if (is.null(proxy_noise_sd)) proxy_noise_sd <- Beta
        d.proxy <- as.data.frame(d.sim.sem |> select(all_of(proxy_nodes)))
        d.proxy <- lapply(d.proxy, \(x) x + rnorm(length(x), 0, proxy_noise_sd))
    }
    
    # fitting BNets
    bn.o =    bnlearn::bn.fit(
        bnlearn::hc(
            d.o,
            blacklist = blacklist), d.o)
    bn.DAG.o <- bn_to_dagitty(bn.o)
    
    if (!is.null(proxy_nodes)) {
        d.opo <- bind_cols(d.o, d.proxy)
        bn.opo = bnlearn::bn.fit(
            bnlearn::hc(
                d.opo,
                blacklist = blacklist), d.opo)
        bn.DAG.opo <- bn_to_dagitty(bn = bn.opo)
    } else {
        d.opo <- NULL
        bn.opo <-  NULL
        bn.DAG.opo <-  NULL
    }
    
    return( # returns a named list of simmed data and models
        list(
            simdata = d.sim.sem,
            # Only observable nodes
            simdata.o     = d.o,
            bn.o    = bn.o,
            bn.DAG.o = bn.DAG.o,
            # Including Potentially Observable (PO), so ALL non-latent nodes
            simdata.opo  = d.opo,
            bn.opo = bn.opo,
            bn.DAG.opo = bn.DAG.opo
    ))
}

bn_to_dagitty <- function(bn) {
  edges <- bnlearn::arcs(bn) |> as_tibble()
  all_nodes <- bnlearn::nodes(bn)
  edge_str <- if (nrow(edges) > 0)
      str_c(edges$from, " -> ", edges$to, collapse = "; ") else ""
  dagitty(str_c("dag{ ",
                str_c(all_nodes, collapse = "; "),
                if (nchar(edge_str)) str_c("; ", edge_str) else "",
                " }"))
}

# Thinking of a simple dag X -> Y, we have 
# Var(Y) = b^2 * Var(X) + Var(eps) -- eps is the noise, and here it is set to 1 by default
# then Var(Y) = b^2 + 1
# Since the R^2 is the ratio of the variance explained by X:
# R^2 = Var(explained) / Var(total)
# R^2 = b^2 / (b^2 + 1)

dagitty_to_bn <- function(dag) {
  
  edges <- dagitty::edges(dag) |> 
    as_tibble() |> 
    filter(e == "->") |>  # directed edges only, excludes bidirected
    select(v, w) |> 
    rename(from = v, to = w)
  
  nodes <- unique(c(edges$from, edges$to))
  
  # Build empty network then set arcs
  net <- bnlearn::empty.graph(nodes)
  bnlearn::arcs(net) <- as.matrix(edges)
  net
}


causal_estimate_over_all_adj_sets <- function(
        dag, data, latent_nodes = NULL,
        exposure = "S.Kn", outcome = "Grade",
        x1_val = 1, x0_val = 0) {
    # This estimates P(Grade|S.Kn, Adjustment Set) over all adjustment sets
    # This will give a range of that the model suggests is the true effect
    if (is.null(dag)) {
        list(adjustment_set = NULL,
                estimate = NULL,
                std_error = NULL)
    }
    
    adj_sets <- adjustmentSets(
        dag, 
        exposure = exposure, outcome = outcome,
        type = "all")
    
    if (!is.null(latent_nodes)) {
        # remove any with latent nodes 
        adj_sets <- keep(adj_sets, \(x) !any(x %in% latent_nodes))
    }
    # Throw error if none!
    estimate_per_set <- map(adj_sets, function(adj_set) {
        if (length(adj_set) == 0) {
            message("No adjustment needed - regress directly")
            formula_str <- str_c(outcome, " ~ ", exposure)
        } else {
            formula_str <- str_c(outcome, " ~ ", exposure, " + ",
                                 str_c(adj_set, collapse = " + "))
        }
        if (class(data[[1]]) == "factor") {
            fit <- glm(as.formula(formula_str), data = data, family = binomial)
            x1_data <- data |> 
                mutate({{exposure}} := factor(x1_val, levels = c(-1,1)))
            x0_data <- data |> 
                mutate({{exposure}} := factor(x0_val, levels = c(-1,1)))
        } else {
            fit <- lm(as.formula(formula_str), data = data)
            x1_data <- data |> 
                mutate({{exposure}} := x1_val)
            x0_data <- data |> 
                mutate({{exposure}} := x0_val)
        }
        # Simulating for potential outcomes
        y1_x1 <- predict(fit,
                     newdata = x1_data,
                     se.fit = TRUE)
        y1_x0 <- predict(fit,
                     newdata = x0_data,
                     se.fit = TRUE)
        
        preds.fit <- y1_x1$fit - y1_x0$fit
        se1 <- sqrt(mean(y1_x1$se.fit^2))
        se0 <- sqrt(mean(y1_x0$se.fit^2))
        se <- max(se1, se0)
        
        # X1 <- model.matrix(delete.response(terms(fit)), x1_data)
        # X0 <- model.matrix(delete.response(terms(fit)), x0_data)
        # d  <- colMeans(X1 - X0)
        # est <- drop(d %*% coef(fit))
        # se  <- sqrt(drop(t(d) %*% vcov(fit) %*% d))
        
    list(
        y1_x0 = y1_x0,
        y1_x1 = y1_x1,
        mean = mean(preds.fit),
        se   = se)})
    
    estimates <- map_dbl(estimate_per_set, "mean")
    std_errors <- map_dbl(estimate_per_set, "se")
    return(list(adjustment_set = adj_sets,
                estimate = estimates,
                std_error = std_errors))
}


#' Estimate a causal contrast under every valid adjustment set
#'
#' Returns, as before:
#'   $adjustment_set  list of adjustment sets (dagitty), after filtering
#'   $estimate        numeric vector, one per set, same order
#'   $std_error       numeric vector, one per set, same order
#'
#' All valid adjustment sets identify the same estimand, so under a CORRECT
#' graph the spread across sets is noise. Under a deliberately misspecified
#' graph it is a measure of structural disagreement - but a lower bound on
#' error, not error itself: sets can agree closely and all be biased the same
#' way. Compare against the known truth as well as across sets.

causal_estimate_over_all_adj_sets <- function(
    dag, data,
    latent_nodes = NULL,
    exposure     = "S.Kn",
    outcome      = "Grade",
    x1_val       = 1,
    x0_val       = 0,
    type         = "all",
    max_sets     = Inf) {

  null_out <- list(adjustment_set = NULL, estimate = NULL, std_error = NULL)

  if (is.null(dag)) return(null_out)          # was missing its return()

  stopifnot(exposure %in% names(data), outcome %in% names(data))

  adj_sets <- dagitty::adjustmentSets(
    dag, exposure = exposure, outcome = outcome,
    type = type, max.results = max_sets
  )

  if (!is.null(latent_nodes))
    adj_sets <- purrr::keep(adj_sets, \(s) !any(s %in% latent_nodes))

  ## covariates must actually be measured, else lm() errors
  adj_sets <- purrr::keep(adj_sets, \(s) all(s %in% names(data)))

  if (length(adj_sets) == 0) {
    warning("no usable adjustment set for ", exposure, " -> ", outcome)
    return(list(adjustment_set = adj_sets,
                estimate = numeric(0), std_error = numeric(0)))
  }

  y        <- data[[outcome]]                 # was data[[1]]
  is_binom <- is.factor(y) || is.logical(y) ||
              (is.numeric(y) && all(y %in% c(0, 1), na.rm = TRUE))
  ex_lev   <- if (is.factor(data[[exposure]])) levels(data[[exposure]]) else NULL

  ## levels taken from the data, so x0_val = 0 can't become NA
  set_exposure <- function(d, v) {
    d[[exposure]] <- if (is.null(ex_lev)) v else factor(v, levels = ex_lev)
    d
  }

  estimate_per_set <- purrr::map(adj_sets, function(adj_set) {

    rhs <- if (length(adj_set) == 0) exposure
           else paste(c(exposure, adj_set), collapse = " + ")
    fit <- if (is_binom)
             stats::glm(stats::as.formula(paste(outcome, "~", rhs)),
                        data = data, family = stats::binomial)
           else
             stats::lm(stats::as.formula(paste(outcome, "~", rhs)), data = data)

    tt <- stats::delete.response(stats::terms(fit))
    X1 <- stats::model.matrix(tt, set_exposure(data, x1_val), xlev = fit$xlevels)
    X0 <- stats::model.matrix(tt, set_exposure(data, x0_val), xlev = fit$xlevels)

    b <- stats::coef(fit)
    V <- stats::vcov(fit)

    if (is_binom) {
      ## response scale (risk difference), delta-method SE
      p1   <- stats::plogis(drop(X1 %*% b))
      p0   <- stats::plogis(drop(X0 %*% b))
      est  <- mean(p1 - p0)
      grad <- colMeans(X1 * (p1 * (1 - p1)) - X0 * (p0 * (1 - p0)))
    } else {
      grad <- colMeans(X1 - X0)
      est  <- drop(grad %*% b)
    }

    ## SE of the contrast: both arms come from one fit, so the covariance
    ## between them must be kept, not dropped as in max(se1, se0)
    list(mean = est,
         se   = sqrt(drop(t(grad) %*% V %*% grad)))
  })

  list(adjustment_set = adj_sets,
       estimate       = purrr::map_dbl(estimate_per_set, "mean"),
       std_error      = purrr::map_dbl(estimate_per_set, "se"))
}



generate_node_group_causal_estimate_comparisons <- function(
        Model = "M5",
        Beta = 0.4, # strength of signal along paths
        # might need to change for different models depending on present nodes
        exposure = "S.Kn", outcome = "Grade",
        use_potential_observable_proxies = TRUE,
        x1_val = 1, x0_val = 0, # should be 1 and 0 for continuous
        # for placiing proxy nodes
        x_p_nudge = 0, y_p_nudge = -2
) {
    ground_truth_dag <- 
        CDAGs[[Model]] |> 
        tidygraph_to_dagitty()
    obs_nds = CDAGs[[Model]]$observable 
    proxy_nds = CDAGs[[Model]] |> 
        activate(nodes) |> 
        as_tibble() |> 
        filter(measurability_node_group == "Potentially observable") |> 
        pull(name)
    
    # Ground truth uses full simmed data
    est <- partial(causal_estimate_over_all_adj_sets,
                   exposure = exposure, outcome = outcome,
                   x1_val = x1_val, x0_val = x0_val)
    
    if (use_potential_observable_proxies) {
        bn_list <- bn_fit_simulations(
            dag = ground_truth_dag, 
            observable_nodes = obs_nds,
            proxy_nodes = proxy_nds,
            Beta = Beta)
        obs_data <- bn_list$simdata.opo
        learned_dag <- bn_list$bn.DAG.opo
    } else {
        bn_list <- bn_fit_simulations(
            dag = ground_truth_dag, 
            observable_nodes = obs_nds,
            Beta = Beta)
        obs_data <- bn_list$simdata.o
        learned_dag <- bn_list$bn.DAG.o
    }

    ground_truth <- est(
        dag = ground_truth_dag, 
        data = bn_list$simdata)
    
    expert_dag <- est(
        dag = ground_truth_dag, 
        data = obs_data,
        latent_nodes = setdiff(
            names(ground_truth_dag),
            names(obs_data)))
    
    learned_dag <- est(
        dag = learned_dag, 
        data = obs_data,
        latent_nodes = setdiff(
            names(learned_dag),
            names(obs_data)))

    estimates <- tibble(
        Model = Model,
        DAG = c(
            "Ground Truth", 
            "Expert DAG",
            "Learned DAG") |> 
            factor(levels = c(
                "Ground Truth", 
                "Expert DAG", 
                "Learned DAG")),
        exposure = exposure, 
        outcome = outcome,
        adjustment_sets = list(
            ground_truth$adjustment_set,
            expert_dag$adjustment_set,
            learned_dag$adjustment_set
        ),
        estimate_mean = c(
            mean(ground_truth$estimate),
            mean(expert_dag$estimate),
            mean(learned_dag$estimate)
        ),
        estimate_std_errors = c(
            mean(ground_truth$std_error),
            mean(expert_dag$std_error),
            mean(learned_dag$std_error)
        ),
        estimate_sd = c(
            sd(ground_truth$estimate),
            sd(expert_dag$estimate),
            sd(learned_dag$estimate)
        )
    )
    message(str_c("Simulation complete for ", Model))
    
    return( # Monster list of everything we need
        list(
            Model = Model,
            DAG = ground_truth_dag,
            fitted_BN = bn_list,
            estimates = estimates
        )
    )
}
```

### Model simulations with `dagitty` and `bnlearn`

```{r simulating-cluster-models}
if (rerun_simulations) {
    simulations_ng_cluster_models <- 
        c("M1", "M2", "M3", "M4", "M5", "M6", "M7", "M8") |> 
        set_names() |> 
        imap(
            \(M, n)
            generate_node_group_causal_estimate_comparisons(
                Model = M,
                Beta = sim_signal_strength[["typical"]]
            )
        )
    saveRDS(simulations_ng_cluster_models, "cache/model_simulations.RDS")
} else {
    simulations_ng_cluster_models <- readRDS("cache/model_simulations.RDS")
}

```

```{r all-CDAG-estimates}
all_ng_estimates <-
    map(
        simulations_ng_cluster_models,
        \(s)
           s$estimates
        ) |> 
    list_rbind() |> 
    group_by(Model) |> 
    mutate(
        gt_est_diff = estimate_mean - estimate_mean[DAG == "Ground Truth"] )|> 
    ungroup() |> 
    mutate(error = estimate_std_errors + replace_na(estimate_sd, 0)) |> 
    select(Model, DAG, estimate_mean, error, gt_est_diff, everything())
```









### Structure learning using ICD


ICD (Iterative Causal Discovery) learns a Partial Ancestry Graph (PAG), an equivalence class of Maximal Ancestry Graphs (MAGs),
allowing for latent confounders and selection bias. This code is taken from https://github.com/IntelLabs/causality-lab/blob/main/notebooks/causal_discovery_with_latent_confounders.ipynb, with some AI translation into here. 
It is implemented only in Intel Labs' `causality-lab` (Python), so this document runs it through `reticulate` and brings the learned graph back into R as an adjacency matrix.

#### One-time environment setup

Run this once in the console, not on every render.

```{r}
#| label: install
#| eval: false

library(reticulate)
virtualenv_create("icd", packages = c("numpy", "scipy"))
virtualenv_install("icd", "git+https://github.com/IntelLabs/causality-lab.git")
```

`causality-lab` is not on PyPI, so it installs from the repo. If `git+https`
is blocked, download the zip and point pip at the unpacked folder instead:

```{r}
#| label: install-fallback
#| eval: false
virtualenv_install("icd", "C:/path/to/causality-lab-main")
```

#### Setup

```{r}
#| label: setup
library(reticulate)
use_virtualenv("icd", required = TRUE)
```

```{python}
#| label: py-imports
import numpy as np
from causal_discovery_algs import LearnStructICD, LearnStructFCI
from causal_discovery_utils.cond_indep_tests import CondIndepParCorr
```


#### Data - batched

`sim_list` is a **named** list of numeric matrices, names `M1`–`M8`. Rows are
records, columns are variables. Models may have different numbers of columns —
the node set is derived per matrix, so a shared vocabulary isn't required.

```{r}
#| label: data

sim_list <- simulations_ng_cluster_models |> 
    map(
        \(sim)
        sim$fitted_BN$simdata.opo
    )

stopifnot(is.list(sim_list), !is.null(names(sim_list)))

sim_list <- lapply(sim_list, function(m) {
  m <- as.matrix(m)
  storage.mode(m) <- "double"
  m
})

# keep column names on the R side; Python works in 0-based indices only
var_names <- lapply(sim_list, function(m) {
  nm <- colnames(m)
  if (is.null(nm)) paste0("V", seq_len(ncol(m))) else nm
})

alpha <- 0.01

data.frame(
  model   = names(sim_list),
  records = vapply(sim_list, nrow, integer(1)),
  nodes   = vapply(sim_list, ncol, integer(1))
)
```


## Learn

The loop runs inside Python so each matrix crosses the boundary once. A failure
on one model is caught and recorded rather than aborting the render.

```{python}
#| label: py-helpers
import numpy as np
from causal_discovery_algs import LearnStructICD, LearnStructFCI
from causal_discovery_utils.cond_indep_tests import CondIndepParCorr


def learn_one(mat, alpha=0.01, run_fci=True):
    dataset = np.asarray(mat, dtype=float)
    if dataset.ndim != 2 or dataset.shape[1] < 3:
        raise ValueError(f"need a 2-D matrix with >=3 columns, got {dataset.shape}")

    obs = set(range(dataset.shape[1]))          # 0-based, matches column order

    ci = CondIndepParCorr(dataset=dataset, threshold=alpha,
                          count_tests=True, use_cache=True)
    icd = LearnStructICD(obs, ci)
    icd.learn_structure()

    out = {"adj_icd": icd.graph.get_adj_mat(),
           "tests_icd": ci.test_counter,
           "error": None}

    if run_fci:
        ci2 = CondIndepParCorr(dataset=dataset, threshold=alpha,
                               count_tests=True, use_cache=True)
        fci = LearnStructFCI(obs, ci2)
        fci.learn_structure()
        out["adj_fci"] = fci.graph.get_adj_mat()
        out["tests_fci"] = ci2.test_counter

    return out


def learn_many(mats, alpha=0.01, run_fci=True):
    res = {}
    for name, mat in mats.items():
        try:
            res[name] = learn_one(mat, alpha, run_fci)
        except Exception as e:
            res[name] = {"adj_icd": None, "tests_icd": None,
                         "error": f"{type(e).__name__}: {e}"}
    return res

```

```{python}
#| label: py-run
results = learn_many(r.sim_list, alpha=r.alpha, run_fci=True)
```

## Back into R

```{r}
#| label: pull-back
res <- py$results
res <- res[names(sim_list)]   # dicts don't guarantee order; restore M1..M8

failed <- vapply(res, function(x) !is.null(x$error), logical(1))
if (any(failed)) {
  message("failed: ", paste(names(res)[failed], collapse = ", "))
  print(vapply(res[failed], function(x) x$error, character(1)))
}

adj_icd <- Map(function(x, nm) {
  if (is.null(x$adj_icd)) return(NULL)
  a <- x$adj_icd
  dimnames(a) <- list(nm, nm)
  a
}, res, var_names)

adj_fci <- Map(function(x, nm) {
  if (is.null(x$adj_fci)) return(NULL)
  a <- x$adj_fci
  dimnames(a) <- list(nm, nm)
  a
}, res, var_names)
```

```{r}
#| label: edge-list
pag_edges <- function(adj, nodes = colnames(adj)) {
  left  <- c("1" = "o", "2" = "<", "3" = "-")   # mark at the a-end, mirrored
  right <- c("1" = "o", "2" = ">", "3" = "-")   # mark at the b-end
  p <- nrow(adj)
  out <- list()

  for (a in seq_len(p - 1)) {
    for (b in (a + 1):p) {
      if (adj[a, b] == 0 && adj[b, a] == 0) next
      out[[length(out) + 1]] <- data.frame(
        from      = nodes[a],
        to        = nodes[b],
        mark_from = left[[as.character(adj[b, a])]],
        mark_to   = right[[as.character(adj[a, b])]],
        stringsAsFactors = FALSE
      )
    }
  }

  if (length(out) == 0) {
    return(data.frame(from = character(), to = character(),
                      mark_from = character(), mark_to = character(),
                      edge = character()))
  }

  res <- do.call(rbind, out)
  res$edge <- paste0(res$mark_from, "-", res$mark_to)
  res
}

edges_icd <- lapply(adj_icd[!vapply(adj_icd, is.null, logical(1))], pag_edges)

# one long frame across models, if that suits the downstream comparison better
edges_all <- do.call(rbind, Map(function(e, nm) {
  if (nrow(e) == 0) return(NULL)
  cbind(model = nm, e)
}, edges_icd, names(edges_icd)))
rownames(edges_all) <- NULL

head(edges_all, 20)
```

## Summary across models

```{r}
#| label: summary
edge_summary <- function(e) {
  types <- c("o-o", "o->", "<-o", "-->", "<--", "<->", "---")
  counts <- table(factor(e$edge, levels = types))
  c(n_edges = nrow(e), counts)
}

summ <- t(vapply(edges_icd, edge_summary, numeric(8)))
summ <- as.data.frame(summ)
summ$total_tests <- vapply(res[rownames(summ)],
                           function(x) sum(unlist(x$tests_icd)), numeric(1))
summ
```

Note `<-o` and `<--` are the same edge types as `o->` and `-->` with the
endpoints swapped; they appear separately here only because `pag_edges()` fixes
the column order by node index rather than by orientation.

## CI tests per conditioning set size

`test_counter` runs to the largest conditioning set reached, which differs by
model, so pad before binding.

```{r}
#| label: test-counts
pad <- function(x, n) c(x, rep(0, n - length(x)))

tests_icd <- lapply(res, function(x) unlist(x$tests_icd))
tests_icd <- tests_icd[!vapply(tests_icd, is.null, logical(1))]
maxlen <- max(vapply(tests_icd, length, integer(1)))

tests_mat <- t(vapply(tests_icd, pad, numeric(maxlen), n = maxlen))
colnames(tests_mat) <- seq_len(maxlen) - 1L
tests_mat

matplot(seq_len(maxlen) - 1L, t(tests_mat), type = "b", pch = 1, lty = 1,
        xlab = "conditioning set size", ylab = "number of CI tests")
legend("topright", rownames(tests_mat), col = seq_len(nrow(tests_mat)),
       lty = 1, bty = "n", cex = 0.8)
```

``` {r}
pag_to_dagitty <- function(e) {
  f <- function(a, b, mf, mt) {
    switch(paste0(mf, mt),
      "oo" = sprintf("%s @-@ %s", a, b),
      "o>" = sprintf("%s @-> %s", a, b),
      "->" = sprintf("%s -> %s",  a, b),
      "<o" = sprintf("%s @-> %s", b, a),   # flip so the circle leads
      "<-" = sprintf("%s -> %s",  b, a),
      "<>" = sprintf("%s <-> %s", a, b),
      "o-" = sprintf("%s @-- %s", a, b),
      "-o" = sprintf("%s @-- %s", b, a),
      "--" = sprintf("%s -- %s",  a, b),
      stop("unhandled edge: ", mf, mt)
    )
  }
  stmts <- mapply(f, e$from, e$to, e$mark_from, e$mark_to, USE.NAMES = FALSE)
  paste0("pag { ", paste(stmts, collapse = "; "), " }") |> 
      dagitty()
}

icd_PAGs <- 
    edges_icd |> 
    map(pag_to_dagitty)

```

Using the ICD PAGs to run simulations

``` {r}
sim_from_pag <- function(Model = "M6") {
    causal_estimate_over_all_adj_sets(
        dag = icd_PAGs[[Model]], 
        data = simulations_ng_cluster_models[[Model]]$fitted_BN$simdata.opo
        ) 
}

simulations_ng_PAGs <- 
    str_c("M",1:8) |> 
    set_names() |> 
    map(sim_from_pag)

PAG_ng_estimates <-
    map(
        simulations_ng_PAGs,
        \(s)
           tibble(
               DAG = "Learned PAG",
               exposure = "S.Kn",
               outcome = "Grade",
               adjustment_sets = list(s$adjustment_set),
               estimate_mean = mean(s$estimate),
               error = mean(s$std_error)
           )
        ) |> 
    list_rbind(names_to = "Model") 


all_simulation_estimates <- 
    bind_rows(all_ng_estimates, PAG_ng_estimates) 

```
















## Simulation comparisons

Between Ground Truth, Expert DAGs, and Learned DAGs from observable
data.

```{r plotting-cluster-model-simulation-results}
dodge_amt <- 0.1

ground_truth_estimate_mean_estimate <- 
    all_ng_estimates |> 
    filter(DAG == "Ground Truth") |> 
    filter(!is.nan(estimate_mean)) |> 
    pull(estimate_mean) |> 
    mean()

ground_truth_error_estimate <- 
    ( # the largest base error
        all_ng_estimates |> 
    filter(DAG == "Ground Truth") |> 
    filter(!is.nan(error)) |> 
    pull(error) |> 
    max())


est_to_plot <- 
    all_simulation_estimates |>
    mutate(estimated_ground_truth = is.nan(estimate_mean)) |> 
    mutate(
        estimate_mean = if_else(is.nan(estimate_mean) & DAG == "Ground Truth",
                                ground_truth_estimate_mean_estimate,
                                estimate_mean),
        error = if_else(is.nan(error) & DAG == "Ground Truth",
                                ground_truth_error_estimate,
                                error)
    ) |> 
    group_by(Model) |> 
    mutate(
        gt_est_diff = if_else(
            is.nan(gt_est_diff) | is.na(gt_est_diff),
            estimate_mean - estimate_mean[DAG == "Ground Truth"],
            gt_est_diff
        )
    ) |> 
    ungroup() |> 
    mutate(ogDAG = DAG) |> 
    mutate(DAG = if_else(estimated_ground_truth & DAG == "Ground Truth",
                         "Ground Truth (est)",
                         DAG) |> 
               factor(levels = c(
                   "Ground Truth",
                   "Ground Truth (est)",
                   "Expert DAG",
                   "Learned DAG",
                   "Learned PAG"
               )))
    
pal_cols <- palette.colors()
pal_cols <- c(pal_cols[1], "#777777", pal_cols[2], pal_cols[3], pal_cols[4])
```

```{r plotting-individual-simulation-results}
# Would be nice to replicate this, but group 
# models into those identifiable and those not
# Maybe change this to abs effect?
est_to_plot |> 
    filter(DAG != "Potentially Observable") |> 
    mutate(DAG = if_else(DAG == "Observable", "Learned DAG", 
                         DAG) |> 
               factor(levels = c(
                   "Ground Truth",
                   "Ground Truth (est)",
                   "Expert DAG",
                   "Learned DAG",
                   "Learned PAG"
               ))) |> 
    mutate(
        y_pos = as.numeric(factor(Model)) + 
            (as.numeric(factor(ogDAG)) - mean(as.numeric(factor(ogDAG)))) * dodge_amt
    ) |> 
    ggplot(aes(color = DAG, y = y_pos)) + 
    # geom_pointinterval(
    #     aes(x = gt_est_diff, 
    #         xmin = gt_est_diff - 2 * error, 
    #         xmax = gt_est_diff + 2 * error),
    #        alpha = 0.3, size = 1
    # ) +
    geom_pointinterval(
        aes(x = estimate_mean,
            xmin = estimate_mean - 1 *error,
            xmax = estimate_mean + 1 *error),
           alpha = 0.8, size = 3
    ) +
    scale_y_continuous(
        breaks = seq_along(levels(factor(est_to_plot$Model))),
        labels = levels(factor(est_to_plot$Model)),
        name = "Model"
    ) +
    annotate("label", x = -0.0, y = c(2.2,3.2,5.2,6.2,8.2), label = "?",
             color = pal_cols[3], alpha = 0.4) +
    scale_colour_discrete(palette = pal_cols) +
    xlab("Estimate") +
    theme_minimal()
```

```{r plotting-simulation-results-split-by-idXY}
est_to_plot_grouped <- 
    est_to_plot |> 
    group_by(Model) |> 
    mutate(model_group = case_when(
        any(DAG == "Ground Truth (est)") ~ "Unknowable models:\n",
        is.nan(estimate_mean[DAG == "Expert DAG"]) ~ "Non-identifiable models\n(from potentially\nobservable data):\n", 
        TRUE ~ "Identifiable models\n(from potentially\nobservable data):\n"
        )) |> 
    ungroup() |> 
    filter(DAG %in% c(
        "Learned DAG",
        "Learned PAG",
        "Ground Truth",
        "Expert DAG")) |>
    group_by(model_group) |> 
    mutate(model_group_list = str_c(
        unique(model_group), 
        str_c(unique(sort(Model)) , collapse = ", "))) |> 
    ungroup() |> 
    mutate(
        y_pos = as.numeric(factor(model_group_list)) + 
            (as.numeric(factor(ogDAG)) - mean(as.numeric(factor(ogDAG)))) * dodge_amt
    )

p.est.sims <- 
    est_to_plot_grouped |> 
    ggplot(aes(color = DAG, y = y_pos)) + 
    geom_pointinterval(
        aes(x = gt_est_diff,
            xmin = gt_est_diff - 1 *error,
            xmax = gt_est_diff + 1 *error),
           alpha = 0.6, size = 5
    ) +
    scale_y_continuous(
        breaks = seq_along(levels(factor(est_to_plot_grouped$model_group_list))),
        labels = levels(factor(est_to_plot_grouped$model_group_list)),
        name = "Model estimates"
    ) +
    annotate("label", x = -0.0, y = c(1.8), label = "?",
             color = pal_cols[3], alpha = 0.4) +
    annotate("label", x = -0.0, y = c(2.8), label = "?",
             color = pal_cols[3], alpha = 0.4) +
    annotate("label", x = -0.0, y = c(3), label = "?",
             color = "grey20", alpha = 0.4) +
    scale_colour_discrete(palette = palette.colors()) +
    xlab("Deviation from ground truth estimate (SD)") +
    theme_minimal() +
    theme(legend.position = "bottom")

p.est.sims


p.est.sims.2 <- 
    est_to_plot_grouped |>
    filter(!str_detect(model_group_list, "Unk")) |> 
    ggplot(aes(color = DAG, y = y_pos)) + 
    geom_pointinterval(
        aes(x = gt_est_diff,
            xmin = gt_est_diff - 1 *error,
            xmax = gt_est_diff + 1 *error),
           alpha = 0.6, size = 5
    ) +
    scale_y_continuous(
        breaks = seq_along(levels(factor(est_to_plot_grouped$model_group_list))),
        labels = levels(factor(est_to_plot_grouped$model_group_list)),
        name = "Model estimates"
    ) +
    annotate("label", x = -0.0, y = c(1.85), label = "?",
             color = pal_cols[3], alpha = 0.4) +
    scale_colour_discrete(palette = palette.colors()) +
    xlab("Deviation from ground truth estimate (SD)") +
    theme_minimal() +
    theme(legend.position = "bottom")

p.est.sims.2

ggsave("figures/FIG--simulation-results.png", p.est.sims.2,
       width = 10, height = 3)

```

```{r}
plot_bnlearn_coarse_comparison <- function(s_estimates) {
    p_orig <- s_estimates$DAG |> 
        dag_to_tidy_graph() %N>%
        mutate(name = str_remove(name, "_p$")) |>
        gcm_cluster_graph(code_node_group) |> 
        g_plot_coarse()
    
    x_limits <- layer_scales(p_orig)$x$get_limits()
    y_limits <- layer_scales(p_orig)$y$get_limits()
    
    p_opo <- s_estimates$fitted_BN$bn.opo |>
        bn_to_dagitty() |> 
        dag_to_tidy_graph() %N>%
        mutate(name = str_remove(name, "_p$")) |> 
        gcm_cluster_graph(code_node_group) |> 
        g_plot_coarse() +
        xlim(x_limits) +
        ylim(y_limits) 
    
    p_o <- s_estimates$fitted_BN$bn.o |>
        bn_to_dagitty() |> 
        dag_to_tidy_graph() %N>%
        mutate(name = str_remove(name, "_p$")) |> 
        gcm_cluster_graph(code_node_group) |> 
        g_plot_coarse() +
        xlim(x_limits) +
        ylim(y_limits) 
    
    return(list(
        p_orig = p_orig, 
        p_opo = p_opo, 
        p_o = p_o))
    
}

pcomp_m1 <-  plot_bnlearn_coarse_comparison(s_estimates = simulations_ng_cluster_models$M1)
pcomp_m1$p_orig / pcomp_m1$p_o +
    ggtitle("Model 1: Actual coarsened DAG (top)
            vs coarsened learned DAG from observable nodes")

pcomp_m2 <-  plot_bnlearn_coarse_comparison(s_estimates = simulations_ng_cluster_models$M2)
pcomp_m2$p_orig / pcomp_m2$p_o +
    ggtitle("Model 2: Actual coarsened DAG (top)
            vs coarsened learned DAG from observable nodes")

pcomp_m3 <-  plot_bnlearn_coarse_comparison(s_estimates = simulations_ng_cluster_models$M3)
pcomp_m3$p_orig / pcomp_m3$p_o + 
    ggtitle("Model 3: Actual coarsened DAG (top)
            vs coarsened learned DAG from observable nodes")

pcomp_m4 <-  plot_bnlearn_coarse_comparison(s_estimates = simulations_ng_cluster_models$M4)
pcomp_m4$p_orig / pcomp_m4$p_o + 
ggtitle("Model 4: Actual coarsened DAG (top)
            vs coarsened learned DAG from observable nodes")

pcomp_m5 <-  plot_bnlearn_coarse_comparison(
    s_estimates = simulations_ng_cluster_models$M5)
pcomp_m5$p_orig / pcomp_m5$p_o + 
    ggtitle("Model 5: Actual coarsened DAG (top)
            vs coarsened learned DAG from observable nodes")

pcomp_m6 <-  plot_bnlearn_coarse_comparison(
    s_estimates = simulations_ng_cluster_models$M6)
pcomp_m6$p_orig / pcomp_m6$p_o + 
ggtitle("Model 6: Actual coarsened DAG (top)
            vs coarsened learned DAG from observable nodes")

pcomp_m7 <-  plot_bnlearn_coarse_comparison(
    s_estimates = simulations_ng_cluster_models$M7)
pcomp_m7$p_orig / pcomp_m7$p_o + 
    ggtitle("Model 7: Actual coarsened DAG (top)
            vs coarsened learned DAG from observable nodes")

pcomp_m8 <-  plot_bnlearn_coarse_comparison(
    s_estimates = simulations_ng_cluster_models$M8)
pcomp_m8$p_orig / pcomp_m8$p_o + 
    ggtitle("Model 8: Actual coarsened DAG (top)
            vs coarsened learned DAG from observable nodes")

```

# Dissimilarity between all models

```{r combining-all-expert-and-learned-CDAGs}
learned_CDAGs <- 
    simulations_ng_cluster_models |> 
    set_names(nm = str_replace(names(simulations_ng_cluster_models), "M", "D")) |> 
    imap(
        \(est,m)
        est$fitted_BN$bn.opo |> 
            bn_to_dagitty() |> 
            dagitty_to_tidygraph() |> 
            activate(nodes) |> 
            mutate(name = str_remove(name, "_p$"))
    )

all_CDAGs <- 
    c(CDAGs, learned_CDAGs)
```

```{r L0-between-all-CDAGs}
if (recalculate_distance_matrices) {
    all_L0_d_matrix <- 
        compute_dist_matrix(all_CDAGs, L0_d, symmetric = TRUE)
    saveRDS(all_L0_d_matrix, "cache/all_L0_d_matrix.RDS")
} else {
    all_L0_d_matrix <- readRDS("cache/all_L0_d_matrix.RDS")
}

```

```{r L1-between-all-CDAGs}
if (recalculate_distance_matrices) {
    # Dist matrix
    all_L1_d_matrix <- 
        compute_dist_matrix(
            all_CDAGs, 
            .df = L1_d)
    saveRDS(all_L1_d_matrix, 
            "cache/all_L1_d_matrix.RDS")
} else {
    all_L1_d_matrix <- readRDS("cache/all_L1_d_matrix.RDS")
}

```

```{r L2-between-all-CDAGs}
# L2 distance between two models
if (recalculate_distance_matrices) {
    # Dist matrix
    all_L2_d_matrix <- 
        compute_dist_matrix(
            all_CDAGs,
            L2_d,
            symmetric = TRUE)
    saveRDS(all_L2_d_matrix, 
            "cache/all_L2_d_matrix.RDS")
} else {
    all_L2_d_matrix <- readRDS("cache/all_L2_d_matrix.RDS")
}
```

```{r tanglegram-L1-L2-clustering-all}
library(viridis)

L1_hclust_all <- hclust(as.dist(all_L1_d_matrix),
                        method = "average")
L2_hclust_all <- hclust(as.dist(all_L2_d_matrix),
                        method = "average")

dend_L12_all <- dendlist(
    L1_hclust_all |> as.dendrogram() |> highlight_branches_col(rev(mako(100))[5:100]),
    L2_hclust_all |> as.dendrogram() |> highlight_branches_col(rev(magma(100))[9:100])
)

tanglegram(dend_L12_all, main_left = "L1-consistency", main_right = "L2-consistency",
           sort = TRUE, edge.lwd = 3, lwd = 2,
           common_subtrees_color_lines = TRUE, 
           highlight_distinct_edges  = FALSE)
```


```{r L012-dist-clusters-across-all}

all_L12_dist_matrix <- 
    (all_L1_d_matrix + all_L2_d_matrix) / 2

L12_hclust_full <- hclust(
    as.dist(all_L12_dist_matrix),
    method = "average")

all_L012_dist_matrix <- 
    (all_L0_d_matrix + all_L1_d_matrix + all_L2_d_matrix) / 3

L012_hclust_full <- hclust(
    as.dist(all_L012_dist_matrix),
    method = "average")

```

## Grouping learned and expert CDAGs by similarity

```{r dendo-full}
dend <- 
    L012_hclust_full |> 
    as.dendrogram() 

lab  <- labels(dend)
is_M <- substr(lab, 1, 1) == "M"
is_D <- substr(lab, 1, 1) == "D"

dend <- set(dend, "by_labels_branches_lty",
            value = lab[is_D], TF_values = c(2, 1))

leaf_cols <- c(
    "#000000", "#E69F00", "#56B4E9", "#009E73",  "#0072B2", "#D55E00", "#CC79A7", "#999999","#F0E442")
num <- as.integer(sub("^\\D+", "", lab))
pal  <- setNames(leaf_cols, 1:8)
labels_colors(dend) <- pal[as.character(num)]

png("figures/FIG--CDAGs-all-dendrogram.png", 
    width = 10, height = 4, units = "in", res = 300)
dend |> 
    hang.dendrogram(hang = 0.1) |>  
    plot(axes = F, horiz = F) 


dev.off()

dend |> 
    hang.dendrogram(hang = 0.1) |>  
    plot(axes = F, horiz = F) 


```


```{r}
labels <- c(paste0("M", 1:8), paste0("D", 1:8))
nums <- as.integer(gsub("\\D", "", labels))

d_mat <- outer(nums, nums, function(a, b) as.numeric(a != b))
dimnames(d_mat) <- list(labels, labels)

d_pair <- as.dist(d_mat)

L_hclust_perfect_learn <- hclust(d_pair ,method = "complete")

dend_perfect <- L_hclust_perfect_learn |> 
    as.dendrogram()

labp  <- labels(dend_perfect)

nump <- as.integer(sub("^\\D+", "", labp))
palp  <- setNames(leaf_cols, 1:8)
labels_colors(dend_perfect) <- palp[as.character(nump)]

tanglegram(dend1 = dend_perfect,
           dend2 = dend,
           main_left = "Target learning", main_right = "L-dissimilarity",
           sort = TRUE, edge.lwd = 2, lwd = 2,
           common_subtrees_color_lines = TRUE, 
           highlight_distinct_edges  = F,
           axes = F,
           rank_branches = F
           )
```


# Appendices

## All models plotted

Highlighting all models in turn, original graph first and then CDAG.

```{r render-all-highlighted-fine-models}
if (export_and_save_every_model) {
    p_fine_show_and_save(1)
    p_coarse_show_and_save(1)
    
    p_fine_show_and_save(2)
    p_coarse_show_and_save(2)
    
    p_fine_show_and_save(3)
    p_coarse_show_and_save(3)
    
    p_fine_show_and_save(4)
    p_coarse_show_and_save(4)
    
    p_fine_show_and_save(5)
    p_coarse_show_and_save(5)
    
    p_fine_show_and_save(6)
    p_coarse_show_and_save(6)
    
    p_fine_show_and_save(7)
    p_coarse_show_and_save(7)
    
    p_fine_show_and_save(8)
    p_coarse_show_and_save(8)
} else {
    print("Models saved in 'figures' folder, names beginning with 'plt--'")
}
```

## Future work

- String diagrams as a next formalism step: maybe they offer cleaner
  graph surgery operations.

### Category theory of DAGs

https://arxiv.org/pdf/2201.06981 - follow up this paper On the
Equivalence of Causal Models: A Category-Theoretic Approach Jun Otsuka
JOTSUKA\@BUN.KYOTO-U.AC.JP Philosophy Dept., Kyoto University, Kyoto,
Japan. Hayato Saigo HARMONIAHAYATO\@GMAIL.COM Dept. of Bioscience,
Nagahama Institute of Bio-Science and Technology, Shiga, Japan.

Or from an earlier paper: \> \[DAG based\] causal reasoning relies
crucially on a delicate interplay between syntax and semantics, which is
often not made explicit in the literature. The syntactic object of
interest is the causal structure (e.g. a causal graph), which captures
something about our understanding of the world, and the mechanisms which
gave rise to some observed phenomena. The semantic object of interest is
the data: joint and conditional probability distributions on some
variables. Fixing a causal structure entails certain constraints on
which probability distributions can arise, hence it is natural to see
distributions satisfying those constraints as models of the syntax.

**What if the semantic object is the expert?**

### What is the case for using string diagrams?

There seems to be a good case for using string diagrams after the
initial run of GCMing with domain experts. They are an incremental step
in the formalisation process and have other affordances as well.

> Is it possible that graph merging, coarsening, or L_i distance metrics
> are simpler in causal string diagrams than they are in causal DAGs?

### Write functions to help node coding with causal abstraction theory

- Given a graph, suggest L1-L2 permissible abstractions (to some degree)
- Given a coarsening, check for L1-L2 validity
- Would be good to have some workflow documented

### A representation of Hierarchical DAGs

Given a a mix of DAGs / CDAGs / coarsenings is there a way to represent both the fine grained structure and the coarse grained structure at the same time? For instance, an initial method might be having the fine-grain nodes sitting within a larger cluster node. However this still needs to have rules for how causal flow / influence moves between the lower and higher abstraction levels. 

For instance, if we are given these two DAGs: 

``` {r}
lwrDAG <- vnt.tdags$Fine1a |> 
    bind_nodes(tibble(name = "T", x = 1, y = -0.2, group = "T")) |> 
    bind_edges(
        tibble(
            from = c(1,6), 
            to = c(6,3), 
            type = rep("->",2))
    )

hrDAG <- vnt.tdags$Cluster1W
    
lwrDAG |> 
    ggraph(layout = "manual", x = x, y = y)+
        geom_edge_fan(
            arrow = arrow(length = unit(4, 'mm'), type = "closed"),
            start_cap = circle(6, 'mm'),
            end_cap = circle(6, 'mm')) +
        geom_node_point(aes(colour = group), size = 6) +
        geom_node_text(aes(label = name), colour = "white", alpha = 0.9) +
        scale_colour_manual(
            values = first_letter_pal,
            guide = "none") +
        xlim(c(-0.1,2.1)) +
        theme_graph() +
    ggtitle("A model with fine-grained descriptions")
```

``` {r}
hrDAG |> 
    plot_vnt_dag() +
    ggtitle("A cluster model that is in agreement, but broader scope")
```

> Can we represent both these fully? We see that $Z \rightarrow W$, but do not know whether $Z_1 \rightarrow W$, $Z_2 \rightarrow W$, or both. There are similar challenges examining the path $W \rightarrow X$. 

``` {r}
mrgDag <- 
    gcm_merge(lwrDAG, hrDAG) |> 
    activate(nodes) |> 
    mutate(
        x = case_when(
            name == "X1" ~ x + 0.1,
            TRUE ~ x
        ),
        y = case_when(
            name == "Z" ~ y + 0.2,
            name == "X1" ~ y + 0.2,
            TRUE ~ y
        )) 

mrgDag |> 
    ggraph(layout = "manual", x = x, y = y)+
        geom_edge_fan(
            arrow = arrow(length = unit(4, 'mm'), type = "closed"),
            start_cap = circle(6, 'mm'),
            end_cap = circle(6, 'mm')) +
        geom_node_point(aes(colour = group), size = 6) +
        geom_node_text(aes(label = name), colour = "white", alpha = 0.9) +
        scale_colour_manual(
            values = first_letter_pal,
            guide = "none") +
        theme_graph()

```

We can see that even if both models are regarded with high confidence we have some unanswered questions. For instance, is there a path from $W$ to $T$?

A proposal: 

- Draw both high and low level abstractions on the same graph. Clusters *might* be drawn larger for clarity. 
- Connect low level (finer grained) abstractions with their higher level abstraction via a `-<` edge, which is a little like a less than or subset sign, with the `-` end to the finer grained, lower level abstraction. 
- Influence flows as normal from lower to higher (left to right along `-<`). A change in the sub-component of a thing changes the whole. 
- The challenge is to imagine what flows from the higher to lower level path (left to right along `>-` paths). 

The last point is the crux of the matter. Just because we influence the higher level construct does not tell us *which* sub-component is influenced. This (if no other information is available) would need to be a *weaker* flow of information, perhaps distributed across the sub-components (and maybe a noise parameter or something to represent that there are 'all the other' sub-components in the model). This could maybe work in a similar way to "conditioning on a decsendant" that partially controls for a variable. 
